### 2021-12-07 SampleExtendedAbstract_SDSS2020.tex
Adding Abstract

% File SDSS2020_SampleExtendedAbstract.tex
\documentclass[10pt]{article}
\usepackage{sdss2020} % Uses Times Roman font (either newtx or times package)
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{algorithm, algorithmic}  
\usepackage{graphicx}

\title{Symposium on Data Science and Statistics (SDSS 2022) \\
Submission and Formatting Instructions for Extended Abstracts}

\author{
  Denise Bradford \\
  University of Nebraska - Lincoln \\
  Lincoln, Nebraska \\
  {\tt denise.bradford@huskers.unl.edu} \\\And
  Susan VanderPlas \\
  University of Nebraska - Lincoln \\
  Lincoln, Nebraska \\
  {\tt susan.vanderplas@unl.edu} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in settings when the analysis is a novice to data. We will demonstrate the surrounding. Our framework will suggest a number of research directions to better support dashboard design, implementation and use for the every day analyst. 
\end{abstract}

{\bf Keywords:} Interactive Dashboards, Exploratory Data Analysis (EDA), Design Space(??)

\section{Introduction}
With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{Fisher}. A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making. Stephen Few defines a dashboard as:
\begin{quotation}
\small A visual display of the most important information needed to achieve one or more objectives, consolidated and arranged on a single screen so the information can be monitored at a glance. \cite{Few}
\end{quotation}

\section{Methods}
Using an exploratory methodology, we derived a design space consisting of 3-4 visual and functional aspects of dashboards with statistical exploratory data analysis. see Section~\ref{info-on-sections}.

\subsection{Purpose}
The intended use of a dashboard for a novice user drives the choices in the visual and functional design allowances. The factors here capture the roles of each dashboard in the process of analysis and communication. We find that the purpose of a dashboard has been substantially expanded from the operational dashboard to capture decision-making at higher levels, and may not be supportive in decision-making.

{\bf Decision Support} (Strategic, Tactical, Operational): The decision support dimension reflects on the sorts of actionable decisions that dashboards are designed to support.

{\bf Communication and Learning:} We encountered several examples of dashboards that did not solicit decision-making on any temporal scale.

\subsection{Audience}
The visual and functional aspects of our dashboards typically reflect the intended audience, which has a wide-range of characteristics in their domain and visualization aspects of a dashboard and their experience with data.

\section{Related Work}
From previous research has explained how visual analytics addresses the issue of information overload and enables analysts to transform raw data into salient information and knowledge

\section{Data/Results}

We encourage the selective use of tables and/or figures as appropriate.

\subsection{Graphics and Tables}

Place figures and tables in the
paper near where they are first discussed, rather than at the end, if
possible.  It is acceptable that wide illustrations
may run across both columns.  The use of color
is allowed in extended abstract graphics.

\section{Discussion/Conclusions}

\subsection{Appendices and acknowledgements}

Please avoid the use of appendices and acknowledgements 
as separate sections.  Any pertinent material should be included in the
body of the extended abstract rather than in an appendix.  If 
acknowledgements are necessary then they should be made using footnotes.

If you wish to provide links to accompanying code, this is allowable because
SDSS encourages best practices for reproducible research.  However, all
conclusions and exposition relevant to the refereeing process should appear in
the body of the extended abstract; furthermore, we cannot guarantee that 
code will be checked or even consulted during the refereeing process.

\subsection{Refereeing Decisions}
It is expected that accepted papers will be presented in person at SDSS in Pittsburgh 
from June 4 to 6, 2020.
Accepted papers' authors will be notified in plenty of time to make registration and travel 
arrangements and for papers not accepted to submit an e-poster to the 
conference if they so choose.

\bibliographystyle{sdss2020} % Please do not change the bibliography style
\bibliography{SampleReferencesForExtendedAbstract}

\end{document}

### 2021-12-07 SampleExtendedAbstract_SDSS2020.Rnw
Rnw file instead of tex file
code chunk
Update SampleExtendedAbstract_SDSS2020.Rnw

--- 2021-12-07 SampleExtendedAbstract_SDSS2020.tex
@@
 \begin{document}
+
 \maketitle
@@
 
-\section{Introduction}
+<<setup, include = F>>=
+# This is a code chunk
+@
+
+\section{Research Problem}
 With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{Fisher}.
@@
 \end{quotation}
+Dashboards are used to improve performance, improve strategic decision making, track processes effectiveness, minimize data complexity and communicate organizations' data-driven decision making \cite{...}.
+Some communities continue to thrive as they lose population because they adapt and stay focused on quality of life, community services, and investing in the future.
+This is what we call rural smart shrinkage.
+Our research team is developing strategies and sharing examples of successful shrink-smart towns to help similar communities to improve quality of life for their residents\cite{SCC}.
+Our work focuses on developing and challenges novice analysts in advanced statistical concepts and data visualizations to help make decisions.
+Specifically, what data can be used to help make decisions, what strategies overcome the challenges of those analysts feel empowered to make data-driven decisions.
+Our ultimate aim is to help inform Iowa Rural Shrink Smart communities by using effective dashboards to help those towns build the Quality of Life (QoL) Measures.
 
-\section{Methods}
-Using an exploratory methodology, we derived a design space consisting of 3-4 visual and functional aspects of dashboards with statistical exploratory data analysis.
-see Section~\ref{info-on-sections}.
+\section{Data Description}
+Using an exploratory methodology, we collected open source data from the Iowa Data Gov website to understand the
 
-\subsection{Purpose}
+we derived a design space consisting of 3-4 visual and functional aspects of dashboards with statistical exploratory data analysis. see Section~\ref{info-on-sections}.
+
+\subsection{Challenges}
+While data
+
 The intended use of a dashboard for a novice user drives the choices in the visual and functional design allowances.

### 2021-12-08 SCCAbstract_SDSS2022-knitr.Rnw
Updating the folder with useful files with name changes

--- 2021-12-07 SampleExtendedAbstract_SDSS2020.Rnw
@@
 \documentclass[10pt]{article}
+\usepackage{newtxtext, newtxmath, times}
 \usepackage{sdss2020} % Uses Times Roman font (either newtx or times package)
@@
 
-\title{Symposium on Data Science and Statistics (SDSS 2022) \\
-Submission and Formatting Instructions for Extended Abstracts}
+%\title{Symposium on Data Science and Statistics (SDSS 2022) \\
+%Submission and Formatting Instructions for Extended Abstracts}
+
+\title{Exploring Rural Shrink Smart Through Guided Discovery Dashboards}
 
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
+
+<<setup, include=FALSE, cache = FALSE>>=
+library(knitr)
+opts_chunk$set(
+concordance=TRUE
+)
+@
 
@@
 
-{\bf Keywords:} Interactive Dashboards, Exploratory Data Analysis (EDA), Design Space(??)
-
-<<setup, include = F>>=
-# This is a code chunk
-@
+{\bf Keywords:} Interactive Dashboards, Exploratory Data Analysis (EDA), Design Space, Guided Discovery
 
@@
 
-\section{Data Description}
-Using an exploratory methodology, we collected open source data from the Iowa Data Gov website to understand the
+\section{Data Collection}
+Iowa Data were used to create the SCC dashboard.
+Iowa Data is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
+Iowa Data website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
+For the purpose of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details.
+These data are collected on all towns that are available.
+We will use statistical clustering methods will be used to determine the town classifications to help improve the rural towns.
 
-we derived a design space consisting of 3-4 visual and functional aspects of dashboards with statistical exploratory data analysis. see Section~\ref{info-on-sections}.
+\section{Data Challenges}
+Iowa Data website has an exhaustive and detailed data, but that data collected has missing information.
+This information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
+For example, many small towns do not have a hospital that will only cater to the community, but they will have access to a hospital in a neighboring town.
+The second type of missing data are when data should be available and are not currently being collected in the data but are available in a different data source.
+For example, our public school data does not actual represent public schools in bigger cities, suggesting that the closest Elementary school is 20 miles away, which we understood that to be untrue.
+This type of missing data is very typical in data collection, but could led our research in a poor direction.
 
-\subsection{Challenges}
-While data
+\section{Guiding Design Principles}
+Dashboards have a variety of challenges, our research will focus on users' challenges with understanding data and statistical relationships of dashboards.
+We investigate user interaction strategies that will not challenge the knowledge of the analyst in a variety of backgrounds.
+With little research on the adaptation of dashboards, we will particularly focus on the techniques that to solve interaction and information presentation challenges.
+In our research, we would like to use a guiding design principles to develop and design the following:
+\begin{itemize}
+\item Showing a town analyst with peers to help with changes that improved the QoL
+\item Dashboard design that will make the town the main focus without distracting the analysts with peer comparisons
+\item Creating a feedback loop that will, assist in the guided discovery from the analysts
+\end{itemize}
 
-The intended use of a dashboard for a novice user drives the choices in the visual and functional design allowances.
-The factors here capture the roles of each dashboard in the process of analysis and communication.
-We find that the purpose of a dashboard has been substantially expanded from the operational dashboard to capture decision-making at higher levels, and may not be supportive in decision-making.
+\section{Current Progress and Future Work}
 
-{\bf Decision Support} (Strategic, Tactical, Operational): The decision support dimension reflects on the sorts of actionable decisions that dashboards are designed to support.
+\subsection{Current Progress}
+Our research as stated above, we would like to produce and develop a guided discovery dashboard that will allow rural analysts the ability to access information that will help with the Quality of Life in their town.
+This dashboard will be used to help the analyst make data-driven decisions about making effective changes in their communities.
+Our dashboard will be used to
+\subsection{Future Work}
 
-{\bf Communication and Learning:} We encountered several examples of dashboards that did not solicit decision-making on any temporal scale.
-
-\subsection{Audience}
-The visual and functional aspects of our dashboards typically reflect the intended audience, which has a wide-range of characteristics in their domain and visualization aspects of a dashboard and their experience with data.
-
-\section{Related Work}
-From previous research has explained how visual analytics addresses the issue of information overload and enables analysts to transform raw data into salient information and knowledge
-
-\section{Data/Results}
-
-We encourage the selective use of tables and/or figures as appropriate.
-
-\subsection{Graphics and Tables}
-
-Place figures and tables in the
-paper near where they are first discussed, rather than at the end, if
-possible.  It is acceptable that wide illustrations
-may run across both columns.  The use of color
-is allowed in extended abstract graphics.
-
-\section{Discussion/Conclusions}
-
-\subsection{Appendices and acknowledgements}
-
-Please avoid the use of appendices and acknowledgements 
-as separate sections.  Any pertinent material should be included in the
-body of the extended abstract rather than in an appendix.  If 
-acknowledgements are necessary then they should be made using footnotes.
-
-If you wish to provide links to accompanying code, this is allowable because
-SDSS encourages best practices for reproducible research.  However, all
-conclusions and exposition relevant to the refereeing process should appear in
-the body of the extended abstract; furthermore, we cannot guarantee that 
-code will be checked or even consulted during the refereeing process.
-
-\subsection{Refereeing Decisions}
-It is expected that accepted papers will be presented in person at SDSS in Pittsburgh 
-from June 4 to 6, 2020.
-Accepted papers' authors will be notified in plenty of time to make registration and travel 
-arrangements and for papers not accepted to submit an e-poster to the 
-conference if they so choose.
+\section{Acknowledgements}
+We would like to thank our sponsors and grant provider NSF, LSAMP Inspire Program and Iowa League of Cities.
+We would like to thank our partners and team members at Iowa State University and University of Nebraska - Lincoln.
+We would like to thank the State of Iowa for providing Data along with our partners in the Rural Iowa communities.
 

### 2021-12-08 SCCAbstract_SDSS2022-knitr.tex
Updating the folder with useful files with name changes

--- 2021-12-08 SCCAbstract_SDSS2022-knitr.Rnw
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{SCCAbstract_SDSS2022-knitr-concordance}
 
-<<setup, include=FALSE, cache = FALSE>>=
-library(knitr)
-opts_chunk$set(
-concordance=TRUE
-)
-@
+\begin{Schunk}
+\begin{Sinput}
+> library(knitr)
+> opts_chunk$set(
++ concordance=TRUE
++ )
+\end{Sinput}
+\end{Schunk}
 
@@
 \section{Data Collection}
-Iowa Data were used to create the SCC dashboard.
-Iowa Data is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
-Iowa Data website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
+Iowa Data Gov data were used to create the SCC dashboard.
+Iowa Data Gov is a public data website that collects and updates data on the state of Iowa, data are mostly on a town/city level that is not collected in National Data Collection Systems.
+Iowa Data Gov website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
 For the purpose of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details.
@@
 \section{Data Challenges}
-Iowa Data website has an exhaustive and detailed data, but that data collected has missing information.
+Iowa Data Gov website has an exhaustive and detailed data, but that data collected has missing information.
 This information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
@@
 
-\section{Current Progress and Future Work}
+\section{Current Work/ Future Work}
 
-\subsection{Current Progress}
-Our research as stated above, we would like to produce and develop a guided discovery dashboard that will allow rural analysts the ability to access information that will help with the Quality of Life in their town.
-This dashboard will be used to help the analyst make data-driven decisions about making effective changes in their communities.
-Our dashboard will be used to
-\subsection{Future Work}
+We encourage the selective use of tables and/or figures as appropriate.
 
-\section{Acknowledgements}
-We would like to thank our sponsors and grant provider NSF, LSAMP Inspire Program and Iowa League of Cities.
-We would like to thank our partners and team members at Iowa State University and University of Nebraska - Lincoln.
-We would like to thank the State of Iowa for providing Data along with our partners in the Rural Iowa communities.
+\section{Discussion/Conclusions}
+
+\subsection{Appendices and acknowledgements}
+
+Please avoid the use of appendices and acknowledgements 
+as separate sections.  Any pertinent material should be included in the
+body of the extended abstract rather than in an appendix.  If 
+acknowledgements are necessary then they should be made using footnotes.
+
+If you wish to provide links to accompanying code, this is allowable because
+SDSS encourages best practices for reproducible research.  However, all
+conclusions and exposition relevant to the refereeing process should appear in
+the body of the extended abstract; furthermore, we cannot guarantee that 
+code will be checked or even consulted during the refereeing process.
+
+\subsection{Refereeing Decisions}
+It is expected that accepted papers will be presented in person at SDSS in Pittsburgh 
+from June 4 to 6, 2020.
+Accepted papers' authors will be notified in plenty of time to make registration and travel 
+arrangements and for papers not accepted to submit an e-poster to the 
+conference if they so choose.
 

### 2021-12-08 SCCAbstract_SDSS2022-knitr.Rnw
Updates 12.08.2021

@@
 \begin{abstract}
+Many small and rural places are shrinking.
 Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
-In our paper, we will explore the specific scope of how dashboards are used in settings when the analysis is a novice to data.
-We will demonstrate the surrounding.
-Our framework will suggest a number of research directions to better support dashboard design, implementation and use for the every day analyst.
+In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
+Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
 \end{abstract}
 
-{\bf Keywords:} Interactive Dashboards, Exploratory Data Analysis (EDA), Design Space, Guided Discovery
+{\bf Keywords:} Interactive Dashboards, Exploratory Data Analysis (EDA), Guided Discovery
 
 \section{Research Problem}
-With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{Fisher}.
+With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{fisher}.
 A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.
 Stephen Few defines a dashboard as:
+
 \begin{quotation}
 \small A visual display of the most important information needed to achieve one or more objectives, consolidated and arranged on a single screen so the information can be monitored at a glance.
-\cite{Few}
+\cite{few}
 \end{quotation}
-Dashboards are used to improve performance, improve strategic decision making, track processes effectiveness, minimize data complexity and communicate organizations' data-driven decision making \cite{...}.
+
 Some communities continue to thrive as they lose population because they adapt and stay focused on quality of life, community services, and investing in the future.
 This is what we call rural smart shrinkage.
-Our research team is developing strategies and sharing examples of successful shrink-smart towns to help similar communities to improve quality of life for their residents\cite{SCC}.
+Our research team is developing strategies and sharing examples of successful shrink-smart towns to help similar communities to improve quality of life for their residents \cite{scc}.
 Our work focuses on developing and challenges novice analysts in advanced statistical concepts and data visualizations to help make decisions.
 Specifically, what data can be used to help make decisions, what strategies overcome the challenges of those analysts feel empowered to make data-driven decisions.
-Our ultimate aim is to help inform Iowa Rural Shrink Smart communities by using effective dashboards to help those towns build the Quality of Life (QoL) Measures.
+Our ultimate aim is to help inform Iowa's Rural Shrink Smart communities by using effective dashboards to help those towns build the Quality of Life (QoL) Measures.
 
 \section{Data Collection}
-Iowa Data were used to create the SCC dashboard.
-Iowa Data is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
-Iowa Data website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
+{\it Iowa Data} were used to create the SCC dashboard.
+{\it Iowa Data}is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
+{\it Iowa Data} website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
 For the purpose of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details.
@@
 \section{Data Challenges}
-Iowa Data website has an exhaustive and detailed data, but that data collected has missing information.
+{\it Iowa Data} website has an exhaustive and detailed data, but that data collected has missing information.
 This information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
@@
 \bibliographystyle{sdss2020} % Please do not change the bibliography style
-\bibliography{SampleReferencesForExtendedAbstract}
+\bibliography{SCCAbstract_SDSS2022}
 

### 2021-12-08 SCCAbstract_SDSS2022-knitr.Rnw
Updates - Denise 12.08.2021

@@
 \usepackage{latexsym}
-\usepackage{amsmath, amsthm, amsfonts}
+\usepackage{amsmath, amsfonts}
 \usepackage{algorithm, algorithmic}  
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
 
-\section{Data Collection}
+\section{Data Description}
 {\it Iowa Data} were used to create the SCC dashboard.
-{\it Iowa Data}is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
+{\it Iowa Data} is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
 {\it Iowa Data} website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
@@
 \section{Current Progress and Future Work}
+Our current design work will be broken down into four categories that have been identified in the following:
+\begin{itemize}
+\item User Flexibility 
+\item Visual, Analytic, and Data Literacy
+\item Data Design
+\item Social Impact
+\end{itemize}
+
+In Figure 1, our dashboard will allow the user to select a town name, which will populate the  information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
+A table will be populated that will include information related to five other towns that have similar attributes, but the aspect of QoL will be included in the differences that help the town make a decision for the town.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=75mm]{SCC_Dashboard.png}
+\caption{Rural Smart EDA Dashboard Design}
+\end{figure}
 
 \subsection{Current Progress}
-Our research as stated above, we would like to produce and develop a guided discovery dashboard that will allow rural analysts the ability to access information that will help with the Quality of Life in their town.
-This dashboard will be used to help the analyst make data-driven decisions about making effective changes in their communities.
-Our dashboard will be used to
+{\bf Statistical Techniques} Our research has incorporated statistical clustering methods, such as K-means and Hierarchical Clustering, to determine the similarities in the towns based on distances and services available.
+As a result, of our research partners we were actually able to use clustering methods to determine and identify missing data.
+We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+
+{\bf Data Science Techniques} Our team has been able to utilize direct API calls and web scraping tools to collect data necessary.
+
+{\bf Dashboard Design Techniques} Our dashboard incorporates the following important dashboard techniques:
+{\it User Flexibility} The aspects of user flexibility are directly based on the detail data adjustments; Adaptability: User and Display; Comparison support and Storytelling.
+For example, we want our users to use the data collected on a mass scale that have and should be summarized in detail but without losing anything.
+
+{\it Visual, Analytic, and Data Literacy} The aspects in our dashboard development primary presents in our ability to ease of usage to the user with visual, data  and analytic literacy.
+We have little information on the users' adaptability to data-driven decisions.
+
+{\it Data Design}.
+In our case, the challenges that arise in our dashboard will include data quality, data grouping, too much data, data sources, inaccurate data representation, and key performance indicators and metrics.
+For example, we have information from {\it Iowa Data} will have missing data can be pointed out by our users' that has not been captured by our data sources, this can be a recent closing of a school due to the pandemic.
+
+{\it Social Impact} This aspect will be about making sure that the users are flexible to the following: data-driven thinking, technology resistance, dashboard aversion, and data sharing, security, and privacy.
+For example, we will encounter town users that will have the urge to compare their town to all types of towns that aren't similar.
+In our research, we will use statistical clustering methods and cater dashboard to show only towns that have similar components.
+
 \subsection{Future Work}
+Our dashboard design has created challenges that promote
 
+We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.
+As a result to our dashboard development, we will develop a feedback loop from the analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
+As a contrast, we will researchers with an extensive understanding of the small towns to be used as a counter example of dashboard adaptability.
+Our research, should be a
+
+So dashboard challenges,
+adaptation techniques, and user strategies are interconnected as
+Figure 1 shows. The rationale is that if we are able to establish
+a relationship between strategies to problems and we can detect
+the strategies in real-time, we can adapt the user interface in realtime or in the next iteration. Then we can say that the adaption
+techniques have a positive effect on the challenges.
 \section{Acknowledgements}

### 2021-12-08 SCCAbstract_SDSS2022-knitr.tex
Updates - Denise 12.08.2021

@@
 % File SDSS2020_SampleExtendedAbstract.tex
-\documentclass[10pt]{article}
+\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{newtxtext, newtxmath, times}
@@
 \date{}
-
-\usepackage{Sweave}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-\input{SCCAbstract_SDSS2022-knitr-concordance}
-
-\begin{Schunk}
-\begin{Sinput}
-> library(knitr)
-> opts_chunk$set(
-+ concordance=TRUE
-+ )
-\end{Sinput}
-\end{Schunk}
 

### 2021-12-08 Old files/SampleExtendedAbstract_SDSS2020.Rnw
Moved old files to folder

--- 2021-12-07 SampleExtendedAbstract_SDSS2020.Rnw
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 {\bf Keywords:} Interactive Dashboards, Exploratory Data Analysis (EDA), Design Space(??)
-
-<<setup, include = F>>=
-# This is a code chunk
-@
 
@@
 
-\section{Data Description}
-Using an exploratory methodology, we collected open source data from the Iowa Data Gov website to understand the
+\section{Data Collection}
+Iowa Data Gov data were used to create the SCC dashboard.
+Iowa Data Gov is a public data website that collects and updates data on the state of Iowa, data are mostly on a town/city level that is not collected in National Data Collection Systems.
+Iowa Data Gov website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
+For the purpose of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details.
+These data are collected on all towns that are available.
+We will use statistical clustering methods will be used to determine the town classifications to help improve the rural towns.
 
-we derived a design space consisting of 3-4 visual and functional aspects of dashboards with statistical exploratory data analysis. see Section~\ref{info-on-sections}.
+\section{Data Challenges}
+Iowa Data Gov website has an exhaustive and detailed data, but that data collected has missing information.
+This information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
+For example, many small towns do not have a hospital that will only cater to the community, but they will have access to a hospital in a neighboring town.
+The second type of missing data are when data should be available and are not currently being collected in the data but are available in a different data source.
+For example, our public school data does not actual represent public schools in bigger cities, suggesting that the closest Elementary school is 20 miles away, which we understood that to be untrue.
+This type of missing data is very typical in data collection, but could led our research in a poor direction.
 
-\subsection{Challenges}
-While data
+\section{Guiding Design Principles}
+Dashboards have a variety of challenges, our research will focus on users' challenges with understanding data and statistical relationships of dashboards.
+We investigate user interaction strategies that will not challenge the knowledge of the analyst in a variety of backgrounds.
+With little research on the adaptation of dashboards, we will particularly focus on the techniques that to solve interaction and information presentation challenges.
+In our research, we would like to use a guiding design principles to develop and design the following:
+\begin{itemize}
+\item Showing a town analyst with peers to help with changes that improved the QoL
+\item Dashboard design that will make the town the main focus without distracting the analysts with peer comparisons
+\item Creating a feedback loop that will, assist in the guided discovery from the analysts
+\end{itemize}
 
-The intended use of a dashboard for a novice user drives the choices in the visual and functional design allowances.
-The factors here capture the roles of each dashboard in the process of analysis and communication.
-We find that the purpose of a dashboard has been substantially expanded from the operational dashboard to capture decision-making at higher levels, and may not be supportive in decision-making.
-
-{\bf Decision Support} (Strategic, Tactical, Operational): The decision support dimension reflects on the sorts of actionable decisions that dashboards are designed to support.
-
-{\bf Communication and Learning:} We encountered several examples of dashboards that did not solicit decision-making on any temporal scale.
-
-\subsection{Audience}
-The visual and functional aspects of our dashboards typically reflect the intended audience, which has a wide-range of characteristics in their domain and visualization aspects of a dashboard and their experience with data.
-
-\section{Related Work}
-From previous research has explained how visual analytics addresses the issue of information overload and enables analysts to transform raw data into salient information and knowledge
-
-\section{Data/Results}
+\section{Current Work/ Future Work}
 
 We encourage the selective use of tables and/or figures as appropriate.
-
-\subsection{Graphics and Tables}
-
-Place figures and tables in the
-paper near where they are first discussed, rather than at the end, if
-possible.  It is acceptable that wide illustrations
-may run across both columns.  The use of color
-is allowed in extended abstract graphics.
 

### 2021-12-16 SCCAbstract_SDSS2022-knitr.Rnw
A few comments

@@
 \usepackage{graphicx}
+% Commands for editing
+\newcommand{\db}[1]{{\textcolor{orange}{#1}}}
+\newcommand{\svp}[1]{{\textcolor{blue}{#1}}}
 
@@
 
-\begin{quotation}
+\begin{quotation}% I don't know that we need this... dashboards are very common, so let's focus on the problem and the dashboard as an obvious solution.
 \small A visual display of the most important information needed to achieve one or more objectives, consolidated and arranged on a single screen so the information can be monitored at a glance.
@@
 
-Some communities continue to thrive as they lose population because they adapt and stay focused on quality of life, community services, and investing in the future.
-This is what we call rural smart shrinkage.
-Our research team is developing strategies and sharing examples of successful shrink-smart towns to help similar communities to improve quality of life for their residents \cite{scc}.
-Our work focuses on developing and challenges novice analysts in advanced statistical concepts and data visualizations to help make decisions.
+Some communities continue to thrive as they lose population because they adapt\svp{, maintaining quality of life and community services for residents while} investing in the future.
+% putting sentences on different lines helps with git merges and also makes it easier to figure out where {} match up :)
+\svp{This process,} \emph{smart shrinkage}\svp{, is important for rural areas who have experienced shrinking populations for decades.} 
+\svp{As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our} research team \svp{will provide communities with data about services in small town Iowa in order to assist with} developing strategies to improve quality of life for their residents amid shrinking populations \cite{scc}.
+\svp{We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
+In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.}
+% Our work focuses on developing and challenges novice analysts in advanced statistical concepts and data visualizations to help make decisions.
 Specifically, what data can be used to help make decisions, what strategies overcome the challenges of those analysts feel empowered to make data-driven decisions.

### 2021-12-22 SCCAbstract_SDSS2022-knitr.Rnw
Update SCCAbstract_SDSS2022-knitr.Rnw

@@
 {\it Iowa Data} is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
-{\it Iowa Data} website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
-For the purpose of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details.
-These data are collected on all towns that are available.
-We will use statistical clustering methods will be used to determine the town classifications to help improve the rural towns.
+\db{{\it Iowa Data} }website is a unique and detailed database of information about residents, such as liquor sales and school building locations. 
+\db{For the purpose} of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details. 
+\db{These data} are collected on all towns that are available. We will use statistical clustering methods will be used to determine the town classifications to help improve the rural towns.
 
@@
 {\it Iowa Data} website has an exhaustive and detailed data, but that data collected has missing information. 
-This information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
-For example, many small towns do not have a hospital that will only cater to the community, but they will have access to a hospital in a neighboring town.
+\db{This} information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
+\db{For example}, many small towns do not have a hospital that will only cater to the community, but they will have access to a hospital in a neighboring town. 
 The second type of missing data are when data should be available and are not currently being collected in the data but are available in a different data source.
-For example, our public school data does not actual represent public schools in bigger cities, suggesting that the closest Elementary school is 20 miles away, which we understood that to be untrue.
-This type of missing data is very typical in data collection, but could led our research in a poor direction.
+\db{For example}, our public school data does not actual represent public schools in bigger cities, suggesting that the closest Elementary school is 20 miles away, which we understood that to be untrue.
+\db{This} type of missing data is very typical in data collection, but could led our research in a poor direction.
 

### 2021-12-23 SCCAbstract_SDSS2022-knitr.Rnw
A few more changes

--- 2021-12-16 SCCAbstract_SDSS2022-knitr.Rnw
@@
 \usepackage{graphicx}
+\usepackage[dvipsnames]{xcolor} % colors
+
 % Commands for editing
@@
 \section{Data Description}
-{\it Iowa Data} were used to create the SCC dashboard.
-{\it Iowa Data} is a public data website that collects and updates data on the State of Iowa Government, data are mostly on a town/city level that is not collected in National Data Collection Systems.
-{\it Iowa Data} website is a unique and detailed database of information about residents, such as liquor sales and school building locations.
-For the purpose of the Rural Shrink Smart Dashboard, we will collect as much data as possible, such as obtaining fire department and hospital details.
-These data are collected on all towns that are available.
-We will use statistical clustering methods will be used to determine the town classifications to help improve the rural towns.
+\svp{Data collected from \url{data.iowa.gov}} were used to create the SCC dashboard.
+\svp{Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.} \svp{\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.}
+\svp{Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
+In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.}
 
-\section{Data Challenges}
-{\it Iowa Data} website has an exhaustive and detailed data, but that data collected has missing information.
-This information can be broken down into two separate types, in which we will call a true missing, a particular service that is missing because the town does not have or offer the service to the community.
-For example, many small towns do not have a hospital that will only cater to the community, but they will have access to a hospital in a neighboring town.
-The second type of missing data are when data should be available and are not currently being collected in the data but are available in a different data source.
-For example, our public school data does not actual represent public schools in bigger cities, suggesting that the closest Elementary school is 20 miles away, which we understood that to be untrue.
-This type of missing data is very typical in data collection, but could led our research in a poor direction.
+\svp{One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.}
+\svp{Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
+This makes both visualization and statistical analysis more complicated.}
 
 \section{Guiding Design Principles}
-Dashboards have a variety of challenges, our research will focus on users' challenges with understanding data and statistical relationships of dashboards.
-We investigate user interaction strategies that will not challenge the knowledge of the analyst in a variety of backgrounds.
-With little research on the adaptation of dashboards, we will particularly focus on the techniques that to solve interaction and information presentation challenges.
-In our research, we would like to use a guiding design principles to develop and design the following:
+\svp{An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
+That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
+With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.}
+\svp{Thus, our initial set of dashboard design principles is as follows: }
 \begin{itemize}
-\item Showing a town analyst with peers to help with changes that improved the QoL
-\item Dashboard design that will make the town the main focus without distracting the analysts with peer comparisons
-\item Creating a feedback loop that will, assist in the guided discovery from the analysts
+\item \svp{The town leaders are the focus audience; thus, the town itself should be the central focus of the app.}
+\item \svp{Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.}
+\item \svp{Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.}
+\item \svp{Allow for more detailed data and feature requests to improve the dashboard design over time.}
 \end{itemize}

### 2022-01-05 SCCAbstract_SDSS2022-knitr.tex
Changes 1.5.2022

@@
 \usepackage{graphicx}
+\usepackage[dvipsnames]{xcolor} % colors
+
+% Commands for editing
+\newcommand{\db}[1]{{\textcolor{orange}{#1}}}
+\newcommand{\svp}[1]{{\textcolor{blue}{#1}}}
 

### 2022-01-10 SCCAbstract_SDSS2022-knitr.tex
Abstract Updates 1.10.2022

@@
 A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.
-Stephen Few defines a dashboard as:
 
-\begin{quotation}% I don't know that we need this... dashboards are very common, so let's focus on the problem and the dashboard as an obvious solution.
-\small A visual display of the most important information needed to achieve one or more objectives, consolidated and arranged on a single screen so the information can be monitored at a glance.
+% Stephen Few defines a dashboard as:
+% 
+% \begin{quotation}% I don't know that we need this... dashboards are very common, so let's focus on the problem and the dashboard as an obvious solution.
+% \small A visual display of the most important information needed to achieve one or more objectives, consolidated and arranged on a single screen so the information can be monitored at a glance.
 \cite{few}
-\end{quotation}
+% \end{quotation}
 
@@
 \section{Current Progress and Future Work}
-Our current design work will be broken down into four categories that have been identified in the following:
+In our research, it is important to use visual orientation of dashboards along with the framework of Guided Discovery Learning (GDL), to make sure that the town analyst has the ability to understand the visual orientation easily as well as guided them to new discovery in the dashboard.
+Stephen Few noted in \italize{Information Dashboard Design: The effective visual communication of data} the following:
+\begin{quotation}
+\smaller{The visual orientation of dashboards is important due to the speed of perception that’s usually required to monitor information.
+The faster you must assess what’s going on, the more you should rely on graphical means to display the information.
+\cite{few} }
+\end{quotation}
+
+Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we will focus on making sure that the following three common mistakes are avoided to make sure that the town analyst have the best path to success:
 \begin{itemize}
-\item User Flexibility 
-\item Visual, Analytic, and Data Literacy
-\item Data Design
-\item Social Impact
+\item{Choosing inappropriate display media}
+\item{Ineffectively highlighting what’s important}
+\item{Misusing or overusing color}
 \end{itemize}
+
+This list will be the guiding list that can push our work in the direction that is far away from the work done in the workplace in the past.
+With a great deal of data, many organizations fall into the trap of over ineffectively highlighting what's important along with choosing inappropriate display media and misusing color.
+This is mostly done, in order to make sure that all of the information is given to the user.
+This could be a major hurdle when we are collecting a great deal of open-source data for the small towns of Iowa.
+
+A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge. DeDonno stated:
+ \begin{quotation}
+\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
+ \end{quotation}
+
+Our current work will incorporate the ideas of GDL, in which dashboard design will be used to display a panel with vital statistics making sure that the town's analyst small town is the center of focus.
+The vital statistics will show and highlight the town's financial health, shortest distances from city center to critical resources, such as the fire department, hospital, post offices and public schools.
+The dashboard will include Quality of Life (QoL) Metrics and the financial health of the small town.
+Based on the number of variables that are used in making sure that the dashboard is extensive and useful, we will use parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+
+In the work of the data visualization portion of our research work, a parallel (coordinates) plot allows to compare the feature of several individual observations (series) on a set of numeric variables.
+Each vertical bar represents a variable and often has its own scale.
+(The units can even be different).
+Values are then plotted as series of lines connected across each axis.
+A parallel coordinate plot is used traditionally in air traffic control to help with collision avoidance algorithms, but has been adapted in the statistical data visualization space with an emphasis on three important considerations: the order, the rotation, and the scaling of the axes.
+
+Using parallel (coordinates) plot will allow
+
+To determine the five similar towns, we will implement unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, based on distances and services available.
+We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+Using GDL,  our dashboard will make sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
+Our research partners found out that the small towns do not want to be told what to do but would like to make decisions that are of those who have been apart of the community for generations.
+Our dashboard will hopefully make sure that the town's analyst are guided in the direction that will help the town's population shrink smart.
 
@@
 
-\subsection{Current Progress}
-{\bf Statistical Techniques} Our research has incorporated statistical clustering methods, such as K-means and Hierarchical Clustering, to determine the similarities in the towns based on distances and services available.
-As a result, of our research partners we were actually able to use clustering methods to determine and identify missing data.
-We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+\subsection{Future Work}
+We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.
+As a result to our dashboard development, we will develop a continual feedback loop from the town analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
+Our team plans to capture useful metrics on the use of the product allowing the quantitative and qualitative data collection to allow for a better user experience.
+This can promote challenges in the event of towns feeling "studied" by people in cities outside of their own.
 
-{\bf Data Science Techniques} Our team has been able to utilize direct API calls and web scraping tools to collect data necessary.
-
-{\bf Dashboard Design Techniques} Our dashboard incorporates the following important dashboard techniques:
-{\it User Flexibility} The aspects of user flexibility are directly based on the detail data adjustments; Adaptability: User and Display; Comparison support and Storytelling.
-For example, we want our users to use the data collected on a mass scale that have and should be summarized in detail but without losing anything.
-
-{\it Visual, Analytic, and Data Literacy} The aspects in our dashboard development primary presents in our ability to ease of usage to the user with visual, data  and analytic literacy.
-We have little information on the users' adaptability to data-driven decisions.
-
-{\it Data Design}.
-In our case, the challenges that arise in our dashboard will include data quality, data grouping, too much data, data sources, inaccurate data representation, and key performance indicators and metrics.
-For example, we have information from {\it Iowa Data} will have missing data can be pointed out by our users' that has not been captured by our data sources, this can be a recent closing of a school due to the pandemic.
-
-{\it Social Impact} This aspect will be about making sure that the users are flexible to the following: data-driven thinking, technology resistance, dashboard aversion, and data sharing, security, and privacy.
-For example, we will encounter town users that will have the urge to compare their town to all types of towns that aren't similar.
-In our research, we will use statistical clustering methods and cater dashboard to show only towns that have similar components.
-
-\subsection{Future Work}
-Our dashboard design has created challenges that promote
-
-We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.
-As a result to our dashboard development, we will develop a feedback loop from the analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
-As a contrast, we will researchers with an extensive understanding of the small towns to be used as a counter example of dashboard adaptability.
-Our research, should be a
-
-So dashboard challenges,
-adaptation techniques, and user strategies are interconnected as
-Figure 1 shows. The rationale is that if we are able to establish
-a relationship between strategies to problems and we can detect
-the strategies in real-time, we can adapt the user interface in realtime or in the next iteration. Then we can say that the adaption
-techniques have a positive effect on the challenges.
 \section{Acknowledgements}

### 2022-01-13 SCCAbstract_SDSS2022-knitr.tex
Removed Quotes and added language on work done

@@
 
-\section{Current Progress and Future Work}
+\section{Current Progress}
 In our research, it is important to use visual orientation of dashboards along with the framework of Guided Discovery Learning (GDL), to make sure that the town analyst has the ability to understand the visual orientation easily as well as guided them to new discovery in the dashboard.
-Stephen Few noted in \italize{Information Dashboard Design: The effective visual communication of data} the following:
-\begin{quotation}
-\smaller{The visual orientation of dashboards is important due to the speed of perception that’s usually required to monitor information.
+Stephen Few noted in \italize{Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is important due to the speed at which perception \cite{few}.
+
+%\begin{quotation}
+%\smaller{The visual orientation of dashboards is important due to the speed of perception that’s usually required to monitor information.
 The faster you must assess what’s going on, the more you should rely on graphical means to display the information.
 \cite{few} }
-\end{quotation}
+%\end{quotation}
 
-Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we will focus on making sure that the following three common mistakes are avoided to make sure that the town analyst have the best path to success:
+Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we will focus on making sure that the following three common mistakes on workplace dashboard development are avoided to make sure that the town analyst have the best path to success:
 \begin{itemize}
@@
 
-This list will be the guiding list that can push our work in the direction that is far away from the work done in the workplace in the past.
-With a great deal of data, many organizations fall into the trap of over ineffectively highlighting what's important along with choosing inappropriate display media and misusing color.
+The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace.
+With the amount of data, our analyst fall into the trap of over ineffectively highlighting what's important along with choosing inappropriate display media and misusing color.
 This is mostly done, in order to make sure that all of the information is given to the user.
-This could be a major hurdle when we are collecting a great deal of open-source data for the small towns of Iowa.
+This has been a major hurdle when displaying the amount data collected from open-source data for the small towns of Iowa.
 
-A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge. DeDonno stated:
- \begin{quotation}
-\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
- \end{quotation}
+%Continue to cite this in the documentation but remove the quotation.
 
-Our current work will incorporate the ideas of GDL, in which dashboard design will be used to display a panel with vital statistics making sure that the town's analyst small town is the center of focus.
-The vital statistics will show and highlight the town's financial health, shortest distances from city center to critical resources, such as the fire department, hospital, post offices and public schools.
-The dashboard will include Quality of Life (QoL) Metrics and the financial health of the small town.
-Based on the number of variables that are used in making sure that the dashboard is extensive and useful, we will use parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
+%\begin{quotation}
+%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
+%\end{quotation}
 
-In the work of the data visualization portion of our research work, a parallel (coordinates) plot allows to compare the feature of several individual observations (series) on a set of numeric variables.
-Each vertical bar represents a variable and often has its own scale.
-(The units can even be different).
-Values are then plotted as series of lines connected across each axis.
-A parallel coordinate plot is used traditionally in air traffic control to help with collision avoidance algorithms, but has been adapted in the statistical data visualization space with an emphasis on three important considerations: the order, the rotation, and the scaling of the axes.
+Our current work incorporates the ideas of GDL, in which dashboard design have been used to display a panel with vital statistics making sure that the town's analyst small town is the center of focus.
+The vital statistics show and highlight the town's financial health, shortest distances from city center to critical resources, such as the fire department, hospital, post offices and public schools.
+The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
+Based on the number of variables that are used in making sure that the dashboard is extensive and useful, we implemented the use of parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+A parallel (coordinates) plot allowed our town analyst to compare the important features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms, but has been adapted in the statistical data visualization space with an emphasis on three important considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots, will display variables in categorically and numerically by scaling the data.
 
-Using parallel (coordinates) plot will allow
-
-To determine the five similar towns, we will implement unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, based on distances and services available.
-We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
-Using GDL,  our dashboard will make sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
-Our research partners found out that the small towns do not want to be told what to do but would like to make decisions that are of those who have been apart of the community for generations.
+To determine the five similar towns, we have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, based on distances and services available.
+We continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+Using GDL, our dashboard make sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
+Our research partners have worked with five towns directly, who have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.As a a result, we found out that the small towns do not want to be told what to do but would like to make decisions that are of those who have been apart of the community for generations.
 Our dashboard will hopefully make sure that the town's analyst are guided in the direction that will help the town's population shrink smart.
 
-In Figure 1, our dashboard will allow the user to select a town name, which will populate the information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
-A table will be populated that will include information related to five other towns that have similar attributes, but the aspect of QoL will be included in the differences that help the town make a decision for the town.
+Our dashboard development has become more focus on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
+Our team has captured useful metrics on the use of the product allowing the quantitative and qualitative data collection to better user experience.
+This can promote challenges in the event of towns feeling "studied" by people in cities outside of their own.
+
+In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
+A table populates in the vital statistics sections that has information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similar based in the most common variables in the towns, such as the distance of fire department.
 
@@
 
-\subsection{Future Work}
-We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.
-As a result to our dashboard development, we will develop a continual feedback loop from the town analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
-Our team plans to capture useful metrics on the use of the product allowing the quantitative and qualitative data collection to allow for a better user experience.
-This can promote challenges in the event of towns feeling "studied" by people in cities outside of their own.
+%To-Do: Write this in the way that we have done this because by the time that we present this will be done. 
+%\subsection{Future Work}
 

### 2022-01-13 SCCAbstract_SDSS2022-knitr.Rnw
More updates

@@
 
-\section{Current Progress and Future Work}
-Our current design work will be broken down into four categories that have been identified in the following:
+\section{Current Progress}
+In our research, it is important to use visual orientation of dashboards along with the framework of Guided Discovery Learning (GDL), to make sure that the town analyst has the ability to understand the visual orientation easily as well as guided them to new discovery in the dashboard.
+Stephen Few noted in \italize{Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is important due to the speed at which perception \cite{few}.
+
+%\begin{quotation}
+%\smaller{The visual orientation of dashboards is important due to the speed of perception that’s usually required to monitor information.
+The faster you must assess what’s going on, the more you should rely on graphical means to display the information.
+\cite{few} }
+%\end{quotation}
+
+%Continue to cite this in the documentation but remove the quotation.
+
+A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
+%\begin{quotation}
+%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
+%\end{quotation}
+Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
 \begin{itemize}
-\item User Flexibility 
-\item Visual, Analytic, and Data Literacy
-\item Data Design
-\item Social Impact
+\item{\br{Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualizing data of five towns in parallel coordinate plots.
+}
+\item{\br{Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user, as such we added their information in the focal point in their center of the page.
+The comparison towns are only in the parallel coordinate plots, to remind the user that their town is most important items in the dashboard.}
+\item{\br{Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
 \end{itemize}
 
-In Figure 1, our dashboard will allow the user to select a town name, which will populate the  information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
-A table will be populated that will include information related to five other towns that have similar attributes, but the aspect of QoL will be included in the differences that help the town make a decision for the town.
+The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace.
+With the amount of data, our analyst fall into the trap of over ineffectively highlighting what's important along with choosing inappropriate display media and misusing color.
+This is mostly done, in order to make sure that all of the information is given to the user.
+This has been a major hurdle when displaying the amount data collected from open-source data for the small towns of Iowa.
+
+Our current work incorporates the ideas of GDL, in which dashboard design have been used to display a panel with vital statistics making sure that the town's analyst small town is the center of focus.
+The vital statistics show and highlight the town's financial health, shortest distances from city center to critical resources, such as the fire department, hospital, post offices and public schools.
+The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
+Based on the number of variables that are used in making sure that the dashboard is extensive and useful, we implemented the use of parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+A parallel (coordinates) plot allowed our town analyst to compare the important features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms, but has been adapted in the statistical data visualization space with an emphasis on three important considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots, will display variables in categorically and numerically by scaling the data.
+
+To determine the five similar towns, we have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, based on distances and services available.
+We continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+Using GDL, our dashboard make sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
+Our research partners have worked with five towns directly, who have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.As a a result, we found out that the small towns do not want to be told what to do but would like to make decisions that are of those who have been apart of the community for generations.
+Our dashboard will hopefully make sure that the town's analyst are guided in the direction that will help the town's population shrink smart.
+
+Our dashboard development has become more focus on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
+Our team has captured useful metrics on the use of the product allowing the quantitative and qualitative data collection to better user experience.
+This can promote challenges in the event of towns feeling "studied" by people in cities outside of their own.
+
+In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
+A table populates in the vital statistics sections that has information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similar based in the most common variables in the towns, such as the distance of fire department.
 
@@
 
-\subsection{Current Progress}
-{\bf Statistical Techniques} Our research has incorporated statistical clustering methods, such as K-means and Hierarchical Clustering, to determine the similarities in the towns based on distances and services available.
-As a result, of our research partners we were actually able to use clustering methods to determine and identify missing data.
-We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+In conclusion,
 
-{\bf Data Science Techniques} Our team has been able to utilize direct API calls and web scraping tools to collect data necessary.
+%To-Do: Write this in the way that we have done this because by the time that we present this will be done. 
+%\subsection{Future Work}
 
-{\bf Dashboard Design Techniques} Our dashboard incorporates the following important dashboard techniques:
-{\it User Flexibility} The aspects of user flexibility are directly based on the detail data adjustments; Adaptability: User and Display; Comparison support and Storytelling.
-For example, we want our users to use the data collected on a mass scale that have and should be summarized in detail but without losing anything.
-
-{\it Visual, Analytic, and Data Literacy} The aspects in our dashboard development primary presents in our ability to ease of usage to the user with visual, data  and analytic literacy.
-We have little information on the users' adaptability to data-driven decisions.
-
-{\it Data Design}.
-In our case, the challenges that arise in our dashboard will include data quality, data grouping, too much data, data sources, inaccurate data representation, and key performance indicators and metrics.
-For example, we have information from {\it Iowa Data} will have missing data can be pointed out by our users' that has not been captured by our data sources, this can be a recent closing of a school due to the pandemic.
-
-{\it Social Impact} This aspect will be about making sure that the users are flexible to the following: data-driven thinking, technology resistance, dashboard aversion, and data sharing, security, and privacy.
-For example, we will encounter town users that will have the urge to compare their town to all types of towns that aren't similar.
-In our research, we will use statistical clustering methods and cater dashboard to show only towns that have similar components.
-
-\subsection{Future Work}
-Our dashboard design has created challenges that promote
-
-We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.
-As a result to our dashboard development, we will develop a feedback loop from the analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
-As a contrast, we will researchers with an extensive understanding of the small towns to be used as a counter example of dashboard adaptability.
-Our research, should be a
-
-So dashboard challenges,
-adaptation techniques, and user strategies are interconnected as
-Figure 1 shows. The rationale is that if we are able to establish
-a relationship between strategies to problems and we can detect
-the strategies in real-time, we can adapt the user interface in realtime or in the next iteration. Then we can say that the adaption
-techniques have a positive effect on the challenges.
 \section{Acknowledgements}

### 2022-01-13 SCCAbstract_SDSS2022-knitr.tex
More updates

@@
 
-Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we will focus on making sure that the following three common mistakes on workplace dashboard development are avoided to make sure that the town analyst have the best path to success:
+%Continue to cite this in the documentation but remove the quotation.
+
+A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
+%\begin{quotation}
+%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
+%\end{quotation}
+Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
 \begin{itemize}
-\item{Choosing inappropriate display media}
-\item{Ineffectively highlighting what’s important}
-\item{Misusing or overusing color}
+\item{\br{Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualizing data of five towns in parallel coordinate plots.
+}
+\item{\br{Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user, as such we added their information in the focal point in their center of the page.
+The comparison towns are only in the parallel coordinate plots, to remind the user that their town is most important items in the dashboard.}
+\item{\br{Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
 \end{itemize}
@@
 This has been a major hurdle when displaying the amount data collected from open-source data for the small towns of Iowa.
-
-%Continue to cite this in the documentation but remove the quotation.
-
-A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
-%\begin{quotation}
-%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
-%\end{quotation}
 
@@
 
+In conclusion,
+
 %To-Do: Write this in the way that we have done this because by the time that we present this will be done. 

### 2022-01-17 SCCAbstract_SDSS2022-knitr.tex
Updates 1.17.2022

--- 2022-01-13 SCCAbstract_SDSS2022-knitr.tex
@@
 \definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\makeatletter
+\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}}}
+\makeatother
 \newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
@@
 \definecolor{errorcolor}{rgb}{1, 0, 0}
+\makeatletter
+\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+}}
+\makeatother
 \newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
@@
 \section{Current Progress}
-In our research, it is important to use visual orientation of dashboards along with the framework of Guided Discovery Learning (GDL), to make sure that the town analyst has the ability to understand the visual orientation easily as well as guided them to new discovery in the dashboard.
-Stephen Few noted in \italize{Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is important due to the speed at which perception \cite{few}.
+In our research, it was necessary to use the visual orientation of dashboards and the framework of Guided Discovery Learning (GDL) to make sure that the town analyst can understand the visual orientation easily and guide them to a discovery about possible changes in the dashboard.
+Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.
 
@@
 
-Few outlined common mistakes that are made in \italize{Information Dashboard Design}, we will focus on making sure that the following three common mistakes on workplace dashboard development are avoided to make sure that the town analyst have the best path to success:
+%Continue to cite this in the documentation but remove the quotation.
+
+A bit of background on Guided Discovery, which is defined as an environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
+%\begin{quotation}
+%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
+%\end{quotation}
+Few outlined common mistakes that are made in {\it Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
 \begin{itemize}
-\item{Choosing inappropriate display media}
-\item{Ineffectively highlighting what’s important}
-\item{Misusing or overusing color}
+\item{ {\bf Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualize data of five towns in parallel coordinate plots.}
+\item{ {\bf Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user; as such, we added their information in the focal point in the center of the page.
+The comparison towns are only in the parallel coordinate plots to remind users that their town is the most important item in the dashboard.}
+\item{ {\bf Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
 \end{itemize}
@@
 The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace.
-With the amount of data, our analyst fall into the trap of over ineffectively highlighting what's important along with choosing inappropriate display media and misusing color.
-This is mostly done, in order to make sure that all of the information is given to the user.
-This has been a major hurdle when displaying the amount data collected from open-source data for the small towns of Iowa.
+With the amount of data, our dashboard could fall into the trap of over ineffectively highlighting what's important and choosing inappropriate display media and misusing color to ensure that the analyst's attention and all of the information is given to the user.
+This has been a significant hurdle when displaying the amount of data collected from open-source data for the small towns of Iowa.
 
-%Continue to cite this in the documentation but remove the quotation.
+Our current work incorporates the ideas of GDL, in which dashboard design has been used to display a panel with vital statistics, ensuring that the town's analyst small town is the center of focus.
+The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
+The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
+Based on the number of variables used to ensure that the dashboard is extensive and valuable, we implemented parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots will display variables categorically and numerically by scaling the data.
 
-A bit of background on Guided Discovery, which is defined as a environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
-%\begin{quotation}
-%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
-%\end{quotation}
+We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
+We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
+Using GDL, our dashboard makes sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
+Our research partners have worked with five towns directly, who have agreed to partner with the project to help identify the best practices that are useful in the small and rural Iowa towns.
+As a result, we found out that the small towns do not want to be told what to do but would like to make decisions for those who have been a part of the community for generations.
+Our dashboard will hopefully make sure that the town's analysts are guided in the direction that will help the town's population shrink smart.
 
-Our current work incorporates the ideas of GDL, in which dashboard design have been used to display a panel with vital statistics making sure that the town's analyst small town is the center of focus.
-The vital statistics show and highlight the town's financial health, shortest distances from city center to critical resources, such as the fire department, hospital, post offices and public schools.
-The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
-Based on the number of variables that are used in making sure that the dashboard is extensive and useful, we implemented the use of parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
-A parallel (coordinates) plot allowed our town analyst to compare the important features of several individual towns on a set of numeric and categorical variables collected.
-Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms, but has been adapted in the statistical data visualization space with an emphasis on three important considerations: the order, the rotation, and the scaling of the axes.
-Our coordinate plots, will display variables in categorically and numerically by scaling the data.
+Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
+Our team has captured valuable metrics on using the product allowing the quantitative and qualitative data collection to better user experience.
+This can promote challenges in towns feeling "studied" by people in cities outside of their own.
 
-To determine the five similar towns, we have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, based on distances and services available.
-We continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
-Using GDL, our dashboard make sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
-Our research partners have worked with five towns directly, who have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.As a a result, we found out that the small towns do not want to be told what to do but would like to make decisions that are of those who have been apart of the community for generations.
-Our dashboard will hopefully make sure that the town's analyst are guided in the direction that will help the town's population shrink smart.
-
-Our dashboard development has become more focus on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
-Our team has captured useful metrics on the use of the product allowing the quantitative and qualitative data collection to better user experience.
-This can promote challenges in the event of towns feeling "studied" by people in cities outside of their own.
-
-In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
-A table populates in the vital statistics sections that has information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similar based in the most common variables in the towns, such as the distance of fire department.
+In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
+A table populates in the vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as the distance of fire department.
 

### 2022-01-17 SCCAbstract_SDSS2022-knitr.Rnw
Fix bibliography and clarify some things.

--- 2021-12-23 SCCAbstract_SDSS2022-knitr.Rnw
@@
 
-\section{Current Progress and Future Work}
-Our current design work will be broken down into four categories that have been identified in the following:
-\begin{itemize}
-\item User Flexibility 
-\item Visual, Analytic, and Data Literacy
-\item Data Design
-\item Social Impact
-\end{itemize}
+\section{Current Progress}
+\svp{We leverage} the framework of Guided Discovery Learning (GDL) to guide \svp{the town leaders} to \svp{make discoveries using our interactive visualization}.
+\svp{This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration}\cite{dedonno}.
+% Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.
 
-In Figure 1, our dashboard will allow the user to select a town name, which will populate the  information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals.
-A table will be populated that will include information related to five other towns that have similar attributes, but the aspect of QoL will be included in the differences that help the town make a decision for the town.
+%\begin{quotation}
+%\smaller{The visual orientation of dashboards is important due to the speed of perception that’s usually required to monitor information.
+The faster you must assess what’s going on, the more you should rely on graphical means to display the information.
+\cite{few} }
+%\end{quotation}
+
+%Continue to cite this in the documentation but remove the quotation.
+
+% A bit of background on Guided Discovery, which is defined as an environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information .
+%\begin{quotation}
+%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
+%\end{quotation}
+% Few outlined common mistakes that are made in {\it Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
+% \begin{itemize}
+% \item{ {\bf Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualize data of five towns in parallel coordinate plots.}
+% \item{ {\bf Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user; as such, we added their information in the focal point in the center of the page.
+The comparison towns are only in the parallel coordinate plots to remind users that their town is the most important item in the dashboard.}
+% \item{ {\bf Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
+% \end{itemize}
+
+\svp{We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.}
+
+\svp{As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \citep{few}.}
+
+% The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace.
+With the amount of data, our dashboard could fall into the trap of over ineffectively highlighting what's important and choosing inappropriate display media and misusing color to ensure that the analyst's attention and all of the information is given to the user.
+This has been a significant hurdle when displaying the amount of data collected from open-source data for the small towns of Iowa.
+
+Our current work incorporates the ideas of GDL, in which dashboard design has been used to display a panel with vital statistics, ensuring that the town's analyst small town is the center of focus.
+The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
+The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
+Based on the number of variables used to ensure that the dashboard is extensive and valuable, we implemented parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots will display variables categorically and numerically by scaling the data.
+
+We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
+We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
+Using GDL, our dashboard makes sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions.
+Our research partners have worked with five towns directly, who have agreed to partner with the project to help identify the best practices that are useful in the small and rural Iowa towns.
+As a result, we found out that the small towns do not want to be told what to do but would like to make decisions for those who have been a part of the community for generations.
+Our dashboard will hopefully make sure that the town's analysts are guided in the direction that will help the town's population shrink smart.
+
+Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
+Our team has captured valuable metrics on using the product allowing the quantitative and qualitative data collection to better user experience.
+This can promote challenges in towns feeling "studied" by people in cities outside of their own.
+
+In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
+A table populates in the vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as the distance of fire department.
 
@@
 
-\subsection{Current Progress}
-{\bf Statistical Techniques} Our research has incorporated statistical clustering methods, such as K-means and Hierarchical Clustering, to determine the similarities in the towns based on distances and services available.
-As a result, of our research partners we were actually able to use clustering methods to determine and identify missing data.
-We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.
+%To-Do: Write this in the way that we have done this because by the time that we present this will be done. 
+%\subsection{Future Work}
 
-{\bf Data Science Techniques} Our team has been able to utilize direct API calls and web scraping tools to collect data necessary.
-
-{\bf Dashboard Design Techniques} Our dashboard incorporates the following important dashboard techniques:
-{\it User Flexibility} The aspects of user flexibility are directly based on the detail data adjustments; Adaptability: User and Display; Comparison support and Storytelling.
-For example, we want our users to use the data collected on a mass scale that have and should be summarized in detail but without losing anything.
-
-{\it Visual, Analytic, and Data Literacy} The aspects in our dashboard development primary presents in our ability to ease of usage to the user with visual, data  and analytic literacy.
-We have little information on the users' adaptability to data-driven decisions.
-
-{\it Data Design}.
-In our case, the challenges that arise in our dashboard will include data quality, data grouping, too much data, data sources, inaccurate data representation, and key performance indicators and metrics.
-For example, we have information from {\it Iowa Data} will have missing data can be pointed out by our users' that has not been captured by our data sources, this can be a recent closing of a school due to the pandemic.
-
-{\it Social Impact} This aspect will be about making sure that the users are flexible to the following: data-driven thinking, technology resistance, dashboard aversion, and data sharing, security, and privacy.
-For example, we will encounter town users that will have the urge to compare their town to all types of towns that aren't similar.
-In our research, we will use statistical clustering methods and cater dashboard to show only towns that have similar components.
-
-\subsection{Future Work}
-Our dashboard design has created challenges that promote
-
-We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns.
-As a result to our dashboard development, we will develop a feedback loop from the analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa.
-As a contrast, we will researchers with an extensive understanding of the small towns to be used as a counter example of dashboard adaptability.
-Our research, should be a
-
-So dashboard challenges,
-adaptation techniques, and user strategies are interconnected as
-Figure 1 shows. The rationale is that if we are able to establish
-a relationship between strategies to problems and we can detect
-the strategies in real-time, we can adapt the user interface in realtime or in the next iteration. Then we can say that the adaption
-techniques have a positive effect on the challenges.
 \section{Acknowledgements}

### 2022-01-17 SCCAbstract_SDSS2022-knitr.tex
App updates

@@
 \section{Current Progress}
-In our research, it was necessary to use the visual orientation of dashboards and the framework of Guided Discovery Learning (GDL) to make sure that the town analyst can understand the visual orientation easily and guide them to a discovery about possible changes in the dashboard.
-Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.
+\svp{We leverage} the framework of Guided Discovery Learning (GDL) to guide \svp{the town leaders} to \svp{make discoveries using our interactive visualization}.
+\svp{This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration}\cite{dedonno}.
+% Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.
 
@@
 
-A bit of background on Guided Discovery, which is defined as an environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information \cite{dedonno}.
+% A bit of background on Guided Discovery, which is defined as an environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information .
 %\begin{quotation}
@@
 %\end{quotation}
-Few outlined common mistakes that are made in {\it Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
-\begin{itemize}
-\item{ {\bf Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualize data of five towns in parallel coordinate plots.}
-\item{ {\bf Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user; as such, we added their information in the focal point in the center of the page.
+% Few outlined common mistakes that are made in {\it Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
+% \begin{itemize}
+% \item{ {\bf Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualize data of five towns in parallel coordinate plots.}
+% \item{ {\bf Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user; as such, we added their information in the focal point in the center of the page.
 The comparison towns are only in the parallel coordinate plots to remind users that their town is the most important item in the dashboard.}
-\item{ {\bf Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
-\end{itemize}
+% \item{ {\bf Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
+% \end{itemize}
 
-The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace.
+\svp{We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.}
+
+\svp{As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \cite{few}.}
+
+% The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace.
 With the amount of data, our dashboard could fall into the trap of over ineffectively highlighting what's important and choosing inappropriate display media and misusing color to ensure that the analyst's attention and all of the information is given to the user.

### 2022-01-17 SCCAbstract_SDSS2022-knitr.Rnw
cite not citep
Updating more about the GDL piece

@@
 We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \citep{few}.}
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \cite{few}.}
 
@@
 
-Our current work incorporates the ideas of GDL, in which dashboard design has been used to display a panel with vital statistics, ensuring that the town's analyst small town is the center of focus.
+Our current work incorporates the ideas of GDL, in which our dashboard design has been used to display a panel with vital statistics, ensuring that the town leaders of a small town is the center of focus on the dashboard.
 The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
-The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
-Based on the number of variables used to ensure that the dashboard is extensive and valuable, we implemented parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
+To complement the Rural Shrink Smart NSF Grant, our dashboard includes Quality of Life (QoL) Metrics, to help identify communities social networks along with financial health data collected to ensure a robust identification of our small towns.
+
+To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
+The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
+As stated by DeDonno \cite{dedonno}, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
+
 A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
@@
 \centering
-\includegraphics[width=75mm]{SCC_Dashboard.png}
+\includegraphics[width=75mm]{SCC_Dashboard_pcp2.png}
 \caption{Rural Smart EDA Dashboard Design}

### 2022-01-18 SCCAbstract_SDSS2022-knitr.tex
More updates with a better focus on the work done in the dashboard

@@
 
-Our current work incorporates the ideas of GDL, in which our dashboard design has been used to display a panel with vital statistics, ensuring that the town leaders of a small town is the center of focus on the dashboard.
-The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
-To complement the Rural Shrink Smart NSF Grant, our dashboard includes Quality of Life (QoL) Metrics, to help identify communities social networks along with financial health data collected to ensure a robust identification of our small towns.
+\db{The Rural Shrink Smart EDA dashboard incorporates two major components: 
+  \begin{enumerate}
+  \item{Data Visualization Interface component with GDL principles.} 
+  \item{Exploratory Data Analysis component utilizing statistical clustering methods.}
+  \end{enumerate} }
 
-To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
-The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
-As stated by DeDonno \cite{dedonno}, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
+\db{ The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
+Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
+The graphical representations including a parallel coordinate plot and geographic maps.[Figure 1].
+The Rural Smart EDA Dashboard map visuals were built using osrm route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.}
 
-A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
-Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
-Our coordinate plots will display variables categorically and numerically by scaling the data.
+\db{The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of the data, with constraints, to similar towns.
+The Rural Smart EDA Dashboard features offer geospatial and  structural data to assist town leaders in gaining a comprehensive understanding of the town in which they support.
+}
 
+\db{ Each component provides the town leaders with analytical aspects related to the performance of their town.
+As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=75mm]{SCC_Dashboard_pcp2.png}
+\caption{Figure 1: Rural Smart EDA Dashboard Design}
+\end{figure}
+
+\subsection{Statistical Methods}
 We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
@@
 
+%, in which our dashboard design has been used to display a panel with vital statistics, ensuring that the town leaders of a small town is the center of focus on the dashboard.
+The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
+To complement the Rural Shrink Smart NSF Grant, our dashboard includes Quality of Life (QoL) Metrics, to help identify communities social networks along with financial health data collected to ensure a robust identification of our small towns.
+
+\db{ To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
+The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
+As stated by DeDonno \cite{dedonno}, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
+}
+
+A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots will display variables categorically and numerically by scaling the data.
+
 Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
@@
 
-In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
-A table populates in the vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as the distance of fire department.
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=75mm]{SCC_Dashboard_pcp2.png}
-\caption{Rural Smart EDA Dashboard Design}
-\end{figure}
-
-%To-Do: Write this in the way that we have done this because by the time that we present this will be done. 
 %\subsection{Future Work}

### 2022-01-18 SCCAbstract_SDSS2022-knitr.tex
Updates

@@
 
-Some communities continue to thrive as they lose population because they adapt\svp{, maintaining quality of life and community services for residents while} investing in the future.
+Some communities continue to thrive as they lose population because they adapt, maintaining quality of life and community services for residents while investing in the future.
 % putting sentences on different lines helps with git merges and also makes it easier to figure out where {} match up :)
-\svp{This process,} \emph{smart shrinkage}\svp{, is important for rural areas who have experienced shrinking populations for decades.} 
-\svp{As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our} research team \svp{will provide communities with data about services in small town Iowa in order to assist with} developing strategies to improve quality of life for their residents amid shrinking populations \cite{scc}.
-\svp{We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
-In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.}
+This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades. 
+As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \cite{scc}.
+We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
+In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
 % Our work focuses on developing and challenges novice analysts in advanced statistical concepts and data visualizations to help make decisions.
@@
 \section{Data Description}
-\svp{Data collected from \url{data.iowa.gov}} were used to create the SCC dashboard.
-\svp{Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.} \svp{\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.}
-\svp{Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
-In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.}
+Data collected from \url{data.iowa.gov} were used to create the SCC dashboard.
+Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.
+\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
+In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.
 
-\svp{One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.}
-\svp{Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
-This makes both visualization and statistical analysis more complicated.}
+One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
+Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
+This makes both visualization and statistical analysis more complicated.
 
 \section{Guiding Design Principles}
-\svp{An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
+An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
-With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.}
-\svp{Thus, our initial set of dashboard design principles is as follows: }
+With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
+Thus, our initial set of dashboard design principles is as follows: 
 \begin{itemize}
-\item \svp{The town leaders are the focus audience; thus, the town itself should be the central focus of the app.}
-\item \svp{Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.}
-\item \svp{Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.}
-\item \svp{Allow for more detailed data and feature requests to improve the dashboard design over time.}
+\item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
+\item Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+\item Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+\item Allow for more detailed data and feature requests to improve the dashboard design over time.
 \end{itemize}
@@
 \section{Current Progress}
-\svp{We leverage} the framework of Guided Discovery Learning (GDL) to guide \svp{the town leaders} to \svp{make discoveries using our interactive visualization}.
-\svp{This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration}\cite{dedonno}.
+We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
+This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \cite{dedonno}.
 % Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.
@@
 
-\svp{We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.}
+We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
 
-\svp{As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
 We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \cite{few}.}
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \cite{few}.
 
@@
 
-\db{The Rural Shrink Smart EDA dashboard incorporates two major components: 
-  \begin{enumerate}
-  \item{Data Visualization Interface component with GDL principles.} 
-  \item{Exploratory Data Analysis component utilizing statistical clustering methods.}
-  \end{enumerate} }
+The Rural Shrink Smart EDA dashboard incorporates two major components: Data Visualization Interface component with GDL principles; and Exploratory Data Analysis component utilizing statistical clustering methods.
 
-\db{ The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
+The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
 Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
-The graphical representations including a parallel coordinate plot and geographic maps.[Figure 1].
-The Rural Smart EDA Dashboard map visuals were built using osrm route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.}
+The graphical representations including a parallel coordinate plot and geographic maps [Figure 1].
+The Rural Smart EDA Dashboard map visuals were built using osrm route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
 
-\db{The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of the data, with constraints, to similar towns.
+The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of data, with constraints, to similar towns.
 The Rural Smart EDA Dashboard features offer geospatial and structural data to assist town leaders in gaining a comprehensive understanding of the town in which they support.
-}
 
-\db{ Each component provides the town leaders with analytical aspects related to the performance of their town.
-As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.}
+Each component provides the town leaders with analytical aspects related to the performance of their town.
+As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.
 
@@
 \centering
-\includegraphics[width=75mm]{SCC_Dashboard_pcp2.png}
-\caption{Figure 1: Rural Smart EDA Dashboard Design}
+\includegraphics[width=60mm]{SCC_Dashboard_pcp2.png}
+\caption{Rural Smart EDA Dashboard Design}
 \end{figure}
@@
 
-\db{ To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
+To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
 The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
-As stated by DeDonno \cite{dedonno}, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
-}
+As stated by DeDonno, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
 
-A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
-Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
-Our coordinate plots will display variables categorically and numerically by scaling the data.
+A parallel (coordinates) plot allowed our town leaders to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
+Our coordinate plots will display variables categorically and numerically by scaling the data to demonstrate differences in necessary services.
 
-Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
-Our team has captured valuable metrics on using the product allowing the quantitative and qualitative data collection to better user experience.
-This can promote challenges in towns feeling "studied" by people in cities outside of their own.
+Our dashboard development final stages incorporates GDL allowing a continual feedback loop from town leaders in our small subset of towns has helped provide our team with valuable notes that will be adaptive to all of the greater 999 small towns in Iowa.
+Our team has captured valuable metrics, related to traditional design techniques, on product usage allowing a combination of quantitative and qualitative data to better the user experience.
+%This can promote challenges in towns feeling "studied" by people in cities outside of their own.
 

### 2022-01-18 SCCAbstract_SDSS2022-knitr.Rnw
adjust sectioning and add user testing section

--- 2022-01-17 SCCAbstract_SDSS2022-knitr.Rnw
@@
 
-\section{Current Progress}
+% \section{Current Progress}
 \svp{We leverage} the framework of Guided Discovery Learning (GDL) to guide \svp{the town leaders} to \svp{make discoveries using our interactive visualization}.
@@
 
-Our current work incorporates the ideas of GDL, in which dashboard design has been used to display a panel with vital statistics, ensuring that the town's analyst small town is the center of focus.
-The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
-The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town.
-Based on the number of variables used to ensure that the dashboard is extensive and valuable, we implemented parallel coordinate plots to show comparisons between the focus small town with the five similar towns.
-A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
-Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
-Our coordinate plots will display variables categorically and numerically by scaling the data.
+\db{The Rural Shrink Smart EDA dashboard incorporates two major components: 
+  \begin{enumerate}
+  \item{Data Visualization Interface component with GDL principles.} 
+  \item{Exploratory Data Analysis component utilizing statistical clustering methods.}
+  \end{enumerate} }
 
-We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
+\db{ The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
+Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
+The graphical representations including a parallel coordinate plot and geographic maps.[Figure 1].
+The Rural Smart EDA Dashboard map visuals were built using osrm route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.}
+
+\db{The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of the data, with constraints, to similar towns.
+The Rural Smart EDA Dashboard features offer geospatial and  structural data to assist town leaders in gaining a comprehensive understanding of the town in which they support.
+}
+
+\db{ Each component provides the town leaders with analytical aspects related to the performance of their town.
+As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=75mm]{SCC_Dashboard_pcp2.png}
+\caption{Figure 1: Rural Smart EDA Dashboard Design}
+\end{figure}
+
+\section{User Testing}
+\svp{In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
+We examine which aspects of dashboard design ``common knowledge" hold up in practice, and assess why these principles may not hold in our use-case.}
+
+\svp{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge" design principles stemming from more controlled experiments hold up in real-world situations.
+}
+
+% \subsection{Statistical Methods}
+% We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
 We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
@@
 Our dashboard will hopefully make sure that the town's analysts are guided in the direction that will help the town's population shrink smart.
-
-Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
+% 
+% %, in which our dashboard design has been used to display a panel with vital statistics, ensuring that the town leaders of a small town is the center of focus on the dashboard.
+The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
+To complement the Rural Shrink Smart NSF Grant, our dashboard includes Quality of Life (QoL) Metrics, to help identify communities social networks along with financial health data collected to ensure a robust identification of our small towns.
+% 
+% \db{ To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
+The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
+As stated by DeDonno \cite{dedonno}, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
+}
+% 
+% A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots will display variables categorically and numerically by scaling the data.
+% 
+% 
+% Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
 Our team has captured valuable metrics on using the product allowing the quantitative and qualitative data collection to better user experience.
@@
 
-In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
-A table populates in the vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as the distance of fire department.
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=75mm]{SCC_Dashboard.png}
-\caption{Rural Smart EDA Dashboard Design}
-\end{figure}
-
-%To-Do: Write this in the way that we have done this because by the time that we present this will be done. 
 %\subsection{Future Work}

### 2022-01-18 SCCAbstract_SDSS2022-knitr.Rnw

@@
 \section{Data Description}
-\svp{Data collected from \url{data.iowa.gov}} were used to create the SCC dashboard.
-\svp{Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.} \svp{\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.}
-\svp{Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
-In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.}
+Data collected from \url{data.iowa.gov} were used to create the SCC dashboard.
+Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.
+\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
+In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.
 
-\svp{One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.}
-\svp{Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
-This makes both visualization and statistical analysis more complicated.}
+One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
+Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
+This makes both visualization and statistical analysis more complicated.
 
 \section{Guiding Design Principles}
-\svp{An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
+An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
-With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.}
-\svp{Thus, our initial set of dashboard design principles is as follows: }
+With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
+Thus, our initial set of dashboard design principles is as follows: 
 \begin{itemize}
-\item \svp{The town leaders are the focus audience; thus, the town itself should be the central focus of the app.}
-\item \svp{Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.}
-\item \svp{Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.}
-\item \svp{Allow for more detailed data and feature requests to improve the dashboard design over time.}
+\item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
+\item Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+\item Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+\item Allow for more detailed data and feature requests to improve the dashboard design over time.
 \end{itemize}
 
-% \section{Current Progress}
-\svp{We leverage} the framework of Guided Discovery Learning (GDL) to guide \svp{the town leaders} to \svp{make discoveries using our interactive visualization}.
-\svp{This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration}\cite{dedonno}.
+\section{Current Progress}
+We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
+This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \cite{dedonno}.
 % Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.
@@
 
-\svp{We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.}
+We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
 
-\svp{As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
 We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \cite{few}.}
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \cite{few}.
 
@@
 Our coordinate plots will display variables categorically and numerically by scaling the data.
-% 
-% 
-% Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa.
-Our team has captured valuable metrics on using the product allowing the quantitative and qualitative data collection to better user experience.
-This can promote challenges in towns feeling "studied" by people in cities outside of their own.
+
+% Our dashboard development final stages incorporates GDL allowing a continual feedback loop from town leaders in our small subset of towns has helped provide our team with valuable notes that will be adaptive to all of the greater 999 small towns in Iowa.
+Our team has captured valuable metrics, related to traditional design techniques, on product usage allowing a combination of quantitative and qualitative data to better the user experience.
+%This can promote challenges in towns feeling "studied" by people in cities outside of their own.
 

### 2022-01-18 SCCAbstract_SDSS2022-knitr.tex

@@
 \definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\makeatletter
-\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}}}
-\makeatother
 \newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
@@
 \definecolor{errorcolor}{rgb}{1, 0, 0}
-\makeatletter
-\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-}}
-\makeatother
 \newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
@@
 
-As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
 We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
@@
 
-\subsection{Statistical Methods}
-We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
+\section{User Testing}
+\svp{In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
+We examine which aspects of dashboard design ``common knowledge" hold up in practice, and assess why these principles may not hold in our use-case.}
+
+\svp{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge" design principles stemming from more controlled experiments hold up in real-world situations.
+}
+
+% \subsection{Statistical Methods}
+% We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available.
 We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
@@
 Our dashboard will hopefully make sure that the town's analysts are guided in the direction that will help the town's population shrink smart.
-
-%, in which our dashboard design has been used to display a panel with vital statistics, ensuring that the town leaders of a small town is the center of focus on the dashboard.
+% 
+% %, in which our dashboard design has been used to display a panel with vital statistics, ensuring that the town leaders of a small town is the center of focus on the dashboard.
 The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools.
 To complement the Rural Shrink Smart NSF Grant, our dashboard includes Quality of Life (QoL) Metrics, to help identify communities social networks along with financial health data collected to ensure a robust identification of our small towns.
+% 
+% \db{ To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
+The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
+As stated by DeDonno \cite{dedonno}, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
+}
+% 
+% A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
+Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes.
+Our coordinate plots will display variables categorically and numerically by scaling the data.
 
-To restate, our dashboard incorporates an extensive number of variables used to ensure a valuable tool to town leaders, we implemented parallel coordinate plots to show comparisons between the user's small town with the five similar towns.
-The user has an option to select similarities, such as economy, location and size, allowing the town leaders the flexibility to guided their interests in the small town.
-As stated by DeDonno, our dashboard leverages hints for the town leaders to learn through feedback on the next steps to take in their towns.
-
-A parallel (coordinates) plot allowed our town leaders to compare the essential features of several individual towns on a set of numeric and categorical variables collected.
-Our coordinate plots will display variables categorically and numerically by scaling the data to demonstrate differences in necessary services.
-
-Our dashboard development final stages incorporates GDL allowing a continual feedback loop from town leaders in our small subset of towns has helped provide our team with valuable notes that will be adaptive to all of the greater 999 small towns in Iowa.
+% Our dashboard development final stages incorporates GDL allowing a continual feedback loop from town leaders in our small subset of towns has helped provide our team with valuable notes that will be adaptive to all of the greater 999 small towns in Iowa.
 Our team has captured valuable metrics, related to traditional design techniques, on product usage allowing a combination of quantitative and qualitative data to better the user experience.

### 2022-01-18 SCCAbstract_SDSS2022-knitr.Rnw
Incorporate language on statistical modeling and fix quotes

@@
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
+The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.
 
@@
 The graphical representations including a parallel coordinate plot and geographic maps [Figure 1].
-The Rural Smart EDA Dashboard map visuals were built using osrm route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
+The Rural Smart EDA Dashboard map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
 
@@
 \svp{In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
-We examine which aspects of dashboard design ``common knowledge" hold up in practice, and assess why these principles may not hold in our use-case.}
+We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.}
 
 \svp{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge" design principles stemming from more controlled experiments hold up in real-world situations.
+As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
 }

### 2022-01-18 SCCAbstract_SDSS2022-knitr.Rnw
acknowledgement

@@
 \section{Acknowledgements}
-We would like to thank our sponsors and grant provider NSF, LSAMP Inspire Program and Iowa League of Cities.
-We would like to thank our partners and team members at Iowa State University and University of Nebraska - Lincoln.
-We would like to thank the State of Iowa for providing Data along with our partners in the Rural Iowa communities.
+This material is based upon work supported by the National Science Foundation under Grant No.
+1952007.
+We would also like to thank our collaborators at Iowa State, the Iowa League of Cities, and the various Iowa government organizations which publish their data at \url{data.iowa.gov}.
 

### 2022-03-16 SCCAbstract_SDSS2022-knitr.Rnw
Updating Github documents

@@
 \centering
-\includegraphics[width=60mm]{SCC_Dashboard_pcp2.png}
+\includegraphics[width=55mm]{SCC_Dashboard_pcp2.png}
 \caption{Rural Smart EDA Dashboard Design}
@@
 \section{User Testing}
-\svp{In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
-We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.}
+In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
+We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.
 
-\svp{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
 As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
-}
 

### 2022-04-08 Graphics Group/Presentation/Graphics Group 4.7.2022.Rmd
Adding  Graphics Group Presentation

---
title: "Graphics Group Presentation"
author: Denise Bradford
date: April 7, 2022
output: 
  revealjs::revealjs_presentation:
    theme: sky
    highlight: pygments
    center: true
    transition: fade
    background_transition: slide
    self_contained: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## Project Information

Some communities continue to thrive as they lose population because they adapt and stay focused on quality of life, community services, and investing in the future. This is what we call rural smart shrinkage.

![](Shrink_Smart_logo.png)

(https://ruralshrinksmart.org/)

## Motivation
With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry. A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.

## Data Description
Data collected from https://data.iowa.gov were used to create the SCC dashboard. Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution. https://data.iowa.gov contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

## Guiding Design Principles

Our initial set of dashboard design principles is as follows:

- The town leaders are the focus audience (central focus of the app)

- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions

- Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.

- Allow for more detailed data and feature requests to improve the dashboard design over time.

## Guided Discovery Learning with Dashboard
We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration

## Dashboard Visual Iterations

- It should be worth noting that the dashboard design is an iterative process that starts with an idea without limitations of the actual implementation pieces.
- After implementation has been completed and the team notes limitations, we change the interface.
- Now incorporate more literature on dashboard design and user experience, which removes some of the "cool" implementations

## The Original Mockup 
![](SCC_Dashboard_pcp2.png)

## Back to the drawing board
![](SCC Dashboard Markup 1.png)

## Back to the drawing board... AGAIN
![](SCC Dashboard Markup 2.png)

## Dashboard Visual Right Now

### 2022-04-08 Abstract Files/SCCAbstract_SDSS2022-knitr.tex
Cleaning up the repo

--- 2022-01-18 SCCAbstract_SDSS2022-knitr.tex
@@
 \section{User Testing}
-\svp{In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
-We examine which aspects of dashboard design ``common knowledge" hold up in practice, and assess why these principles may not hold in our use-case.}
+In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
+We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.
 
-\svp{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge" design principles stemming from more controlled experiments hold up in real-world situations.
-}
+Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
 

### 2022-05-30 SDSS Presentation/SDSS Presentation.Rmd
Adding presentation for SDSS

---
title: "Exploring Rural Shrink Smart Through Guided Discovery Dashboards"
author: "Denise R. Bradford and Susan R. Vanderplas, PhD"
date: "June 8, 2022"
output:
  xaringan::moon_reader:
    css: ["default", "unl.css"]
    lib_dir: libs
    includes:
      in_header: header.html
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class:inverse
## About Me

- PhD student in Statistics at University of Nebraska - Lincoln
- MS/BS in Applied Statistics from Purdue University

- Research Areas
  - Data Science - automation, data pipelines, tools
  - Visualization - perception, design, and implementation
  - Sports Analytics - deploying useful statistical methods for sports-related problems
---

## Rural Smart Shrinkage

<!-- Insert the shrink smart Logo -->

- Rural towns are losing population steadily

  - Shrinking tax bases lead to fewer services
  - Shrinking population may shrink social interaction opportunities
  - Schools, businesses, etc. may move to nearby towns

--

- .red[Project Question:] How do some cities manage to maintain quality of life for residents amid shrinking populations?
  - Can this be taught?

---
## Complications

- Quality of Life is a subjective assessment
  - Iowa Small Town poll has measured subjective QOL by decade for one small town in each of the 99 counties in Iowa (same town each year)
  - Survey response rates keep decreasing over time

- There are objective measures of things that contribute to QOL available in public datasets
.small.emph[
- School ratings
- Town budgets for services
- Access to medical care, shopping, etc.
- Transportation trends
- Demographic shifts
- City level cooperative agreements with other government entities and NGOs
]

---
## Complications

- Quality of Life is a subjective assessment
  - Iowa Small Town poll has measured subjective QOL by decade for one small town in each of the 99 counties in Iowa (same town each year)
  - Survey response rates keep decreasing over time

- There are objective measures of things that contribute to QOL available in public datasets

- .red[Project Question:] How do we leverage .emph.cerulean[data science] and .emph.cerulean[survey design] to reduce the length of the survey and maintain reliability?
    - Can objective measures of QOL contributors be incorporated into the survey design?
    - Can we focus on estimating only things that objective measurements don't capture?

---
## Basic Idea

- Assemble publicly available data sets
    - Available over lots of different geographic measurements    
    Lat/Long, Zip code (5 or 9 digit, changes over time), Address, City, Census units, School District, County
    - May not be complete for small towns    
    (ACS estimates aren't useful in these areas - not enough people surveyed)

- Assess correlation between public data and Iowa Small Town Poll QOL estimates on different dimensions

- .emph.cerulean[Survey Stats problems]: 
    - Incorporate this information into survey designs
    - Change up design of the survey to split questions across participants, etc.

---
## Guiding Dashboard Design Principles

- The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
- Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
- Allow for more detailed data and feature requests to improve the dashboard design over time.

--

We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration

---
## Statistical Analysis

- Compile a single data set
    - rows: Each town
    - columns: Each variable with appropriate spatial aggregation    
    e.g. distance to service instead of lat/long

- Unsupervised analysis
    - Which towns are most similar based only on the public data?

- Supervised analysis
    - Which variables are most associated with subjectively reported quality of life (in the 99 towns in the ISTP)?

---
## Visualization

- Create a town-focused dashboard that lets leaders see how their town compares to other similar towns (geographically and as measured by the unsupervised statistical analysis)

- Goal is to help towns see what strategies for maintaining QOL work and don't

- Emphasis on things towns are empowered to change
    - Can't focus on e.g. agricultural policy
    - City budgets, partnerships with outside entities, services that matter and don't

- How do we best design charts/graphs/UI to make people feel empowered and help them explore the data with an open mind?
    - How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.

### 2022-05-30 SDSS Presentation/SDSS Presentation.Rmd
styling updates

@@
   xaringan::moon_reader:
-    css: ["default", "unl.css"]
+    seal: false
+    includes:
+      after_body:
+        "js-addins.html"
+    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
+    css: ["default", "metropolis-fonts", "metropolis" ,"css/modal.css", "css/sizeformat.css"]
     lib_dir: libs
-    includes:
-      in_header: header.html
     nature:
       highlightStyle: github
-      highlightLines: true
-      countIncrementalSlides: false
+      highlightlines: true
+      countIncrementalSlides: true
 ---
+class:title-slide-custom
 
-```{r setup, include=FALSE}
-options(htmltools.dir.version = FALSE)
+```{r, child = "style.Rmd"}
 ```
 
-class:inverse
-## About Me
+```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
+# Packages
+library(emoji)
+library(purrr)
+library(tidyverse)
+library(gridExtra)
+library(nullabor)
+library(scales)
+library(knitr)
+library(kableExtra)
+library(RefManageR)
+library(iconr)
+library(fontawesome)
+# download_fontawesome()
 
-- PhD student in Statistics at University of Nebraska - Lincoln
-- MS/BS in Applied Statistics from Purdue University
+# References
+bib <- ReadBib("bib/references.bib", check = FALSE)
+ui <- "- "
 
-- Research Areas
-  - Data Science - automation, data pipelines, tools
-  - Visualization - perception, design, and implementation
-  - Sports Analytics - deploying useful statistical methods for sports-related problems
+# R markdown options
+knitr::opts_chunk$set(echo = FALSE, 
+                      message = FALSE, 
+                      warning = FALSE, 
+                      cache = FALSE,
+                      dpi = 300)
+options(htmltools.dir.version = FALSE)
+options(knitr.kable.NA = '')
+```
+
+```{r, include = F, eval = T, cache = F}
+clean_file_name <- function(x) {
+  basename(x) %>% str_remove("\\..*?$") %>% str_remove_all("[^[A-z0-9_]]")
+}
+img_modal <- function(src, alt = "", id = clean_file_name(src), other = "") {
+
+  other_arg <- paste0("'", as.character(other), "'") %>%
+    paste(names(other), ., sep = "=") %>%
+    paste(collapse = " ")
+
+  js <- glue::glue("<script>
+        /* Get the modal*/
+          var modal{id} = document.getElementById('modal{id}');
+        /* Get the image and insert it inside the modal - use its 'alt' text as a caption*/
+          var img{id} = document.getElementById('img{id}');
+          var modalImg{id} = document.getElementById('imgmodal{id}');
+          var captionText{id} = document.getElementById('caption{id}');
+          img{id}.onclick = function(){{
+            modal{id}.style.display = 'block';
+            modalImg{id}.src = this.src;
+            captionText{id}.innerHTML = this.alt;
+          }}
+          /* When the user clicks on the modalImg, close it*/
+          modalImg{id}.onclick = function() {{
+            modal{id}.style.display = 'none';
+          }}
+</script>")
+
+  html <- glue::glue(
+     " <!-- Trigger the Modal -->
+<img id='img{id}' src='{src}' alt='{alt}' {other_arg}>
+<!-- The Modal -->
+<div id='modal{id}' class='modal'>
+  <!-- Modal Content (The Image) -->
+  <img class='modal-content' id='imgmodal{id}'>
+  <!-- Modal Caption (Image Text) -->
+  <div id='caption{id}' class='modal-caption'></div>
+</div>
+"
+  )
+  write(js, file = "js-addins.html", append = T)
+  return(html)
+}
+# Clean the file out at the start of the compilation
+write("", file = "js-addins.html")
+```
+
+<br><br>
+# Exploring Rural Shrink Smart Through Guided Discovery Dashboards
+## SDSS 2022
+### June 8, 2022
+#### Denise Bradford
+#### Department of Statistics, University of Nebraska - Lincoln 
+##### `r fa("envelope", fill = "black")` [denise.bradford@huskers.unl.edu](denise.bradford@huskers.unl.edu) 
+##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/) 
+##### `r fa("github", fill = "black")` [drbradford12](https://github.com/drbradford12)
+
+???
+Thank you for inviting me for a visit today and for attending my talk.
+
+---
+class:primary
+# Outline
+
++ Introduction to Rural Shrink Smart Project with Iowa State
+
++ Why Dashboards?
+
++ Dashboard Design and Guided Discovery Learning (GDL)
+
++ Parallel Coordinate Plots (PCPs) and the naive user
+
++ Data Collection and Feedback the loop
+
++ Future Work
+
+???
+In this talk today, I will provide a background of the Rural Shrink Smart Project.
+I will then provide a framework to way using a dashboard can be helpful to these population discovery of gaining insights from other towns.
+I will continue to discuss the usefulness of parallel coordinate plots on the naive user.
+I will conclude with future work.
+
 ---
@@
 - There are objective measures of things that contribute to QOL available in public datasets
-.small.emph[
+.medium.emph[
 - School ratings
@@
 
-- .red[Project Question:] How do we leverage .emph.cerulean[data science] and .emph.cerulean[survey design] to reduce the length of the survey and maintain reliability?
-    - Can objective measures of QOL contributors be incorporated into the survey design?
+- .red[Project Question:] How do we leverage .emph.cerulean[data science] and .emph.cerulean[exploratory data methods] to inform town leaders to reduce the shrinkage and maintain reliability?
+    - Can traditional dashboard design of open source data be incorporated into overall decision-making?
     - Can we focus on estimating only things that objective measurements don't capture?
@@
 
-- Assess correlation between public data and Iowa Small Town Poll QOL estimates on different dimensions
+- Assess usability, intelligibility and adaptability to new statistical methodology.
 
-- .emph.cerulean[Survey Stats problems]: 
-    - Incorporate this information into survey designs
-    - Change up design of the survey to split questions across participants, etc.
+- .emph.cerulean[Exploratory Data Analysis (EDA)]: 
+    - Incorporate this information into an exploratory dashboard
+    - Change up design of the dashboard to manage the easy of adaptability across our rural towns
 

### 2022-05-30 SDSS Presentation/SDSS Presentation.Rmd
Updates for 5.30.2022

@@
     includes:
+      in_footer: footer.html
       after_body:
@@
 ##### `r fa("envelope", fill = "black")` [denise.bradford@huskers.unl.edu](denise.bradford@huskers.unl.edu) 
-##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/) 
+<!-- ##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/)  -->
 ##### `r fa("github", fill = "black")` [drbradford12](https://github.com/drbradford12)
 
+<br><br><br>
+.medium.emph[
+Theme designed by Emily A. Robinson [www.emilyarobinson.com](https://www.emilyarobinson.com/)
+]
 ???
@@
 
-+ Introduction to Rural Shrink Smart Project with Iowa State
++ Introduction to Rural Shrink Smart Project with Iowa State University
 
@@
 
-## Rural Smart Shrinkage
-
-<!-- Insert the shrink smart Logo -->
+# Rural Shrink Smart Project
+.center[
+![](Shrink_Smart_logo.png)
+]
+.center[
+(https://ruralshrinksmart.org/)
+]
 
@@
 
---
+---
+# Rural Shrink Smart Project
+.center[
+![](Shrink_Smart_logo.png)
+]
+.center[
+(https://ruralshrinksmart.org/)
+]
 
@@
 
+.medium.emph[
+This material is based upon work supported by the National Science Foundation under Grant No.
+1952007.
+We would also like to thank our collaborators at Iowa State, the Iowa League of Cities, and the various Iowa government organizations which publish their data at (https://data.iowa.gov).
+]
+
 ---
-## Complications
+# Complications
 
@@
 ---
-## Complications
+# Complications
 
@@
 ---
-## Basic Idea
+# Basic Idea
 
@@
 
-- Assess usability, intelligibility and adaptability to new statistical methodology.
+- Assess usability, interpretability and adaptability to new statistical methodology.
 
-- .emph.cerulean[Exploratory Data Analysis (EDA)]: 
-    - Incorporate this information into an exploratory dashboard
-    - Change up design of the dashboard to manage the easy of adaptability across our rural towns
+- .emph.cerulean[The Rural Shrink Smart EDA dashboard incorporates two major components]: 
+    - Data Visualization Interface component with Guided Discovery Learning (GDL) principles
+    - Exploratory Data Analysis component utilizing statistical clustering methods.
 
 ---
-## Guiding Dashboard Design Principles
+#
 
-- The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
-- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+---
+# Guiding Design Principles
+
+Our initial set of dashboard design principles are:
+
+- The town leaders are the focus audience (central focus of the app)
+
+- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions
+
 - Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+
 - Allow for more detailed data and feature requests to improve the dashboard design over time.
 
---
+---
+# Guided Discovery Learning with Dashboard Design
 
-We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
-This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration
+We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. By leveraging:
+
+- hints
+
+- feedback
+
+- other helpful information to guide users in interactive exploration
+    e.g. user ignoring the radio buttons to compare similar cities, etc.
 
 ---
-## Statistical Analysis
+# Dashboard Design Workflow
 
@@
 
-- Supervised analysis
-    - Which variables are most associated with subjectively reported quality of life (in the 99 towns in the ISTP)?
+- Dashboard layout
+    - Which layout will allow the user to engage the most without overwhelming their curiosity?
+
+- Dashboard Feedback
+    - What elements on the dashboard are useful, confusing, misleading or incorrect from town leaders?
 
 ---
-## Visualization
+# Visualization
 
@@
     - How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.
+
+---
+# Dashboard Visual Iterations
+
+- Version 1: An idea without limitations of the actual implementation pieces.
+
+- Version 2: After implementation has been completed and the team notes limitations, we change the interface.
+
+- Version 3: Now incorporate more literature on dashboard design and user experience, which removes some of the "cool" implementations
+
+- Version 4: Finally incorporate a collective feedback from team-leaders that suggests lack of understanding of information
+
+---
+
+#

### 2022-06-02 SDSS Presentation/SDSS Presentation.Rmd
Updates to slides 6.2.2022

@@
 
-<br><br><br>
+<br><br>
 .medium.emph[
@@
 
+<br>
+.large.emph[
 + Introduction to Rural Shrink Smart Project with Iowa State University
@@
 + Future Work
+]
 
@@
 I will conclude with future work.
-
----
-
-# Rural Shrink Smart Project
-.center[
-![](Shrink_Smart_logo.png)
-]
-.center[
-(https://ruralshrinksmart.org/)
-]
-
-- Rural towns are losing population steadily
-
-  - Shrinking tax bases lead to fewer services
-  - Shrinking population may shrink social interaction opportunities
-  - Schools, businesses, etc. may move to nearby towns
 
@@
 
-- .red[Project Question:] How do some cities manage to maintain quality of life for residents amid shrinking populations?
-  - Can this be taught?
++ <span style="color:darkred"> Project Question:</span> How do some cities manage to maintain quality of life for residents amid shrinking populations?
+  + Can this be taught?
 
@@
 ---
-# Complications
+# Basic Idea
 
-- Quality of Life is a subjective assessment
-  - Iowa Small Town poll has measured subjective QOL by decade for one small town in each of the 99 counties in Iowa (same town each year)
-  - Survey response rates keep decreasing over time
++ Assemble publicly available data sets
+    + Available over lots of different geographic measurements    
+    Lat/Long, Zip code (5 or 9 digit, changes over time), Address, City, Census units, School District, County
+    + May not be complete for small towns    
+    (ACS estimates aren't useful in these areas - not enough people surveyed)
 
-- There are objective measures of things that contribute to QOL available in public datasets
++ Assess usability, interpretability and adaptability to new statistical methodology.
+
++ .emph.cerulean[The Rural Shrink Smart EDA dashboard incorporates two major components]: 
+    + Data Visualization Interface component with Guided Discovery Learning (GDL) principles, which engages the town leaders. 
+    + Exploratory Data Analysis component utilizing statistical clustering methods, which allow town leaders the opportunity explore new elements to add or remove.
+
+---
+# Challenges
+
++ Quality of Life is a subjective assessment
+  + Iowa Small Town poll has measured subjective QOL by decade for one small town in each of the 99 counties in Iowa (same town each year)
+  + Survey response rates keep decreasing over time
+
++ There are objective measures of things that contribute to QOL available in public datasets
 .medium.emph[
-- School ratings
-- Town budgets for services
-- Access to medical care, shopping, etc.
-- Transportation trends
-- Demographic shifts
-- City level cooperative agreements with other government entities and NGOs
+  + School ratings
+  + Town budgets for services
+  + Access to medical care, shopping, etc.
+  + Transportation trends
+  + Demographic shifts
+  + City level cooperative agreements with other government entities and NGOs
 ]
@@
 ---
-# Complications
+# Challenges
 
-- Quality of Life is a subjective assessment
-  - Iowa Small Town poll has measured subjective QOL by decade for one small town in each of the 99 counties in Iowa (same town each year)
-  - Survey response rates keep decreasing over time
++ Quality of Life is a subjective assessment
+  + Iowa Small Town poll has measured subjective QOL by decade for one small town in each of the 99 counties in Iowa (same town each year)
+  + Survey response rates keep decreasing over time
 
-- There are objective measures of things that contribute to QOL available in public datasets
++ There are objective measures of things that contribute to QOL available in public datasets
 
-- .red[Project Question:] How do we leverage .emph.cerulean[data science] and .emph.cerulean[exploratory data methods] to inform town leaders to reduce the shrinkage and maintain reliability?
-    - Can traditional dashboard design of open source data be incorporated into overall decision-making?
-    - Can we focus on estimating only things that objective measurements don't capture?
++ <span style="color:darkred">Project Question:</span> How do we leverage <span style="color:blue">data science</span>  and <span style="color:blue">exploratory data methods</span> to inform town leaders to reduce the shrinkage and maintain reliability?
+    + Can traditional dashboard design of open source data be incorporated into overall decision-making?
+    + Can we focus on estimating only things that objective measurements don't capture?
 
 ---
-# Basic Idea
 
-- Assemble publicly available data sets
-    - Available over lots of different geographic measurements    
-    Lat/Long, Zip code (5 or 9 digit, changes over time), Address, City, Census units, School District, County
-    - May not be complete for small towns    
-    (ACS estimates aren't useful in these areas - not enough people surveyed)
+# Identifying Key Metrics
 
-- Assess usability, interpretability and adaptability to new statistical methodology.
+<b>Audience</b>
 
-- .emph.cerulean[The Rural Shrink Smart EDA dashboard incorporates two major components]: 
-    - Data Visualization Interface component with Guided Discovery Learning (GDL) principles
-    - Exploratory Data Analysis component utilizing statistical clustering methods.
+.medium.emph[
++ As small rural towns, traditionally, do not have access:
+     + Data scientists 
+     + Ability to easily leverage data collected locally to support decisions
+
+Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations
+]
+--
+
+<b>Visualization Goals</b>
+
+.medium.emph[
++ Create a central place for the excess of data to be conveyed in a single comprehensive visual.
+    + We will utilize Parallel Coordinate Plot (PCPs)
++ Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g. Fire Dept, Schools, etc.)
++ Add an overview of town general basics related to the seven QoL factors.
+]
+--
+
+<b>Engage People</b>
+
+.medium.emph[
++ We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+]
 
 ---
-#
+# Visualization
+
++ Create a town-focused dashboard that lets leaders see how their town compares to other similar towns (geographically and as measured by the unsupervised statistical analysis)
+
++ Goal is to help towns see what strategies for maintaining QOL work and don't
+
++ Emphasis on things towns are empowered to change
+    + Can't focus on e.g. agricultural policy
+    + City budgets, partnerships with outside entities, services that matter and don't
+--
+
++ How do we best design charts/graphs/UI to make people feel empowered and help them explore the data with an open mind?
+    + How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.
 
@@
 
-- The town leaders are the focus audience (central focus of the app)
++ The town leaders are the focus audience (central focus of the app)
+    + Local Map (Local Essential Services)
+    + Vital Statistics
 
-- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions
++ Facilitate comparisons with other towns in order to allow the user to explore other potential solutions
 
-- Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
++ Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+    + PCP to compare the towns with varying similar features (e.g. economy, population, rurality or user input)
 
-- Allow for more detailed data and feature requests to improve the dashboard design over time.
++ Allow for more detailed data and feature requests to improve the dashboard design over time.
 
 ---
+
+# Dashboard Design Workflow
+
++ Compile a single data set
+    + rows: Each town
+    + columns: Each variable with appropriate spatial aggregation    
+    e.g. distance to service instead of lat/long
+
++ Unsupervised Analysis
+    + Which towns are most similar based only on the public data?
+
++ Dashboard Layout
+    + Which layout will allow the user to engage the most without overwhelming their curiosity?
+
++ Dashboard Feedback
+    + What elements on the dashboard are useful, confusing, misleading or incorrect from town leaders?
+
+---
+# Unsupervised Analysis
+<br>
+
+- Unsupervised clustering methods were used to determine data quality
+
+--
+
+<br>
+-  During the process, we found that schools, fire departments, post offices and hospitals were missing
+
+--
+
+<br>
+-  Resulted in finding other open source datasets that were used for sanity checks along with adding information of missing data
+
+???
+
+We found that in towns that the project leaders understood were off.
+For example, we found that many of the schools that were known to be in the towns were not in the dataset that we were using.
+Along with the fact that we have the leaflet feature, we could see that schools were in particular parts of towns that we did not see.
+
+---
+
+# Dashboard Layout Iterations
+
+.large.emph[
++ Version 1: An idea without limitations of the actual implementation pieces.
+]
+
+--
+
+<br>
+.large.emph[
++ Version 2: After implementation has been completed and the team notes limitations, we change the interface.
+]
+
+--
+
+<br>
+.large.emph[
++ Version 3: Now incorporate more literature on dashboard design and user experience, which removes some of the "cool" implementations 
+]
+
+--
+
+<br>
+.large.emph[
++ Version 4: Finally incorporate a collective feedback from team-leaders that suggests lack of understanding of information
+]
+
+---
+
+# Dashboard Components
+
+.large.emph[
++ Map (Town's closest services)
+]
+
+--
+ <br>
+
+.large.emph[
++ Vital Statistics (Town-centric and county level)
+]
+--
+ <br>
+
+.large.emph[
++ Paralllel Coordinate Plot with objective measures related to the QoL Survey
+]
+
+---
+
+# Reasoning behind using PCP
+
+<br>
+
+.large.emph[
++ We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. 
+]
+
+--
+<br>
+
+.large.emph[
++ We allow users to modify a set of suggested comparison towns to indicate other towns of interest.
+]
+---
+class:primary
+
+# PCPs Effectiveness & Novice Users
+
+**Who?** Sukwon Lee, et al. wrote How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking
+
+**What?** Propose a grounded model of `NO`vice’s information `VI`sualization `S`ensemaking (`NOVIS` model)
++ Framework for assessing a person's ability to interpret a new graphical map
+    + Two Basic Components:
+      + Visual Object (textual objects and non-textual objects)
+      + Frame (frame of content and frame of visual encoding)
+
+**Why?** Advantages of using NOVIS model includes a reproducible framework of understanding graphics.
+<br><br>
+[Reference Journal](https://ieeexplore.ieee.org/document/7192668)`r Citep(bib[[c("lee et al")]])`
+
+???
+
+NOVIS Model has been used to compare other graphs such as PCP, Tree Maps and Chord diagrams, which helps with the understanding of why we could use the PCPs.
+We will use this model to ask questions of our town leaders to gain feedback on what they understand.
+---
+
+# PCP in Dashboard Develop Reasoning
+<br>
++ The large set of publicly available data we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+
+  + A small subset of variables of interest to start with, and then
+
+  + Allow the user to increase the complexity of the display in accordance with their interest
+
++ This avoids some of the pitfalls of dashboard design that can easily lead to user overload [Few](https://www.stephen-few.com/) `r Citep(bib[[c("few")]])`
+
+---
+class:primary
+
+# Background of Guided Discovery Learning
+
+**Who?** [Michael A. DeDonno](https://doi.org/10.1016/j.lindif.2016.05.023) wrote The influence of IQ on pure discovery and guided discovery learning of a complex real-world task.
+
+**What?** To investigate the predictive ability of IQ on pure discovery and guided discovery learning of a complex real-world task.
+
+**How is this useful?** As mentioned previously, we not only want to present a complex graphic for novice users, we also want to empower the town-leader users to explore the dashboard, comprehensively.
+We would like to use the general principles by making sure that the users understands what is being presented to them.
+
+**Experimentation** We will collect how the users are clicking on the app. This will help us understand if the users are getting better understanding of the app over time.
+
+???
+
+---
+
 # Guided Discovery Learning with Dashboard Design
-
+<br>
 We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. By leveraging:
 
-- hints
++ hints
 
-- feedback
++ feedback
 
-- other helpful information to guide users in interactive exploration
++ other helpful information to guide users in interactive exploration
     e.g. user ignoring the radio buttons to compare similar cities, etc.
@@
 ---
-# Dashboard Design Workflow
+# User Testing in Beta
 
-- Compile a single data set
-    - rows: Each town
-    - columns: Each variable with appropriate spatial aggregation    
-    e.g. distance to service instead of lat/long
+**Audience**
 
-- Unsupervised analysis
-    - Which towns are most similar based only on the public data?
++ Groups of Users:
+    + Graphics Group Members (Users know Statistical Graphics)
 
-- Dashboard layout
-    - Which layout will allow the user to engage the most without overwhelming their curiosity?
+    + Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics)
 
-- Dashboard Feedback
-    - What elements on the dashboard are useful, confusing, misleading or incorrect from town leaders?
+    + Others (Member of doctoral community that has background in educational adaptivity)
 
 ---
-# Visualization
+# Dashbaord Feedback
 
-- Create a town-focused dashboard that lets leaders see how their town compares to other similar towns (geographically and as measured by the unsupervised statistical analysis)
++ Our feedback from users has two components:
 
-- Goal is to help towns see what strategies for maintaining QOL work and don't
+   + User prompted (using a survey and screenshots)
 
-- Emphasis on things towns are empowered to change
-    - Can't focus on e.g. agricultural policy
-    - City budgets, partnerships with outside entities, services that matter and don't
+   + User behaviors being record in the background:
 
-- How do we best design charts/graphs/UI to make people feel empowered and help them explore the data with an open mind?
-    - How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.
+      + This will help validate issues that are being mentioned in the survey.
+
+      + e.g. "I'm having a hard time understanding the PCP information", we will see that users are not spending any time on the PCP plot.
+---
+class:primary
+
+# Future Work
+
++ Allow the....???
 
 ---
-# Dashboard Visual Iterations
+class:inverse
+<br>
+<br>
+<br>
+.center[
+# Thank you! 
+### Acknowledgments: Dr. Susan VanderPlas and the ISU - Rural Shrink Smart Project Team
+<br>
+<br>
+`r fa("envelope", fill = "white")` **denise.bradford@huskers.unl.edu**
+<!-- ##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/)  -->
+`r fa("github", fill = "black")` **@drbradford12**
+]
 
-- Version 1: An idea without limitations of the actual implementation pieces.
+???
 
-- Version 2: After implementation has been completed and the team notes limitations, we change the interface.
-
-- Version 3: Now incorporate more literature on dashboard design and user experience, which removes some of the "cool" implementations
-
-- Version 4: Finally incorporate a collective feedback from team-leaders that suggests lack of understanding of information
-
----
-
-#
+Thank you for the support of my adviser Dr. Susan VanderPlas and the ISU Rural Shrink Smart

### 2022-06-02 SDSS Presentation/SDSS Presentation.Rmd
format

@@
 ---
-# Unsupervised Analysis
+# Dashboard Design Workflow
+
+## Unsupervised Analysis
 <br>
@@
 ---
+# Dashboard Design Workflow (Layout)
 
-# Dashboard Layout Iterations
+## Dashboard Layout Iterations
 
@@
 ---
+# Dashboard Design Workflow (Layout)
 
-# Dashboard Components
+## Dashboard Components
 
@@
 ---
+# Dashboard Philosophy
 
-# Reasoning behind using PCP
+## Reasoning behind using PCP
 
@@
 
-# PCPs Effectiveness & Novice Users
+# Dashboard Philosophy
+
+## PCPs Effectiveness & Novice Users
 
@@
 ---
+# Dashboard Philosophy
 
-# PCP in Dashboard Develop Reasoning
+## PCP in Dashboard Develop Reasoning
+
 <br>
@@
 class:primary
+# Dashboard Philosophy
 
-# Background of Guided Discovery Learning
+## Background of Guided Discovery Learning
 
@@
 ---
+# Dashboard Philosophy
 
-# Guided Discovery Learning with Dashboard Design
+## Guided Discovery Learning with Dashboard Design
 <br>
@@
 ---
-# User Testing in Beta
+# Dashboard Design WorkFlow (Feedback)
+
+## User Testing in Beta
 
@@
 ---
-# Dashbaord Feedback
+# Dashboard Design WorkFlow (Feedback)
+
+## Dashbaord Feedback
 

### 2022-06-05 SDSS Presentation/SDSS Presentation.Rmd
Screenshot updates and Susan comments

@@
 .center[
-![](Shrink_Smart_logo.png)
-]
-.center[
-(https://ruralshrinksmart.org/)
+<img src="images/Shrink_Smart_logo.png" width="35%"/>
 ]
@@
 + .emph.cerulean[The Rural Shrink Smart EDA dashboard incorporates two major components]: 
-    + Data Visualization Interface component with Guided Discovery Learning (GDL) principles, which engages the town leaders. 
-    + Exploratory Data Analysis component utilizing statistical clustering methods, which allow town leaders the opportunity explore new elements to add or remove.
+    + Data Visualization Interface component with Guided Discovery Learning (GDL) principles
+    + Exploratory Data Analysis component utilizing statistical graphical elements, such as maps and Parallel Coordinate Plots (PCPs)
 
@@
 + <span style="color:darkred">Project Question:</span> How do we leverage <span style="color:blue">data science</span>  and <span style="color:blue">exploratory data methods</span> to inform town leaders to reduce the shrinkage and maintain reliability?
-    + Can traditional dashboard design of open source data be incorporated into overall decision-making?
+    + How does an interactive dashboard of open-source data integrate into overall decision-making?
     + Can we focus on estimating only things that objective measurements don't capture?
@@
 
---
+.center[
+<img src="images/version1.png" width="75%"/>
+]
 
-<br>
+---
+# Dashboard Design Workflow (Layout)
+
+## Dashboard Layout Iterations
+
 .large.emph[
@@
 
---
+.center[
+<img src="images/version2.png" width="75%"/>
+]
 
-<br>
+---
+# Dashboard Design Workflow (Layout)
+
+## Dashboard Layout Iterations
+
 .large.emph[
@@
 
---
+.center[
+<img src="images/version3.png" width="50%"/>
+]
 
-<br>
+---
+# Dashboard Design Workflow (Layout)
+
+## Dashboard Layout Iterations
+
 .large.emph[
 + Version 4: Finally incorporate a collective feedback from team-leaders that suggests lack of understanding of information
+]
+
+.center[
+<img src="images/version4.png" width="75%"/>
 ]
@@
 .large.emph[
-+ Paralllel Coordinate Plot with objective measures related to the QoL Survey
++ Parallel Coordinate Plot with objective measures related to the QoL Survey
 ]
@@
 
-      + e.g. "I'm having a hard time understanding the PCP information", we will see that users are not spending any time on the PCP plot.
+      + e.g. "I'm having a hard time understanding the PCP information", we will see that users are not spending any time on the PCP.
+
+.center[
+<img src="images/feedback_question.png" width="35%"/>
+]
 ---
@@
 
-+ Allow the....???
++ Continue to incorporate feedback from town leaders.
+<br>
+.center[
+<img src="images/feedback_loop.png" width="50%"/>
+]
 

### 2022-06-06 SDSS Presentation/SDSS Presentation.Rmd
Updates with screenshots

@@
 
-.medium.emph[
+.small.emph[
 + As small rural towns, traditionally, do not have access:
@@
 
-.medium.emph[
+.small.emph[
 + Create a central place for the excess of data to be conveyed in a single comprehensive visual.
@@
 
-.medium.emph[
+.small.emph[
 + We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
@@
 .center[
-<img src="images/version1.png" width="75%"/>
+<img src="images/version1.png" width="80%"/>
 ]
@@
 .center[
-<img src="images/version2.png" width="75%"/>
+<img src="images/version2.png" width="52%"/>
 ]
@@
 .center[
-<img src="images/version3.png" width="50%"/>
+<img src="images/version3.png" width="60%"/>
 ]
@@
 
-.large.emph[
-+ Map (Town's closest services)
+- Map (Town's closest services)
+
+.center[
+<img src="images/map.png" width="60%"/>
 ]
 
---
- <br>
+---
+# Dashboard Design Workflow (Layout)
 
-.large.emph[
-+ Vital Statistics (Town-centric and county level)
-]
---
- <br>
+## Dashboard Components
 
-.large.emph[
-+ Parallel Coordinate Plot with objective measures related to the QoL Survey
+- Vital Statistics (Town-centric and county level)
+
+.column[
+<img src="images/vital1.png" width="50%"/>
+<img src="images/vital2.png" width="50%"/>
 ]
 
+---
+# Dashboard Design Workflow (Layout)
+
+## Dashboard Components
+
+- Parallel Coordinate Plot with objective measures related to the QoL Survey
+
+.center[
+<img src="images/pcp.png" width="75%"/>
+]
+
+[Smart Connected Communities Dashbaord](https://shiny.srvanderplas.com/Smart-Connected-Communities/)
 ---

### 2022-06-06 SDSS Presentation/SDSS Presentation.Rmd
Updates to images

@@
 - Vital Statistics (Town-centric and county level)
-
+.center[
 .column[
-<img src="images/vital1.png" width="50%"/>
-<img src="images/vital2.png" width="50%"/>
+<img src="images/vital1.png" width="45%"/>
+<img src="images/vital2.png" width="45%"/>
+]
 ]
@@
 + Our feedback from users has two components:
-
    + User prompted (using a survey and screenshots)
-
    + User behaviors being record in the background:
-
       + This will help validate issues that are being mentioned in the survey.
-
       + e.g. "I'm having a hard time understanding the PCP information", we will see that users are not spending any time on the PCP.
@@
 .center[
-<img src="images/feedback_question.png" width="35%"/>
+<img src="images/feedback_question.png" width="44%"/>
 ]

### 2022-06-06 SDSS Presentation/SDSS Presentation.Rmd
Updates from Brian adding break slides

@@
 
-<br><br>
-.medium.emph[
-Theme designed by Emily A. Robinson [www.emilyarobinson.com](https://www.emilyarobinson.com/)
-]
 ???
@@
 .large.emph[
-+ Introduction to Rural Shrink Smart Project with Iowa State University
++ Introduction to Rural Shrink Smart Project with Iowa State University and Open-Source Data Challenges
 
-+ Why Dashboards?
++ Motivation - Why Dashboards?
 
-+ Dashboard Design and Guided Discovery Learning (GDL)
++ Guiding Design Principles & Dashboard Design Workflow
 
-+ Parallel Coordinate Plots (PCPs) and the naive user
++ Dashboard Philosophy - PCPs and the Novice User
+  + Using Guided Discovery Learning (GDL)
 
-+ Data Collection and Feedback the loop
++ Dashboard Design Workflow 
+  + Feedback Loop (Active and Passive Methodologies)
 
@@
 I will conclude with future work.
+
+---
+class:inverse
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+.center[
+# Project Introduction and Open-source Data Description
+]
 
@@
     + Can we focus on estimating only things that objective measurements don't capture?
+---
+class:inverse
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
 
+.center[
+# Motivation - Why Use Dashboards?
+]
 ---
@@
 # Identifying Key Metrics
+<b><u>Audience</u></b>
 
-<b>Audience</b>
-
-.small.emph[
+.medium.emph[
 + As small rural towns, traditionally, do not have access:
@@
 --
+<b><u>Visualization Goals</u></b>
 
-<b>Visualization Goals</b>
-
-.small.emph[
+.medium.emph[
 + Create a central place for the excess of data to be conveyed in a single comprehensive visual.
@@
 --
+<b><u>Engage People</u></b>
 
-<b>Engage People</b>
-
-.small.emph[
+.medium.emph[
 + We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
@@
     + How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.
+---
+class:inverse
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
 
+.center[
+# Guiding Design Principles and Dashboard Design Workflow
+]    
 ---
@@
 
-[Smart Connected Communities Dashbaord](https://shiny.srvanderplas.com/Smart-Connected-Communities/)
+[Smart Connected Communities Dashboard](https://shiny.srvanderplas.com/Smart-Connected-Communities/)
 ---
+class:inverse
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+.center[
+# Dashboard Philosophy - PCPs and the Novice User
+]  
+---
+exclude: true
+
 # Dashboard Philosophy
@@
 --
+exclude: true
 <br>
@@
 ]
+
 ---
@@
 
-**Who?** Sukwon Lee, et al. wrote How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking
+**Who?** Sukwon Lee, et al. wrote "How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking"
 
@@
 We will use this model to ask questions of our town leaders to gain feedback on what they understand.
+
 ---
+
 # Dashboard Philosophy
@@
 class:primary
+
 # Dashboard Philosophy
@@
 
-**Who?** [Michael A. DeDonno](https://doi.org/10.1016/j.lindif.2016.05.023) wrote The influence of IQ on pure discovery and guided discovery learning of a complex real-world task.
+**Who?** [Michael A. DeDonno](https://doi.org/10.1016/j.lindif.2016.05.023) wrote "The Influence of IQ on pure discovery and guided discovery learning of a complex real-world task."
 
@@
 ---
+class:inverse
+<br>
+<br>
+<br>
+<br>
+<br>
+<br>
+.center[
+# Dashboard Design Workflow - Feedback Loop 
+## Active & Passive Data Collection
+]
+
+---
 # Dashboard Design WorkFlow (Feedback)
@@
 ## User Testing in Beta
-
-**Audience**
 
@@
 
-## Dashbaord Feedback
+## Dashboard Feedback
 
@@
 .center[
-<img src="images/feedback_question.png" width="44%"/>
+<img src="images/feedback_question.png" width="55%"/>
 ]
+
 ---
@@
 
+???
+With six rural town leaders, we will have enough information needed to understand the usefulness of the dashboard from
+
 ---
@@
 # Thank you! 
-### Acknowledgments: Dr. Susan VanderPlas and the ISU - Rural Shrink Smart Project Team
+### Acknowledgments: Dr. Susan VanderPlas & the ISU - Rural Shrink Smart Project Team
 <br>
@@
 
+<br>
+
+.medium.emph[
+Special Thank You to Emily A. Robinson for creating our UNL theme, find her work at `r fa("github", fill = "black")` **@earobinson95**
+]
+
 ???
 
-Thank you for the support of my adviser Dr. Susan VanderPlas and the ISU Rural Shrink Smart
+Thank you for the support of my adviser Dr. Susan VanderPlas and the ISU Rural Shrink Smart Project Team for the hard work that they have done for not only me but also the rural town leaders.

### 2022-06-14 Abstract Files/SCCAbstract_SDSS2022-knitr.Rnw
updates from conference talk

--- 2022-03-16 SCCAbstract_SDSS2022-knitr.Rnw
@@
 \section{Acknowledgements}
-This material is based upon work supported by the National Science Foundation under Grant No.
-1952007.
-We would also like to thank our collaborators at Iowa State, the Iowa League of Cities, and the various Iowa government organizations which publish their data at \url{data.iowa.gov}.
 

### 2022-06-14 SDSS Presentation/SDSS Presentation.Rmd
updates from conference talk

@@
 .center[
-<img src="images/Shrink_Smart_logo.png" width="35%"/>
+<img src="images/Shrink_Smart_Logo.png" width="35%"/>
 ]
@@
 .center[
-<img src="images/version1.png" width="80%"/>
+<img src="images/Version1.png" width="80%"/>
 ]
@@
 .center[
-<img src="images/version2.png" width="52%"/>
+<img src="images/Version2.png" width="52%"/>
 ]
@@
 .center[
-<img src="images/version3.png" width="60%"/>
+<img src="images/Version3.png" width="60%"/>
 ]
@@
 .center[
-<img src="images/version4.png" width="75%"/>
+<img src="images/Version4.png" width="75%"/>
 ]

### 2022-07-19 SDSS Journal Paper/Paper Draft.Rmd
First Draft Updates

---
title: Exploring Rural Shrink Smart Through Guided Discovery Dashboards
short-title:  
author:
  - first: Denise                 # first name
    first-init: D.              # initial of first name
    last: Bradford                  # last name
    marker: '1'               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: denise.bradford@huskers.unl.edu    # optional email address
  - first: Susan
    first-init: S.
    last: VanderPlas
    marker: '1'
    footnote-id: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: University of Nebraska - Lincoln
    prefix: Department of Statistics        # optional
    country: United States of America
    marker: 1
classoption:
  - letterpaper    # or a4paper, etc.
  - inpress        # or review
# - linenumber     # add line number for submission
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{lipsum}
abstract: >
  Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
keywords:
# alphabetical; excluding anything in the title already; lower case
  - Interactive Dashboards
  - Exploratory Data Analysis (EDA)
  - Parallel Coordinate Plots (PCP)
  - Guided Discovery Learning (GDL)
bibliography: jdsart-sample.bib
output: jds.rmd::pdf_article
---

# Abstract

Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.

# Introduction

With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \texttt{fisher}. 
A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making. An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.

Some communities continue to thrive as they lose population because they adapt, maintaining quality of life and community services for residents while investing in the future. 
This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades. As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \texttt{scc}. We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life. In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.

\section{Data Description}

Data collected from \url{data.iowa.gov} were used to create the SCC dashboard. Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution. \url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education. These data are used for details are a town level, that will provide information about the quality of life.

Data collected from the Index of Relative Rurality (IRR) \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns. The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:
\begin{enumerate}
\item Identifying the dimensions of rurality: population size, density, remoteness, and built-up area.
\item Selecting measureable variables to adequately represent each dimension:
    \begin{itemize}
        \item Size: logarithm of population size
        \item Density: logarithm of population density.
        \item Remoteness: network distance.
        \item Built-up area: urban area (as defined by the US Census Bureau) as a percentage of total land area.
    \end{itemize}
\item Re-scaling the variables onto bounded scales that range from 0 to 1.
\item Selecting a link function: unweighted average of the four re-scaled variable.
\end{enumerate}

Data collected from Rural Urban Commuting Area Codes \texttt{usda} were used to help identify towns with commuting behaviors in our rural areas. The rural-urban commuting area (RUCA) codes classify U.S. census tracts using measures of population density, urbanization, and daily commuting. This data is on a zipcode-level that will help identify those communities that commute to more urban areas. The most recent RUCA codes are based on data from the 2010 decennial census and the 2006-10 American Community Survey. The classification contains two levels. Whole numbers (1-10) delineate metropolitan, micropolitan, small town, and rural commuting areas based on the size and direction of the primary (largest) commuting flows.

Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location. In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.

One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services. This makes both visualization and statistical analysis more complicated.

# Methods

\subsection{Identifying Key Metrics for our Dashboard Design}

\begin{center}
\includegraphics[width=.5\textwidth]{Key_Metrics}
\end{center}

Our audience being small rural towns, traditionally, do not have access:
\begin{itemize}
     \item Data scientists 
      \item Ability to easily leverage data collected locally to support decisions 
\end{itemize}
Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations.

While data is important for data-driven decisions, we need to help minimize the amount of cognitive load for the town-leaders to digest the amount of data will be conveyed, as a result we will create a central place for the excess of data to be conveyed in a single comprehensive visual. We will utilize Parallel Coordinate Plot (PCPs). Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g. Fire Dept, Schools, etc.). Lastly, we will add an overview of town general basics related to the seven QoL factors.

We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.

\subsection{Guiding Dashboard Design Principles}

An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable. That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard. With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.

Thus, our initial set of dashboard design principles is as follows: 
\begin{itemize}
\item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
\item Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
\item Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
\item Allow for more detailed data and feature requests to improve the dashboard design over time.
\end{itemize}

\subsection{Dashboard Design Workflow}

We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest. The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.

Secondly in our workflow, we will look into using Unsupervised Clustering Methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.Our research implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the next five to ten similar towns, based on distances and services available. We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard. We will take into account that towns will make sure to help towns not make comparisons to towns that may be a bit out of reach and not similar in make-up. We will discuss this more in detail in the discussion, our original goal at using the clustering methods to determine similar towns actually led us to using the methods to determine the data quality of our open-source data.

In our final stages of the Dashboard Design Workflow, we will look at two important aspects including the Layout and feedback. Firstly, we will look at the dashboard layout that will allow the user to engage that most without overwhelming their curiosity and understanding of the delivered information. We will conclude with feedback on what elements on the dashboard are useful, confusing, misleading or incorrect based on two modes - clicks on the UI (Passive) and UI survey (Active).

An aspect of our dashboard design, we will iterate over designs, allowing for the design of dashboards to be a spectrum from dream design of the design team to a more functional design from the town leaders feedback.

\begin{center}
\includegraphics[width=.3\textwidth]{Version1}
\includegraphics[width=.3\textwidth]{Version2}
\includegraphics[width=.3\textwidth]{Version3}
\includegraphics[width=.3\textwidth]{Version4}
\end{center}

\subsubsection{Dashboard Components}
The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture. We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest. This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.

While our dashboard layout is vitally important into making sure that the town leaders are using the dashboards, we should discuss the dashboard components that we used along with the philosophy of those key components. Starting with our parallel coordinate plot, which will allow for objective measures related to the QoL Survey. As our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others. The framework mentioned in \texttt{lee},``How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking,'' allowed our work to determine that a novice user seeing a PCP will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding). The PCP in this study were found to be promising for novice users, when comparing to the Treemap and Chord Diagram.

Furthermore, the large set of publicly available data we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.

\subsection{Guided Discovery Learning (GDL)}
We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \texttt{dedonno}.

The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals. Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size. The graphical representations including a parallel coordinate plot and geographic maps [Figure 1]. The Rural Smart EDA Dashboard map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.

The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of data, with constraints, to similar towns. The Rural Smart EDA Dashboard features offer geospatial and structural data to assist town leaders in gaining a comprehensive understanding of the town in which they support.

Each component provides the town leaders with analytical aspects related to the performance of their town. As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.

\subsection{User Testing}
In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices. We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.

Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.

Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives," \texttt{young}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions. We will adapt these questions to better help the town leaders communicate clear feedback for our team to use. We will collect how the users are experiencing the app with these set questions:
\begin{itemize}
\item Is the information the dashboard presents clear? \textit{(Scope)}
\item Is the dashboard easy to navigate? \textit{(Usability)}
\item Is the dashboard used in a socially responsible manner? \textit{(Ethics)}
\item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)}
\item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)}
\end{itemize}

\subsubsection{Beta Testing}
Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups:
\begin{itemize}
    \item Graphics Group Members (Users know Statistical Graphics)
    \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics)
    \item  Others (Member of doctoral community that has background in educational adaptivity)
\end{itemize}

While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes.

\section{Discussion and Future Research}

\subsection{Data Collection Revisited}

As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources. This work empowered our time to look for other open data sources to validate our vital statistics and maps. Since the clustering methods were used differently, we decided to use the USDA's version of rural area definitions. This allowed for a more transparent look into towns that may be similar based on there rurality index. Similar to the Census, the rurality metrics are county based and updated every ten years. This could later cause a problem for towns that may be shrinking faster than the collection of data. It could be worth looking into developing a proxy from our data collected that would be reliable for the years in between the USDA's collected years.

In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns. This may be important to identify the reasons for shrinking. In the event, that the towns are aging, will we be able to identify how fast the younger generation is moving out of the towns. This could be the most important piece of truly identifying the use of data in towns with an aging population.

\subsection{Visualization Revisited}
In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful foor the sake of ease on the town leaders.

In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

\subsection{Feedback}
We will continue to engage with twon leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.

### 2022-07-19 SDSS Journal Paper/Paper-Draft.tex
First Draft Updates

--- 2022-07-19 SDSS Journal Paper/Paper Draft.Rmd
@@
----
-title: Exploring Rural Shrink Smart Through Guided Discovery Dashboards
-short-title:  
-author:
-  - first: Denise                 # first name
-    first-init: D.              # initial of first name
-    last: Bradford                  # last name
-    marker: '1'               # identifier for the affiliation
-    footnote-id: 1              # optional identifier for the footnote
-    email: denise.bradford@huskers.unl.edu    # optional email address
-  - first: Susan
-    first-init: S.
-    last: VanderPlas
-    marker: '1'
-    footnote-id: 1
-footnote:
-  - content: Corresponding author
-    type: corresp               # set type for the corresponding author
-    id: 1
-affiliation:
-  - institution: University of Nebraska - Lincoln
-    prefix: Department of Statistics        # optional
-    country: United States of America
-    marker: 1
-classoption:
-  - letterpaper    # or a4paper, etc.
-  - inpress        # or review
-# - linenumber     # add line number for submission
-preamble: >
-  \usepackage{amsfonts,amsmath,amssymb,amsthm}
-  \usepackage{booktabs}
-  \usepackage{lipsum}
-abstract: >
+\documentclass[letterpaper,inpress]{jdsart}
+
+\setcounter{page}{1}
+\pubmonth{July}
+\pubyear{2020}
+\volume{xx}
+\issue{xx}
+\doi{0000}
+
+\usepackage[utf8]{inputenc}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage{amsfonts,amsmath,amssymb,amsthm} \usepackage{booktabs} \usepackage{lipsum}
+
+\begin{document}
+\begin{frontmatter}
+
+\title{Exploring Rural Shrink Smart Through Guided Discovery Dashboards}
+
+\author[1]{
+  \inits{D.}
+  \fnms{Denise}
+  \snm{Bradford}  \thanksref{1}  \ead{denise.bradford@huskers.unl.edu}}
+\author[1]{
+  \inits{S.}
+  \fnms{Susan}
+  \snm{VanderPlas}  \thanksref{1}}
+
+\thankstext[type=corresp,id=1]{Corresponding author}
+\address[1]{Department of Statistics, 
+  \institution{University of Nebraska - Lincoln}, \cny{United States of America}}
+
+\begin{abstract}
 Many small and rural places are shrinking.
@@
 Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
-keywords:
-# alphabetical; excluding anything in the title already; lower case
-  - Interactive Dashboards
-  - Exploratory Data Analysis (EDA)
-  - Parallel Coordinate Plots (PCP)
-  - Guided Discovery Learning (GDL)
-bibliography: jdsart-sample.bib
-output: jds.rmd::pdf_article
----
+\end{abstract}
 
-# Abstract
+\begin{keywords}
+\kwd{Interactive Dashboards}\kwd{Exploratory Data Analysis (EDA)}\kwd{Parallel Coordinate Plots (PCP)}\kwd{Guided Discovery Learning (GDL)}.
+\end{keywords}
+
+\end{frontmatter}
+
+\hypertarget{abstract}{%
+\section{Abstract}\label{abstract}}
 
@@
 
-# Introduction
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 Four steps are involved in its design:
+
 \begin{enumerate}
@@
 One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
-Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
+Thus, in addition to the usual challenges of working with real-world data that is ``messy'' in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
 This makes both visualization and statistical analysis more complicated.
 
-# Methods
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 Our audience being small rural towns, traditionally, do not have access:
+
 \begin{itemize}
@@
 \end{itemize}
+
 Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations.
@@
 We will utilize Parallel Coordinate Plot (PCPs).
-Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g.
-Fire Dept, Schools, etc.).
+Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g.~Fire Dept, Schools, etc.).
 Lastly, we will add an overview of town general basics related to the seven QoL factors.
@@
 Thus, our initial set of dashboard design principles is as follows:
+
 \begin{itemize}
@@
 \subsubsection{Dashboard Components}
+
 The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
@@
 We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others.
-The framework mentioned in \texttt{lee},``How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking,'' allowed our work to determine that a novice user seeing a PCP will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
+The framework mentioned in \texttt{lee},``How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice's Information Visualization Sensemaking,'\,' allowed our work to determine that a novice user seeing a PCP will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
 The PCP in this study were found to be promising for novice users, when comparing to the Treemap and Chord Diagram.
@@
 \subsection{Guided Discovery Learning (GDL)}
+
 We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
@@
 Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
-The graphical representations including a parallel coordinate plot and geographic maps [Figure 1].
+The graphical representations including a parallel coordinate plot and geographic maps {[}Figure 1{]}.
 The Rural Smart EDA Dashboard map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
@@
 \subsection{User Testing}
+
 In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
-We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.
+We examine which aspects of dashboard design ``common knowledge'\,' hold up in practice, and assess why these principles may not hold in our use-case.
 
 Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
+As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.
 
-Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives," \texttt{young}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
+Using the framework from ``Creating design guidelines for building city dashboards from a user's perspectives,'' \texttt{young}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
 We will adapt these questions to better help the town leaders communicate clear feedback for our team to use.
 We will collect how the users are experiencing the app with these set questions:
+
 \begin{itemize}
@@
 \subsubsection{Beta Testing}
+
 Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups:
+
 \begin{itemize}
@@
 \subsection{Visualization Revisited}
+
 In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
@@
 \subsection{Feedback}
+
 We will continue to engage with twon leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
+
+\bibliography{jdsart-sample.bib}
+\bibliographystyle{jds}
+
+\end{document}

### 2022-07-20 SDSS Journal Paper/Paper Draft.Rmd
Tweaks to the intro, adding additional colors for editing purposes.

@@
   \usepackage{lipsum}
+  \newcommand{\db}[1]{{\textcolor{blue}{#1}}}
+  \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
+
 abstract: >
@@
 bibliography: jdsart-sample.bib
-output: jds.rmd::pdf_article
+output: 
+  jds.rmd::pdf_article:
+    keep_tex: true
 ---
@@
 # Introduction
-
-With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \texttt{fisher}. 
-A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.
-An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.
+\svp{As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
+As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.}
+<!-- With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{fisher}.  -->
+\svp{In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.}
+<!-- A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.  -->
+\svp{Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity.
+One glance at a car's dashboard will tell you the speed, RPM, engine temperature, amount of gas in the tank; more importantly, however, the goal is not for the user to remember all of these characteristics, but to assess whether any of these quantities is outside of the expected range.
+Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.}
+<!-- An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.  -->
+\svp{In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.}
 
@@
 This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades.
-As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \texttt{scc}.
+As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \cite{scc}.
 We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.

### 2022-07-20 SDSS Journal Paper/Paper Draft.Rmd
comments/tweaks

@@
 This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades. 
-As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \cite{scc}.
+As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}.
 We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life. 
 In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+\svp{This paragraph is pretty good, but see if you can tweak it to blend in a bit better with the edits I made above.}
 
@@
 Data collected from \url{data.iowa.gov} were used to create the SCC dashboard.
-Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.
-\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+Most of these data sets are collected at a town/city or county \svp{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}.
+\svp{In addition to the more commonly available statistics derived from e.g.
+the census and American Community Survey,} \url{data.iowa.gov} contains \svp{several} unique \svp{data sets}, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
 
+\svp{Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.}
 Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
 The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
-These data are used for details are a town level, that will provide information about the quality of life.
+\svp{Cite Dave's papers about the ISTP QOL metrics, including education}
+<!-- These data are used for details are a town level, that will provide information about the quality of life. -->
 
-Data collected from the Index of Relative Rurality (IRR) \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns.
+Data collected from the Index of Relative Rurality (IRR) \svp{Create a bibtex entry for this and cite it properly, please.
+Same goes for most of the other URLs.} \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns.
 The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality.
@@
 
-Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
-In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.
-
 One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
@@
 This makes both visualization and statistical analysis more complicated.
+
+<!-- Not sure this paragraph is in the right place and/or necessary -->
+<!-- Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
+In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.
+-->
 

### 2022-07-20 SDSS Journal Paper/Paper Draft.Rmd
Key metrics

@@
 census tracts using measures of population density, urbanization, and daily commuting.
-This data is on a zipcode-level that will help identify those communities that commute to more urban areas.
+This data is on a zip code-level that will help identify those communities that commute to more urban areas.
 The most recent RUCA codes are based on data from the 2010 decennial census and the 2006-10 American Community Survey.
@@
 
-One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
+One of the interesting features of this assembled data set is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
 Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
@@
 
-Our audience being small rural towns, traditionally, do not have access:
-\begin{itemize}
-     \item Data scientists 
-      \item Ability to easily leverage data collected locally to support decisions 
-\end{itemize}
-Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations.
+\svp{One problem we identified early in the process of assessing smart-shrinkage strategies in small towns is that these towns do not have the resources to make data-driven decisions.
+Typically, small towns in Iowa are managed by at most a few part-time employees or volunteers.
+In some cases, essential management functions of the town are paid, but the municipalities we are interested in do not have sufficient funding to hire professionals to gather and analyze data.}
 
-While data is important for data-driven decisions, we need to help minimize the amount of cognitive load for the town-leaders to digest the amount of data will be conveyed, as a result we will create a central place for the excess of data to be conveyed in a single comprehensive visual.
+\svp{As part of a wider project investigating the strategies towns use to maintain quality of life amid shrinking population, our research team provides communities with data about their own town, but also comparable towns across the state which may have a different approach to city services.
+In combination with other engagement strategies that are more qualitative, we hope to use this interactive dashboard approach to assist small Iowa cities with generalizing and developing strategies to improve or maintain quality of life amid shrinking populations.}
+
+\svp{One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data.
+At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.}
+\svp{Our initial strategy to accomplish this task is to utilize parallel coordinate plots (PCPs), a multivariate data visualization method that can accommodate both categorical and numerical data on in the same visualization.
+Supplementing the PCP that is the primary data display, we also provide some contextual information in the form of maps and "quick facts" that allow viewers to quickly assess a new town before comparing to a wider selection of towns.}
+<!-- While data is important for data-driven decisions, we need to help minimize the amount of cognitive load for the town-leaders to digest the amount of data will be conveyed, as a result we will create a central place for the excess of data to be conveyed in a single comprehensive visual.
 We will utilize Parallel Coordinate Plot (PCPs).
@@
 Lastly, we will add an overview of town general basics related to the seven QoL factors.
+-->
 

### 2022-07-28 SDSS Journal Paper/Paper Draft.Rmd
Additional edits

@@
 
-An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
+An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and \svp{hard to apply to more generalized settings}.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
-With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
+With this in mind, we \svp{set out to clearly document our} intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
 
@@
 \item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
-\item Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
-\item Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
-\item Allow for more detailed data and feature requests to improve the dashboard design over time.
+\item \svp{We should facilitate} comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+\item \svp{We will present} the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+\item \svp{We will implement feedback mechanisms that allow us to provide} more detailed data and \svp{respond to} feature requests to improve the dashboard design over time.
 \end{itemize}
 
-\subsection{Dashboard Design Workflow}
+\subsection{Dashboard Design Process Overview}
 
-We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+We provided users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
 The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.
+\svp{The initial design sketch and implementation are shown in} \autoref{fig:v1}.
 
-Secondly in our workflow, we will look into using Unsupervised Clustering Methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.Our research implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the next five to ten similar towns, based on distances and services available.
+We \svp{implemented some suggested town comparisons using} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
+\svp{We initially focused on} determining the next five to ten similar towns, based on distances to services.
 We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
-We will take into account that towns will make sure to help towns not make comparisons to towns that may be a bit out of reach and not similar in make-up.
-We will discuss this more in detail in the discussion, our original goal at using the clustering methods to determine similar towns actually led us to using the methods to determine the data quality of our open-source data.
+\svp{This feature became an important diagnostic for our data quality, but was not particularly useful to the application users, as they already know what towns they want to compare to; our suggested comparisons were just in the way.}
 
-In our final stages of the Dashboard Design Workflow, we will look at two important aspects including the Layout and feedback.
+<!-- \svp{The discussion section will include an assessment of the initial layout design based on both direct user feedback through the survey form and based on UI clicks that provide a passive feedback mechanism to indicate what elements of the UI the users tend to interact with.
+We will also discuss verbal feedback we received when testing the application.} -->
+
+<!-- In our final stages of the Dashboard Design Workflow, we will look at two important aspects including the Layout and feedback.
 Firstly, we will look at the dashboard layout that will allow the user to engage that most without overwhelming their curiosity and understanding of the delivered information.
 We will conclude with feedback on what elements on the dashboard are useful, confusing, misleading or incorrect based on two modes - clicks on the UI (Passive) and UI survey (Active).
+-->
 
-An aspect of our dashboard design, we will iterate over designs, allowing for the design of dashboards to be a spectrum from dream design of the design team to a more functional design from the town leaders feedback.
+\svp{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
+However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}
+<!-- An aspect of our dashboard design, we will iterate over designs, allowing for the design of dashboards to be a spectrum from dream design of the design team to a more functional design from the town leaders feedback.
+-->
 
-\begin{center}
-\includegraphics[width=.3\textwidth]{Version1}
-\includegraphics[width=.3\textwidth]{Version2}
-\includegraphics[width=.3\textwidth]{Version3}
-\includegraphics[width=.3\textwidth]{Version4}
-\end{center}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Version1}
+
+\includegraphics[width=.7\textwidth]{Version2}
+\caption{Initial dashboard design sketch (top) and implementation (bottom).}\label{fig:v1}
+\end{figure}
 
@@
 
-While our dashboard layout is vitally important into making sure that the town leaders are using the dashboards, we should discuss the dashboard components that we used along with the philosophy of those key components.
-Starting with our parallel coordinate plot, which will allow for objective measures related to the QoL Survey.
-As our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+\svp{This design is based on research into visualization sensemaking}[@lee].
 We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others.
-The framework mentioned in \texttt{lee},``How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking,'' allowed our work to determine that a novice user seeing a PCP will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
-The PCP in this study were found to be promising for novice users, when comparing to the Treemap and Chord Diagram.
-
-Furthermore, the large set of publicly available data we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+\svp{When faced with a parallel coordinate plot (PCP), a novice user} will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
+\svp{As PCPs are likely to be unfamiliar to most users, we decided to make the first part of the app less visually intimidating, providing maps showing the distance to local services in version 1 (\autoref{fig:v1}), and showing a condensed service map along with basic facts about the town in version 2 (\autoref{fig:v2}).}
 
@@
 We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
-This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \texttt{dedonno}.
+This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
 
 The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
-Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
-The graphical representations including a parallel coordinate plot and geographic maps [Figure 1].
+\svp{In version 2, as shown in} \autoref{fig:v2}, value boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
+The graphical representations including a parallel coordinate plot and geographic maps.
 The Rural Smart EDA Dashboard map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
+
+<!-- Stopped here -->
 
@@
 
-Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives," \texttt{young}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
+Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@young], we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
 We will adapt these questions to better help the town leaders communicate clear feedback for our team to use.
@@
 While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes.
+
+\begin{figure}
+\includegraphics[width=.8\textwidth]{Version3}
+
+\includegraphics[width=\textwidth]{Version4}
+\caption{A second iteration of the sketched design (top) and the implementation (bottom).}\label{fig:v2}
+\end{figure}
 

### 2022-07-28 SDSS Journal Paper/Paper-Draft.tex
Additional edits

@@
 
-\usepackage{amsfonts,amsmath,amssymb,amsthm} \usepackage{booktabs} \usepackage{lipsum}
+\usepackage{amsfonts,amsmath,amssymb,amsthm} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{blue}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
 
@@
 
-With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \texttt{fisher}.
-A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.
-An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.
+{\textcolor{red}{As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
+As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.}}
+
+{\textcolor{red}{In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.}}
+
+{\textcolor{red}{Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity.
+One glance at a car's dashboard will tell you the speed, RPM, engine temperature, amount of gas in the tank; more importantly, however, the goal is not for the user to remember all of these characteristics, but to assess whether any of these quantities is outside of the expected range.
+Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.}}
+
+{\textcolor{red}{In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.}}
 
@@
 This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades.
-As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \texttt{scc}.
+As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}.
 We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
 In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+{\textcolor{red}{This paragraph is pretty good, but see if you can tweak it to blend in a bit better with the edits I made above.}}
 
@@
 Data collected from \url{data.iowa.gov} were used to create the SCC dashboard.
-Most of these datasets are collected on a town/city or county level, requiring us to carefully join data accounting for differences in spatial resolution.
-\url{data.iowa.gov} contains unique information about residents, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+Most of these data sets are collected at a town/city or county {\textcolor{red}{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}}.
+{\textcolor{red}{In addition to the more commonly available statistics derived from e.g.
+the census and American Community Survey,}} \url{data.iowa.gov} contains {\textcolor{red}{several}} unique {\textcolor{red}{data sets}}, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+
+{\textcolor{red}{Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.}}
 
@@
 The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
-These data are used for details are a town level, that will provide information about the quality of life.
+{\textcolor{red}{Cite Dave's papers about the ISTP QOL metrics, including education}}
 
-Data collected from the Index of Relative Rurality (IRR) \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns.
+Data collected from the Index of Relative Rurality (IRR) {\textcolor{red}{Create a bibtex entry for this and cite it properly, please.
+Same goes for most of the other URLs.}} \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns.
 The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality.
@@
 
-Using this data, we created a dashboard which allows communities to explore these data and compare and contrast their local community to other communities of similar size and location.
-In addition to manual comparisons created by the user, we will use statistical clustering methods to identify groups of towns which employ similar strategies to maintain resident quality of life.
-
-One of the interesting features of this assembled dataset is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
+One of the interesting features of this assembled data set is that missing data can be missing for multiple reasons: not all state data is complete, but data about certain services may also be missing because towns do not offer that service.
 Thus, in addition to the usual challenges of working with real-world data that is ``messy'' in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
@@
 
-Our audience being small rural towns, traditionally, do not have access:
+{\textcolor{red}{One problem we identified early in the process of assessing smart-shrinkage strategies in small towns is that these towns do not have the resources to make data-driven decisions.
+Typically, small towns in Iowa are managed by at most a few part-time employees or volunteers.
+In some cases, essential management functions of the town are paid, but the municipalities we are interested in do not have sufficient funding to hire professionals to gather and analyze data.}}
 
-\begin{itemize}
-     \item Data scientists 
-      \item Ability to easily leverage data collected locally to support decisions 
-\end{itemize}
+{\textcolor{red}{As part of a wider project investigating the strategies towns use to maintain quality of life amid shrinking population, our research team provides communities with data about their own town, but also comparable towns across the state which may have a different approach to city services.
+In combination with other engagement strategies that are more qualitative, we hope to use this interactive dashboard approach to assist small Iowa cities with generalizing and developing strategies to improve or maintain quality of life amid shrinking populations.}}
 
-Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations.
-
-While data is important for data-driven decisions, we need to help minimize the amount of cognitive load for the town-leaders to digest the amount of data will be conveyed, as a result we will create a central place for the excess of data to be conveyed in a single comprehensive visual.
-We will utilize Parallel Coordinate Plot (PCPs).
-Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g.~Fire Dept, Schools, etc.).
-Lastly, we will add an overview of town general basics related to the seven QoL factors.
+{\textcolor{red}{One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data.
+At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.}}
+{\textcolor{red}{Our initial strategy to accomplish this task is to utilize parallel coordinate plots (PCPs), a multivariate data visualization method that can accommodate both categorical and numerical data on in the same visualization.
+Supplementing the PCP that is the primary data display, we also provide some contextual information in the form of maps and "quick facts" that allow viewers to quickly assess a new town before comparing to a wider selection of towns.}}
 
@@
 
-An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and not generalizable.
+An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and {\textcolor{red}{hard to apply to more generalized settings}}.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
-With this in mind, we have clearly documented our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
+With this in mind, we {\textcolor{red}{set out to clearly document our}} intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
 
@@
 \item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
-\item Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
-\item Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
-\item Allow for more detailed data and feature requests to improve the dashboard design over time.
+\item {\textcolor{red}{We should facilitate}} comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+\item {\textcolor{red}{We will present}} the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+\item {\textcolor{red}{We will implement feedback mechanisms that allow us to provide}} more detailed data and {\textcolor{red}{respond to}} feature requests to improve the dashboard design over time.
 \end{itemize}
 
-\subsection{Dashboard Design Workflow}
+\subsection{Dashboard Design Process Overview}
 
-We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+We provided users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
 The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.
+{\textcolor{red}{The initial design sketch and implementation are shown in}} \autoref{fig:v1}.
 
-Secondly in our workflow, we will look into using Unsupervised Clustering Methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.Our research implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the next five to ten similar towns, based on distances and services available.
+We {\textcolor{red}{implemented some suggested town comparisons using}} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
+{\textcolor{red}{We initially focused on}} determining the next five to ten similar towns, based on distances to services.
 We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
-We will take into account that towns will make sure to help towns not make comparisons to towns that may be a bit out of reach and not similar in make-up.
-We will discuss this more in detail in the discussion, our original goal at using the clustering methods to determine similar towns actually led us to using the methods to determine the data quality of our open-source data.
+{\textcolor{red}{This feature became an important diagnostic for our data quality, but was not particularly useful to the application users, as they already know what towns they want to compare to; our suggested comparisons were just in the way.}}
 
-In our final stages of the Dashboard Design Workflow, we will look at two important aspects including the Layout and feedback.
-Firstly, we will look at the dashboard layout that will allow the user to engage that most without overwhelming their curiosity and understanding of the delivered information.
-We will conclude with feedback on what elements on the dashboard are useful, confusing, misleading or incorrect based on two modes - clicks on the UI (Passive) and UI survey (Active).
+{\textcolor{red}{We will conclude with an assessment of the initial layout design based on both direct user feedback through the survey form and based on UI clicks that provide a passive feedback mechanism to indicate what elements of the UI the users tend to interact with.
+We will also discuss verbal feedback we received when testing the application.}}
 
-An aspect of our dashboard design, we will iterate over designs, allowing for the design of dashboards to be a spectrum from dream design of the design team to a more functional design from the town leaders feedback.
+{\textcolor{red}{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
+However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}}
 
-\begin{center}
-\includegraphics[width=.3\textwidth]{Version1}
-\includegraphics[width=.3\textwidth]{Version2}
-\includegraphics[width=.3\textwidth]{Version3}
-\includegraphics[width=.3\textwidth]{Version4}
-\end{center}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Version1}
+
+\includegraphics[width=.7\textwidth]{Version2}
+\caption{Initial dashboard design sketch (top) and implementation (bottom).}\label{fig:v1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.8\textwidth]{Version3}
+
+\includegraphics[width=\textwidth]{Version4}
+\caption{A second iteration of the sketched design (top) and the implementation (bottom).}\label{fig:v2}
+\end{figure}
 

### 2022-07-29 SDSS Journal Paper/Paper Draft.Rmd
Fix bib formatting - consistency

@@
 
-Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@young], we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
+Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@lee], we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
 We will adapt these questions to better help the town leaders communicate clear feedback for our team to use.

### 2022-07-29 SDSS Journal Paper/Paper Draft.Rmd
Updates and restructuring

@@
 
-Data collected from \url{data.iowa.gov} were used to create the SCC dashboard.
-Most of these data sets are collected at a town/city or county \svp{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}.
+\svp{The Smart and Connected Community (SCC) dashaboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.} Most of these data sets are collected at a town/city or county \svp{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}.
 \svp{In addition to the more commonly available statistics derived from e.g.
@@
 Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
-This makes both visualization and statistical analysis more complicated.
+This makes both visualization and statistical analysis more complicated \svp{(and more interesting)}.
 
@@
 
-# Methods
+# Dashboard Design Considerations
 
-\subsection{Identifying Key Metrics for our Dashboard Design}
-
-\begin{center}
+\begin{figure}
 \includegraphics[width=.5\textwidth]{Key_Metrics}
-\end{center}
+\caption{Diagram of considerations for our dashboard design process.}\label{fig:metrics}
+\end{figure}
 
@@
 At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.}
-\svp{Our initial strategy to accomplish this task is to utilize parallel coordinate plots (PCPs), a multivariate data visualization method that can accommodate both categorical and numerical data on in the same visualization.
-Supplementing the PCP that is the primary data display, we also provide some contextual information in the form of maps and "quick facts" that allow viewers to quickly assess a new town before comparing to a wider selection of towns.}
-<!-- While data is important for data-driven decisions, we need to help minimize the amount of cognitive load for the town-leaders to digest the amount of data will be conveyed, as a result we will create a central place for the excess of data to be conveyed in a single comprehensive visual.
-We will utilize Parallel Coordinate Plot (PCPs).
-Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g.
-Fire Dept, Schools, etc.).
-Lastly, we will add an overview of town general basics related to the seven QoL factors.
--->
+\svp{This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a} framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
 
-We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+\svp{We expect that users will be interested in "sets" of variables from the wider dataset, which we assembled based on quality of life factors in the Iowa Small Town Poll [@petersCommunityResiliencyDeclining2019].
+For instance, users might be interested in medical and social services available to residents, such as a local primary care clinic, nursing homes which are within driving distance, and the distance to the nearest emergency room; these factors might be explored separately from variables describing the services provided directly by city government, such as parks and recreation expenditures, snow removal services, and the distance to the closest fire station.}
 
-\subsection{Guiding Dashboard Design Principles}
+\svp{As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have.
+Introduced in the 1880s [@dOcagne:1885], parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation.
+When representing categorical data, parallel set plots may show "blocks" of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables [@Bendix:2005]; modifications of this design, such as common-angle plots [@Hofmann:2013], address the issues which arise due to line-width illusions [@sine, @sineillusion].
+Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp].
+In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
+}
 
-An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and \svp{hard to apply to more generalized settings}.
+\svp{We also anticipate that users will be interested in comparing their town to other, similar towns.
+We will discuss the different ways that this comparison strategy was implemented in each dashboard in the next section, which describes the evolution of the dashboard over time and accounting for feedback from users and other researchers on the wider project.}
+
+\svp{One final component of this project is that our dashboard is part of a wider effort to work with towns to understand the different strategies used to maintain resident quality of life amid shrinking populations.
+Thus, while the town leaders are our primary audience, we also are creating this applet for use in parallel with a team of other researchers: sociologists, economists, city planning specialists, and artists.
+These researchers opinions and feedback about the dashboard are also useful and important, as they regularly work with town leaders in different capacities and have an understanding of what factors are most important to them and what types of questions these leaders may have when faced with data and unfamiliar statistical visualizations.}
+
+\svp{Throughout the design process,} we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+
+## Guiding Design Principles
+
+Research on dashboard creation and interactive visualization tends to be very task-specific and \svp{hard to apply to more generalized settings}.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
@@
 
-\subsection{Dashboard Design Process Overview}
+\svp{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
+However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}
 
-We provided users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
-The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.
+# Dashboard Design Process
+
+## Dashboard Components
+
+\svp{In this section, we discuss the philosophy behind the basic "building blocks" of the dashboard.
+This philosophy is present in all of the iterations of the dashboard that we present in this discussion, and we will evaluate the overall philosophy's effectiveness in the conclusion.}
+
+The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.
+
+Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+\svp{As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest.} \svp{This design is based on research into visualization sensemaking}[@lee], \svp{in that we allow users to explore outward from the familiar to the unknown}.
+The map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
+\svp{XXX cite R, use consistent name for the app, cite OSRM and expand the acronym, and exaplain that in some communities, "as the crow flies" distance isn't a good metric... road distance is. XXX}
+<!-- \svp{As PCPs are likely to be unfamiliar to most users, we decided to make the first part of the app less visually intimidating, providing maps showing the distance to local services in version 1 (\autoref{fig:v1}), and showing a condensed service map along with basic facts about the town in version 2 (\autoref{fig:v2}).} -->
+
+<!-- We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others.  -->
+\svp{When faced with the next component, a parallel coordinate plot (PCP), a novice user} will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
+
+Taken together, the app is a single page; the initial "solid ground" which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
+\svp{In version 2, as shown in} \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.} \svp{This relatively straightforward display is} followed by a parallel coordinate plot that allows the \svp{user} to see \svp{similar} towns along dimensions such as economic indicators or population size.
+
+## Initial Draft
+
 \svp{The initial design sketch and implementation are shown in} \autoref{fig:v1}.
 
-We \svp{implemented some suggested town comparisons using} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
-\svp{We initially focused on} determining the next five to ten similar towns, based on distances to services.
-We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
-\svp{This feature became an important diagnostic for our data quality, but was not particularly useful to the application users, as they already know what towns they want to compare to; our suggested comparisons were just in the way.}
-
-<!-- \svp{The discussion section will include an assessment of the initial layout design based on both direct user feedback through the survey form and based on UI clicks that provide a passive feedback mechanism to indicate what elements of the UI the users tend to interact with.
-We will also discuss verbal feedback we received when testing the application.} -->
-
-<!-- In our final stages of the Dashboard Design Workflow, we will look at two important aspects including the Layout and feedback.
-Firstly, we will look at the dashboard layout that will allow the user to engage that most without overwhelming their curiosity and understanding of the delivered information.
-We will conclude with feedback on what elements on the dashboard are useful, confusing, misleading or incorrect based on two modes - clicks on the UI (Passive) and UI survey (Active).
--->
-
-\svp{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
-However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}
-<!-- An aspect of our dashboard design, we will iterate over designs, allowing for the design of dashboards to be a spectrum from dream design of the design team to a more functional design from the town leaders feedback.
+\svp{Users' towns} are at the center of our application, and comparisons to other, similar towns are secondary.
+As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
+<!-- The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.
 -->
@@
 
-\subsubsection{Dashboard Components}
-The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
-We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.
+We \svp{implemented some suggested town comparisons using} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
+\svp{We initially focused on} determining the next five to ten similar towns, based on distances to services.
+\svp{This feature became an important diagnostic for our data quality, as it became clear that towns which were grouped with big cities but which did not have a large population were so grouped because of missing data.}\svp{Unfortunately, this clustering feature was not as useful to the application users, as they came to the dashboard with a pre-existing set of towns to compare to; our suggested comparisons were in the way.}
 
-Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-\svp{This design is based on research into visualization sensemaking}[@lee].
-We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others.
-\svp{When faced with a parallel coordinate plot (PCP), a novice user} will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
-\svp{As PCPs are likely to be unfamiliar to most users, we decided to make the first part of the app less visually intimidating, providing maps showing the distance to local services in version 1 (\autoref{fig:v1}), and showing a condensed service map along with basic facts about the town in version 2 (\autoref{fig:v2}).}
+\svp{The initial dashboard design featured several responsive maps showing the distance to the nearest hospital, fire department, post office, and school.
+These maps were ineffective for several reasons: }
 
-\subsection{Guided Discovery Learning (GDL)}
-We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
-This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
+- Town residents already know this information (though it was useful for us as the dashboard designers, because we aren't nearly as familiar with the 900+ small towns in Iowa)
+- We computed distance from services relative to the center of town - coordinates provided in the data from \url{data.iowa.gov}.
+Generally speaking, the post office is at the center of town and the fire department is usually very close to the center of town; these two maps were useless.
+The school and hospital maps were less useless, but still did not provide particularly useful information to people already familiar with the town.
+- It became clear that it might be more useful to show the comparison towns on a map (relative to the town of interest) so that users could compare geographical ratings for unfamiliar data to familiar data.
 
-The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
-\svp{In version 2, as shown in} \autoref{fig:v2}, value boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
-The graphical representations including a parallel coordinate plot and geographic maps.
-The Rural Smart EDA Dashboard map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
+\svp{In addition, we received feedback on the parallel coordinate plot at the bottom of the app which was surprising: the viewers (in this case, other researchers on the team) were not as intimidated by the parallel coordinate plot as we had expected.
+They did need some explanation of how to read the plot, and these hints need to be included in the dashboard, but they grasped the fundamental idea of the plot very quickly.}
 
-<!-- Stopped here -->
+\svp{Our conclusion, based on this initial dashboard draft, was that we needed to restructure the application.
+Our attempt to show familiar information first to "build up" to the more unfamiliar structure of a parallel coordinate plot was not effective; there was too much clutter and not enough new information to draw users in.}
 
-The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of data, with constraints, to similar towns.
-The Rural Smart EDA Dashboard features offer geospatial and structural data to assist town leaders in gaining a comprehensive understanding of the town in which they support.
+<!-- An aspect of our dashboard design, we will iterate over designs, allowing for the design of dashboards to be a spectrum from dream design of the design team to a more functional design from the town leaders feedback.
+-->
 
-Each component provides the town leaders with analytical aspects related to the performance of their town.
-As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.
+## Redesign
+
+\begin{figure}
+\includegraphics[width=.8\textwidth]{Version3}
+
+\includegraphics[width=\textwidth]{Version4}
+\caption{A second iteration of the sketched design (top) and the implementation (bottom).}\label{fig:v2}
+\end{figure}
+
+\svp{XXX Denise, can you follow up with the initial design section and explain the 2nd design and approach?
+Why did we change things?
+How did those changes impact feedback from users?
+E.g.
+wanting to show the whole range of values in the PCP vs.
+being visually overwhelmed.
+Incorporate the user feedback into the narrative.
+XXX}
+
+# Discussion
+How did the philosophy work out?
+What worked and didn't (overall)?
+People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with?
+Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis.
+What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare.
+We also should provide much, much more user support for how to interact with and interpret the app and results.
+
+# Future Work
+
+Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
+Transition to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
+Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.
+
+# Conclusions
+Specific takeaways - visualization research (broad) vs.
+dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
+Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
+Test early, test often...
 
@@
 
-\begin{figure}
-\includegraphics[width=.8\textwidth]{Version3}
-
-\includegraphics[width=\textwidth]{Version4}
-\caption{A second iteration of the sketched design (top) and the implementation (bottom).}\label{fig:v2}
-\end{figure}
-
 \section{Discussion and Future Research}

### 2022-07-29 SDSS Journal Paper/Paper-Draft.tex
Updates and restructuring

@@
 
-Data collected from \url{data.iowa.gov} were used to create the SCC dashboard.
-Most of these data sets are collected at a town/city or county {\textcolor{red}{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}}.
+{\textcolor{red}{The Smart and Connected Community (SCC) dashaboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.}} Most of these data sets are collected at a town/city or county {\textcolor{red}{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}}.
 {\textcolor{red}{In addition to the more commonly available statistics derived from e.g.
@@
 Thus, in addition to the usual challenges of working with real-world data that is ``messy'' in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
-This makes both visualization and statistical analysis more complicated.
+This makes both visualization and statistical analysis more complicated {\textcolor{red}{(and more interesting)}}.
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\hypertarget{dashboard-design-considerations}{%
+\section{Dashboard Design Considerations}\label{dashboard-design-considerations}}
 
-\subsection{Identifying Key Metrics for our Dashboard Design}
-
-\begin{center}
+\begin{figure}
 \includegraphics[width=.5\textwidth]{Key_Metrics}
-\end{center}
+\caption{Diagram of considerations for our dashboard design process.}\label{fig:metrics}
+\end{figure}
 
@@
 At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.}}
-{\textcolor{red}{Our initial strategy to accomplish this task is to utilize parallel coordinate plots (PCPs), a multivariate data visualization method that can accommodate both categorical and numerical data on in the same visualization.
-Supplementing the PCP that is the primary data display, we also provide some contextual information in the form of maps and "quick facts" that allow viewers to quickly assess a new town before comparing to a wider selection of towns.}}
 
-We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+{\textcolor{red}{This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a}} framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
 
-\subsection{Guiding Dashboard Design Principles}
+{\textcolor{red}{We expect that users will be interested in "sets" of variables from the wider dataset, which we assembled based on quality of life factors in the Iowa Small Town Poll [@petersCommunityResiliencyDeclining2019].
+For instance, users might be interested in medical and social services available to residents, such as a local primary care clinic, nursing homes which are within driving distance, and the distance to the nearest emergency room; these factors might be explored separately from variables describing the services provided directly by city government, such as parks and recreation expenditures, snow removal services, and the distance to the closest fire station.}}
 
-An additional challenge is that research on dashboard creation and interactive visualization tends to be very task-specific and {\textcolor{red}{hard to apply to more generalized settings}}.
+{\textcolor{red}{As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have.
+Introduced in the 1880s [@dOcagne:1885], parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation.
+When representing categorical data, parallel set plots may show "blocks" of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables [@Bendix:2005]; modifications of this design, such as common-angle plots [@Hofmann:2013], address the issues which arise due to line-width illusions [@sine, @sineillusion].
+Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp].
+In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
+}}
+
+{\textcolor{red}{We also anticipate that users will be interested in comparing their town to other, similar towns.
+We will discuss the different ways that this comparison strategy was implemented in each dashboard in the next section, which describes the evolution of the dashboard over time and accounting for feedback from users and other researchers on the wider project.}}
+
+{\textcolor{red}{One final component of this project is that our dashboard is part of a wider effort to work with towns to understand the different strategies used to maintain resident quality of life amid shrinking populations.
+Thus, while the town leaders are our primary audience, we also are creating this applet for use in parallel with a team of other researchers: sociologists, economists, city planning specialists, and artists.
+These researchers opinions and feedback about the dashboard are also useful and important, as they regularly work with town leaders in different capacities and have an understanding of what factors are most important to them and what types of questions these leaders may have when faced with data and unfamiliar statistical visualizations.}}
+
+{\textcolor{red}{Throughout the design process,}} we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+
+\hypertarget{guiding-design-principles}{%
+\subsection{Guiding Design Principles}\label{guiding-design-principles}}
+
+Research on dashboard creation and interactive visualization tends to be very task-specific and {\textcolor{red}{hard to apply to more generalized settings}}.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
@@
 
-\subsection{Dashboard Design Process Overview}
+{\textcolor{red}{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
+However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}}
 
-We provided users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
-The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.
+\hypertarget{dashboard-design-process}{%
+\section{Dashboard Design Process}\label{dashboard-design-process}}
+
+\hypertarget{dashboard-components}{%
+\subsection{Dashboard Components}\label{dashboard-components}}
+
+{\textcolor{red}{In this section, we discuss the philosophy behind the basic "building blocks" of the dashboard.
+This philosophy is present in all of the iterations of the dashboard that we present in this discussion, and we will evaluate the overall philosophy's effectiveness in the conclusion.}}
+
+The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.
+
+Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
+{\textcolor{red}{As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest.}} {\textcolor{red}{This design is based on research into visualization sensemaking}}\citep{lee}, {\textcolor{red}{in that we allow users to explore outward from the familiar to the unknown}}.
+The map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
+{\textcolor{red}{XXX cite R, use consistent name for the app, cite OSRM and expand the acronym, and exaplain that in some communities, "as the crow flies" distance isn't a good metric...
+road distance is.
+XXX}}
+
+{\textcolor{red}{When faced with the next component, a parallel coordinate plot (PCP), a novice user}} will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
+
+Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
+{\textcolor{red}{In version 2, as shown in}} \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.\} {\textcolor{red}{This relatively straightforward display is}} followed by a parallel coordinate plot that allows the {\textcolor{red}{user}} to see {\textcolor{red}{similar}} towns along dimensions such as economic indicators or population size.
+
+\hypertarget{initial-draft}{%
+\subsection{Initial Draft}\label{initial-draft}}
+
 {\textcolor{red}{The initial design sketch and implementation are shown in}} \autoref{fig:v1}.
 
-We {\textcolor{red}{implemented some suggested town comparisons using}} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
-{\textcolor{red}{We initially focused on}} determining the next five to ten similar towns, based on distances to services.
-We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard.
-{\textcolor{red}{This feature became an important diagnostic for our data quality, but was not particularly useful to the application users, as they already know what towns they want to compare to; our suggested comparisons were just in the way.}}
-
-{\textcolor{red}{We will conclude with an assessment of the initial layout design based on both direct user feedback through the survey form and based on UI clicks that provide a passive feedback mechanism to indicate what elements of the UI the users tend to interact with.
-We will also discuss verbal feedback we received when testing the application.}}
-
-{\textcolor{red}{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
-However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}}
+{\textcolor{red}{Users' towns}} are at the center of our application, and comparisons to other, similar towns are secondary.
+As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
 
@@
 
+We {\textcolor{red}{implemented some suggested town comparisons using}} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
+{\textcolor{red}{We initially focused on}} determining the next five to ten similar towns, based on distances to services.
+{\textcolor{red}{This feature became an important diagnostic for our data quality, as it became clear that towns which were grouped with big cities but which did not have a large population were so grouped because of missing data.}}{\textcolor{red}{Unfortunately, this clustering feature was not as useful to the application users, as they came to the dashboard with a pre-existing set of towns to compare to; our suggested comparisons were in the way.}}
+
+{\textcolor{red}{The initial dashboard design featured several responsive maps showing the distance to the nearest hospital, fire department, post office, and school.
+These maps were ineffective for several reasons: }}
+
+\begin{itemize}
+\tightlist
+\item
+  Town residents already know this information (though it was useful for us as the dashboard designers, because we aren't nearly as familiar with the 900+ small towns in Iowa)
+\item
+  We computed distance from services relative to the center of town - coordinates provided in the data from \url{data.iowa.gov}.
+Generally speaking, the post office is at the center of town and the fire department is usually very close to the center of town; these two maps were useless.
+The school and hospital maps were less useless, but still did not provide particularly useful information to people already familiar with the town.
+\item
+  It became clear that it might be more useful to show the comparison towns on a map (relative to the town of interest) so that users could compare geographical ratings for unfamiliar data to familiar data.
+\end{itemize}
+
+{\textcolor{red}{In addition, we received feedback on the parallel coordinate plot at the bottom of the app which was surprising: the viewers (in this case, other researchers on the team) were not as intimidated by the parallel coordinate plot as we had expected.
+They did need some explanation of how to read the plot, and these hints need to be included in the dashboard, but they grasped the fundamental idea of the plot very quickly.}}
+
+{\textcolor{red}{Our conclusion, based on this initial dashboard draft, was that we needed to restructure the application.
+Our attempt to show familiar information first to "build up" to the more unfamiliar structure of a parallel coordinate plot was not effective; there was too much clutter and not enough new information to draw users in.}}
+
+\hypertarget{redesign}{%
+\subsection{Redesign}\label{redesign}}
+
 \begin{figure}
@@
 
-\subsubsection{Dashboard Components}
+{\textcolor{red}{XXX Denise, can you follow up with the initial design section and explain the 2nd design and approach?
+Why did we change things?
+How did those changes impact feedback from users?
+E.g.
+wanting to show the whole range of values in the PCP vs.
+being visually overwhelmed.
+Incorporate the user feedback into the narrative.
+XXX}}
 
-The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
-We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-While our dashboard layout is vitally important into making sure that the town leaders are using the dashboards, we should discuss the dashboard components that we used along with the philosophy of those key components.
-Starting with our parallel coordinate plot, which will allow for objective measures related to the QoL Survey.
-As our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others.
-The framework mentioned in \texttt{lee},``How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice's Information Visualization Sensemaking,'\,' allowed our work to determine that a novice user seeing a PCP will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
-The PCP in this study were found to be promising for novice users, when comparing to the Treemap and Chord Diagram.
+\hypertarget{future-work}{%
+\section{Future Work}\label{future-work}}
 
-Furthermore, the large set of publicly available data we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
+Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
+Transition to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
+Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.
 
-\subsection{Guided Discovery Learning (GDL)}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
-We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization.
-This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \texttt{dedonno}.
-
-The Data Visualization component with GDL principles is a single page which populates the information in maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals.
-Value Boxes populate vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as similar economy or population size.
-The graphical representations including a parallel coordinate plot and geographic maps {[}Figure 1{]}.
-The Rural Smart EDA Dashboard map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
-
-The Exploratory Data Analysis component utilizing statistical clustering methods enables the town leaders to interact with the parallel coordinate plots at various levels of granularities of data, with constraints, to similar towns.
-The Rural Smart EDA Dashboard features offer geospatial and structural data to assist town leaders in gaining a comprehensive understanding of the town in which they support.
-
-Each component provides the town leaders with analytical aspects related to the performance of their town.
-As town leaders interact with the tool more helps build the knowledge and understanding different dimensions and measures related to the shrinkage of the town to a particular indicator.
+Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
+Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
+Test early, test often\ldots{}
 
@@
 
-Using the framework from ``Creating design guidelines for building city dashboards from a user's perspectives,'' \texttt{young}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
+Using the framework from ``Creating design guidelines for building city dashboards from a user's perspectives'' \citep{lee}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
 We will adapt these questions to better help the town leaders communicate clear feedback for our team to use.

### 2022-07-30 SDSS Journal Paper/Paper Draft.Rmd
Updates from brainstorming meeting

@@
     marker: '1'               # identifier for the affiliation
-    footnote-id: 1              # optional identifier for the footnote
-    email: denise.bradford@huskers.unl.edu    # optional email address
+    #email: denise.bradford@huskers.unl.edu    # optional email address
   - first: Susan
@@
     footnote-id: 1
+    email: svanderplas@unl.edu 
 footnote:
@@
 # Introduction
-\svp{As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
-As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.}
+As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
+As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.
 <!-- With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{fisher}.  -->
-\svp{In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.}
+In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.
 <!-- A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.  -->
-\svp{Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity.
+Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity.
 One glance at a car's dashboard will tell you the speed, RPM, engine temperature, amount of gas in the tank; more importantly, however, the goal is not for the user to remember all of these characteristics, but to assess whether any of these quantities is outside of the expected range.
-Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.}
+Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.
 <!-- An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.  -->
-\svp{In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.}
+In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.
 
@@
 This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades.
-As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}.
-We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life. 
-In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
-\svp{This paragraph is pretty good, but see if you can tweak it to blend in a bit better with the edits I made above.}
+As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations [@scc].
+We hope to allow towns to explore their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
 
@@
 
-\svp{The Smart and Connected Community (SCC) dashaboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.} Most of these data sets are collected at a town/city or county \svp{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}.
-\svp{In addition to the more commonly available statistics derived from e.g.
-the census and American Community Survey,} \url{data.iowa.gov} contains \svp{several} unique \svp{data sets}, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+The Smart and Connected Community (SCC) dashaboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
+Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level.
+In addition to the more commonly available statistics derived from e.g.
+the census and American Community Survey, \url{data.iowa.gov} contains several unique data sets, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
 
-\svp{Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.}
+Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
 Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
 The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
-\svp{Cite Dave's papers about the ISTP QOL metrics, including education}
-<!-- These data are used for details are a town level, that will provide information about the quality of life. -->
+\db{[@petersCommunityResiliencyDeclining2019] -- Change the reference}
+<!-- \svp{Cite Dave's papers about the ISTP QOL metrics, including education}
+ These data are used for details are a town level, that will provide information about the quality of life. -->
 
-Data collected from the Index of Relative Rurality (IRR) \svp{Create a bibtex entry for this and cite it properly, please.
-Same goes for most of the other URLs.} \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns.
+Data collected from the Index of Relative Rurality (IRR) [@Rural_classification]  were used in the SCC dashboard to help classify the towns.
 The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality.
@@
 
-Data collected from Rural Urban Commuting Area Codes \texttt{usda} were used to help identify towns with commuting behaviors in our rural areas.
+Data collected from Rural Urban Commuting Area Codes [@usda] were used to help identify towns with commuting behaviors in our rural areas.
 The rural-urban commuting area (RUCA) codes classify U.S.
@@
 Thus, in addition to the usual challenges of working with real-world data that is "messy" in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
-This makes both visualization and statistical analysis more complicated \svp{(and more interesting)}.
+This makes both visualization and statistical analysis more complicated (and more interesting).
 
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{Key_Metrics}
+\includegraphics[width=.8\textwidth]{Key_Metrics}
 \caption{Diagram of considerations for our dashboard design process.}\label{fig:metrics}
@@
 
-\svp{One problem we identified early in the process of assessing smart-shrinkage strategies in small towns is that these towns do not have the resources to make data-driven decisions.
+One problem we identified early in the process of assessing smart-shrinkage strategies in small towns is that these towns do not have the resources to make data-driven decisions.
 Typically, small towns in Iowa are managed by at most a few part-time employees or volunteers.
-In some cases, essential management functions of the town are paid, but the municipalities we are interested in do not have sufficient funding to hire professionals to gather and analyze data.}
+In some cases, essential management functions of the town are paid, but the municipalities we are interested in do not have sufficient funding to hire professionals to gather and analyze data.
 
-\svp{As part of a wider project investigating the strategies towns use to maintain quality of life amid shrinking population, our research team provides communities with data about their own town, but also comparable towns across the state which may have a different approach to city services.
-In combination with other engagement strategies that are more qualitative, we hope to use this interactive dashboard approach to assist small Iowa cities with generalizing and developing strategies to improve or maintain quality of life amid shrinking populations.}
+As part of a wider project investigating the strategies towns use to maintain quality of life amid shrinking population, our research team provides communities with data about their own town, but also comparable towns across the state which may have a different approach to city services.
+In combination with other engagement strategies that are more qualitative, we hope to use this interactive dashboard approach to assist small Iowa cities with generalizing and developing strategies to improve or maintain quality of life amid shrinking populations.
 
-\svp{One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data.
-At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.}
-\svp{This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a} framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
+One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data.
+At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
+This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
 
-\svp{We expect that users will be interested in "sets" of variables from the wider dataset, which we assembled based on quality of life factors in the Iowa Small Town Poll [@petersCommunityResiliencyDeclining2019].
-For instance, users might be interested in medical and social services available to residents, such as a local primary care clinic, nursing homes which are within driving distance, and the distance to the nearest emergency room; these factors might be explored separately from variables describing the services provided directly by city government, such as parks and recreation expenditures, snow removal services, and the distance to the closest fire station.}
+We expect that users will be interested in "sets" of variables from the wider dataset, which we assembled based on quality of life factors in the Iowa Small Town Poll [@petersCommunityResiliencyDeclining2019].
+For instance, users might be interested in medical and social services available to residents, such as a local primary care clinic, nursing homes which are within driving distance, and the distance to the nearest emergency room; these factors might be explored separately from variables describing the services provided directly by city government, such as parks and recreation expenditures, snow removal services, and the distance to the closest fire station.
 
-\svp{As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have.
+As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have.
 Introduced in the 1880s [@dOcagne:1885], parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation.
@@
 In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
-}
 
-\svp{We also anticipate that users will be interested in comparing their town to other, similar towns.
-We will discuss the different ways that this comparison strategy was implemented in each dashboard in the next section, which describes the evolution of the dashboard over time and accounting for feedback from users and other researchers on the wider project.}
+We also anticipate that users will be interested in comparing their town to other, similar towns.
+We will discuss the different ways that this comparison strategy was implemented in each dashboard in the next section, which describes the evolution of the dashboard over time and accounting for feedback from users and other researchers on the wider project.
 
-\svp{One final component of this project is that our dashboard is part of a wider effort to work with towns to understand the different strategies used to maintain resident quality of life amid shrinking populations.
+One final component of this project is that our dashboard is part of a wider effort to work with towns to understand the different strategies used to maintain resident quality of life amid shrinking populations.
 Thus, while the town leaders are our primary audience, we also are creating this applet for use in parallel with a team of other researchers: sociologists, economists, city planning specialists, and artists.
-These researchers opinions and feedback about the dashboard are also useful and important, as they regularly work with town leaders in different capacities and have an understanding of what factors are most important to them and what types of questions these leaders may have when faced with data and unfamiliar statistical visualizations.}
+These researchers opinions and feedback about the dashboard are also useful and important, as they regularly work with town leaders in different capacities and have an understanding of what factors are most important to them and what types of questions these leaders may have when faced with data and unfamiliar statistical visualizations.
 
-\svp{Throughout the design process,} we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+Throughout the design process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
 
@@
 
-Research on dashboard creation and interactive visualization tends to be very task-specific and \svp{hard to apply to more generalized settings}.
+Research on dashboard creation and interactive visualization tends to be very task-specific and hard to apply to more generalized settings.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
-With this in mind, we \svp{set out to clearly document our} intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
+With this in mind, we set out to clearly document our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
 
@@
 \item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
-\item \svp{We should facilitate} comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
-\item \svp{We will present} the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
-\item \svp{We will implement feedback mechanisms that allow us to provide} more detailed data and \svp{respond to} feature requests to improve the dashboard design over time.
+\item We should facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+\item We will present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+\item We will implement feedback mechanisms that allow us to provide more detailed data and respond to feature requests to improve the dashboard design over time.
 \end{itemize}
 
-\svp{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
-However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}
+As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
+However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.
 
-# Dashboard Design Process
+\section{ Dashboard Design Process}
 
-## Dashboard Components
+\subsection{Dashboard Components }
 
-\svp{In this section, we discuss the philosophy behind the basic "building blocks" of the dashboard.
-This philosophy is present in all of the iterations of the dashboard that we present in this discussion, and we will evaluate the overall philosophy's effectiveness in the conclusion.}
+In this section, we discuss the philosophy behind the basic "building blocks" of the dashboard.
+This philosophy is present in all of the iterations of the dashboard that we present in this discussion, and we will evaluate the overall philosophy's effectiveness in the conclusion.
 
@@
 We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload [@few].
 
 Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-\svp{As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest.} \svp{This design is based on research into visualization sensemaking}[@lee], \svp{in that we allow users to explore outward from the familiar to the unknown}.
-The map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
-\svp{XXX cite R, use consistent name for the app, cite OSRM and expand the acronym, and exaplain that in some communities, "as the crow flies" distance isn't a good metric... road distance is. XXX}
+As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest.
+This design is based on research into visualization sensemaking [@lee], in that we allow users to explore outward from the familiar to the unknown.
+\db{The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point.
+OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.}
+
 <!-- \svp{As PCPs are likely to be unfamiliar to most users, we decided to make the first part of the app less visually intimidating, providing maps showing the distance to local services in version 1 (\autoref{fig:v1}), and showing a condensed service map along with basic facts about the town in version 2 (\autoref{fig:v2}).} -->
@@
 <!-- We believe that this will allow the town leaders the ability to see as much data as possible with relation to their towns and others.  -->
-\svp{When faced with the next component, a parallel coordinate plot (PCP), a novice user} will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
+When faced with the next component, a parallel coordinate plot (PCP), a novice user will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
 
 Taken together, the app is a single page; the initial "solid ground" which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
-\svp{In version 2, as shown in} \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.} \svp{This relatively straightforward display is} followed by a parallel coordinate plot that allows the \svp{user} to see \svp{similar} towns along dimensions such as economic indicators or population size.
+In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.
+This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.
 
-## Initial Draft
+\subsection{Initial Draft}
 
-\svp{The initial design sketch and implementation are shown in} \autoref{fig:v1}.
+The initial design sketch and implementation are shown in \autoref{fig:v1}.
 
-\svp{Users' towns} are at the center of our application, and comparisons to other, similar towns are secondary.
+\svp{Users' towns} \db{Town Users' ??} are at the center of our application, and comparisons to other, similar towns are secondary.
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
@@
 
-We \svp{implemented some suggested town comparisons using} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
-\svp{We initially focused on} determining the next five to ten similar towns, based on distances to services.
-\svp{This feature became an important diagnostic for our data quality, as it became clear that towns which were grouped with big cities but which did not have a large population were so grouped because of missing data.}\svp{Unfortunately, this clustering feature was not as useful to the application users, as they came to the dashboard with a pre-existing set of towns to compare to; our suggested comparisons were in the way.}
+We implemented some suggested town comparisons using unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
+We initially focused on determining the next five to ten similar towns, based on distances to services.
+This feature became an important diagnostic for our data quality, as it became clear that towns which were grouped with big cities but which did not have a large population were so grouped because of missing data.
+Unfortunately, this clustering feature was not as useful to the application users, as they came to the dashboard with a pre-existing set of towns to compare to; our suggested comparisons were in the way.
 
-\svp{The initial dashboard design featured several responsive maps showing the distance to the nearest hospital, fire department, post office, and school.
-These maps were ineffective for several reasons: }
+The initial dashboard design featured several responsive maps showing the distance to the nearest hospital, fire department, post office, and school. These maps were ineffective for several reasons:
 
@@
 
-\svp{In addition, we received feedback on the parallel coordinate plot at the bottom of the app which was surprising: the viewers (in this case, other researchers on the team) were not as intimidated by the parallel coordinate plot as we had expected.
-They did need some explanation of how to read the plot, and these hints need to be included in the dashboard, but they grasped the fundamental idea of the plot very quickly.}
+In addition, we received feedback on the parallel coordinate plot at the bottom of the app which was surprising: the viewers (in this case, other researchers on the team) were not as intimidated by the parallel coordinate plot as we had expected.
+They did need some explanation of how to read the plot, and these hints need to be included in the dashboard, but they grasped the fundamental idea of the plot very quickly.
 
-\svp{Our conclusion, based on this initial dashboard draft, was that we needed to restructure the application.
-Our attempt to show familiar information first to "build up" to the more unfamiliar structure of a parallel coordinate plot was not effective; there was too much clutter and not enough new information to draw users in.}
+Our conclusion, based on this initial dashboard draft, was that we needed to restructure the application.
+Our attempt to show familiar information first to "build up" to the more unfamiliar structure of a parallel coordinate plot was not effective; there was too much clutter and not enough new information to draw users in.
 
@@
 
-## Redesign
+\subsection{ Redesign }
 
@@
 
-# Discussion
+\db{In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
+As a result, we cached map directions from OSRM for each service in our database, which drastically changed the response time for the user.
+Our initial design lack proper dashboard design, which would allow the user's eye to focus on the most important parts of the dashboard.
+The redesign allowed for a cleaner box from the top to the bottom.
+}
+
+\db{In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
+The rurality index allowed for a better classification index, the USDA source also allowed for the town to understand how much or little the closest major city is impacting their towns.
+Town members are commuting into the next town for work or resources, this could be helpful to know which services and resources to add into their own town.}
+
+\db{In regards to the Parallel Coordinate Plots, we wanted to declutter the aspects of the x- and y-axis that will allow for a less overwhelming experience for our users.
+Our x-axis had a large number of variables that our team precevied to be the best variables.
+This sparked the need to make sure that only necessary variables are placed in the graph, along with the way they were placed along the x-axis.
+We decided to limit the variables to those open source variables that were related directly to QoL.
+The variables were then placed in a bulks based on QoL topic.}
+
+\section{ Discussion}
 How did the philosophy work out?
@@
 
-# Future Work
+\section{ Future Work }
 
@@
 
-# Conclusions
+\section {Conclusions}
 Specific takeaways - visualization research (broad) vs.

### 2022-07-30 SDSS Journal Paper/Paper-Draft.tex
Updates from brainstorming meeting

@@
   \fnms{Denise}
-  \snm{Bradford}  \thanksref{1}  \ead{denise.bradford@huskers.unl.edu}}
+  \snm{Bradford}}
 \author[1]{
@@
   \fnms{Susan}
-  \snm{VanderPlas}  \thanksref{1}}
+  \snm{VanderPlas}  \thanksref{1}  \ead{svanderplas@unl.edu}}
 
@@
 
-{\textcolor{red}{As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
-As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.}}
-
-{\textcolor{red}{In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.}}
-
-{\textcolor{red}{Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity.
+As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important.
+As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.
+In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.
+Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity.
 One glance at a car's dashboard will tell you the speed, RPM, engine temperature, amount of gas in the tank; more importantly, however, the goal is not for the user to remember all of these characteristics, but to assess whether any of these quantities is outside of the expected range.
-Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.}}
-
-{\textcolor{red}{In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.}}
+Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.
+In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.
 
@@
 As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}.
-We hope to allow towns to discover their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
-In the process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
-{\textcolor{red}{This paragraph is pretty good, but see if you can tweak it to blend in a bit better with the edits I made above.}}
+We hope to allow towns to explore their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
 
@@
 
-{\textcolor{red}{The Smart and Connected Community (SCC) dashaboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.}} Most of these data sets are collected at a town/city or county {\textcolor{red}{spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level}}.
-{\textcolor{red}{In addition to the more commonly available statistics derived from e.g.
-the census and American Community Survey,}} \url{data.iowa.gov} contains {\textcolor{red}{several}} unique {\textcolor{red}{data sets}}, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
+The Smart and Connected Community (SCC) dashaboard data are primarily assembled from \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
+Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level.
+In addition to the more commonly available statistics derived from e.g.~the census and American Community Survey, \url{data.iowa.gov} contains several unique data sets, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.
 
-{\textcolor{red}{Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.}}
-
+Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
 Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
 The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
-{\textcolor{red}{Cite Dave's papers about the ISTP QOL metrics, including education}}
+{\textcolor{blue}{[@petersCommunityResiliencyDeclining2019] -- Change the reference}}
 
-Data collected from the Index of Relative Rurality (IRR) {\textcolor{red}{Create a bibtex entry for this and cite it properly, please.
-Same goes for most of the other URLs.}} \url{https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/} were used in the SCC dashboard to help classify the towns.
+Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns.
 The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality.
@@
 
-Data collected from Rural Urban Commuting Area Codes \texttt{usda} were used to help identify towns with commuting behaviors in our rural areas.
+Data collected from Rural Urban Commuting Area Codes \citep{usda} were used to help identify towns with commuting behaviors in our rural areas.
 The rural-urban commuting area (RUCA) codes classify U.S.
@@
 Thus, in addition to the usual challenges of working with real-world data that is ``messy'' in a variety of ways, we also have to contend with missing data that is missing due to the size of the community or the lack of services.
-This makes both visualization and statistical analysis more complicated {\textcolor{red}{(and more interesting)}}.
+This makes both visualization and statistical analysis more complicated (and more interesting).
 
@@
 
-{\textcolor{red}{One problem we identified early in the process of assessing smart-shrinkage strategies in small towns is that these towns do not have the resources to make data-driven decisions.
+One problem we identified early in the process of assessing smart-shrinkage strategies in small towns is that these towns do not have the resources to make data-driven decisions.
 Typically, small towns in Iowa are managed by at most a few part-time employees or volunteers.
-In some cases, essential management functions of the town are paid, but the municipalities we are interested in do not have sufficient funding to hire professionals to gather and analyze data.}}
+In some cases, essential management functions of the town are paid, but the municipalities we are interested in do not have sufficient funding to hire professionals to gather and analyze data.
 
-{\textcolor{red}{As part of a wider project investigating the strategies towns use to maintain quality of life amid shrinking population, our research team provides communities with data about their own town, but also comparable towns across the state which may have a different approach to city services.
-In combination with other engagement strategies that are more qualitative, we hope to use this interactive dashboard approach to assist small Iowa cities with generalizing and developing strategies to improve or maintain quality of life amid shrinking populations.}}
+As part of a wider project investigating the strategies towns use to maintain quality of life amid shrinking population, our research team provides communities with data about their own town, but also comparable towns across the state which may have a different approach to city services.
+In combination with other engagement strategies that are more qualitative, we hope to use this interactive dashboard approach to assist small Iowa cities with generalizing and developing strategies to improve or maintain quality of life amid shrinking populations.
 
-{\textcolor{red}{One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data.
-At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.}}
+One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data.
+At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
+This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
 
-{\textcolor{red}{This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a}} framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
+We expect that users will be interested in ``sets'' of variables from the wider dataset, which we assembled based on quality of life factors in the Iowa Small Town Poll \citep{petersCommunityResiliencyDeclining2019}.
+For instance, users might be interested in medical and social services available to residents, such as a local primary care clinic, nursing homes which are within driving distance, and the distance to the nearest emergency room; these factors might be explored separately from variables describing the services provided directly by city government, such as parks and recreation expenditures, snow removal services, and the distance to the closest fire station.
 
-{\textcolor{red}{We expect that users will be interested in "sets" of variables from the wider dataset, which we assembled based on quality of life factors in the Iowa Small Town Poll [@petersCommunityResiliencyDeclining2019].
-For instance, users might be interested in medical and social services available to residents, such as a local primary care clinic, nursing homes which are within driving distance, and the distance to the nearest emergency room; these factors might be explored separately from variables describing the services provided directly by city government, such as parks and recreation expenditures, snow removal services, and the distance to the closest fire station.}}
+As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours \citep{tourr}, require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have.
+Introduced in the 1880s \citep{dOcagne:1885}, parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation.
+When representing categorical data, parallel set plots may show ``blocks'' of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables \citep{Bendix:2005}; modifications of this design, such as common-angle plots \citep{Hofmann:2013}, address the issues which arise due to line-width illusions \citep[\citet{sineillusion}]{sine}.
+Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}.
+In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
 
-{\textcolor{red}{As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have.
-Introduced in the 1880s [@dOcagne:1885], parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation.
-When representing categorical data, parallel set plots may show "blocks" of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables [@Bendix:2005]; modifications of this design, such as common-angle plots [@Hofmann:2013], address the issues which arise due to line-width illusions [@sine, @sineillusion].
-Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp].
-In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
-}}
+We also anticipate that users will be interested in comparing their town to other, similar towns.
+We will discuss the different ways that this comparison strategy was implemented in each dashboard in the next section, which describes the evolution of the dashboard over time and accounting for feedback from users and other researchers on the wider project.
 
-{\textcolor{red}{We also anticipate that users will be interested in comparing their town to other, similar towns.
-We will discuss the different ways that this comparison strategy was implemented in each dashboard in the next section, which describes the evolution of the dashboard over time and accounting for feedback from users and other researchers on the wider project.}}
+One final component of this project is that our dashboard is part of a wider effort to work with towns to understand the different strategies used to maintain resident quality of life amid shrinking populations.
+Thus, while the town leaders are our primary audience, we also are creating this applet for use in parallel with a team of other researchers: sociologists, economists, city planning specialists, and artists.
+These researchers opinions and feedback about the dashboard are also useful and important, as they regularly work with town leaders in different capacities and have an understanding of what factors are most important to them and what types of questions these leaders may have when faced with data and unfamiliar statistical visualizations.
 
-{\textcolor{red}{One final component of this project is that our dashboard is part of a wider effort to work with towns to understand the different strategies used to maintain resident quality of life amid shrinking populations.
-Thus, while the town leaders are our primary audience, we also are creating this applet for use in parallel with a team of other researchers: sociologists, economists, city planning specialists, and artists.
-These researchers opinions and feedback about the dashboard are also useful and important, as they regularly work with town leaders in different capacities and have an understanding of what factors are most important to them and what types of questions these leaders may have when faced with data and unfamiliar statistical visualizations.}}
-
-{\textcolor{red}{Throughout the design process,}} we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
+Throughout the design process, we will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
 
@@
 
-Research on dashboard creation and interactive visualization tends to be very task-specific and {\textcolor{red}{hard to apply to more generalized settings}}.
+Research on dashboard creation and interactive visualization tends to be very task-specific and hard to apply to more generalized settings.
 That is, it is relatively easy to create a dashboard that works for a particular task, but it is hard to generalize from that process what will work for the next dashboard.
-With this in mind, we {\textcolor{red}{set out to clearly document our}} intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
+With this in mind, we set out to clearly document our intentions at each stage of the design and evaluation process, with the goal of gathering some useful information about general dashboard design from the process of creating this specific dashboard.
 
@@
 \item The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
-\item {\textcolor{red}{We should facilitate}} comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
-\item {\textcolor{red}{We will present}} the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
-\item {\textcolor{red}{We will implement feedback mechanisms that allow us to provide}} more detailed data and {\textcolor{red}{respond to}} feature requests to improve the dashboard design over time.
+\item We should facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
+\item We will present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
+\item We will implement feedback mechanisms that allow us to provide more detailed data and respond to feature requests to improve the dashboard design over time.
 \end{itemize}
 
-{\textcolor{red}{As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a "final" dashboard design because the application will continue to evolve.
-However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.}}
+As with many dashboards, this project is under continuous development; while it makes for an unsatisfactory conclusion, we do not have a ``final'' dashboard design because the application will continue to evolve.
+However, we have some useful insights into the process of creating an application designed to invite users to explore a large and complex dataset that we believe to be a useful contribution to work in this area.
 
-\hypertarget{dashboard-design-process}{%
-\section{Dashboard Design Process}\label{dashboard-design-process}}
+\section{ Dashboard Design Process}
 
-\hypertarget{dashboard-components}{%
-\subsection{Dashboard Components}\label{dashboard-components}}
+\subsection{Dashboard Components }
 
-{\textcolor{red}{In this section, we discuss the philosophy behind the basic "building blocks" of the dashboard.
-This philosophy is present in all of the iterations of the dashboard that we present in this discussion, and we will evaluate the overall philosophy's effectiveness in the conclusion.}}
+In this section, we discuss the philosophy behind the basic ``building blocks'' of the dashboard.
+This philosophy is present in all of the iterations of the dashboard that we present in this discussion, and we will evaluate the overall philosophy's effectiveness in the conclusion.
 
@@
 We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
-This avoids some of the pitfalls of dashboard design that can easily lead to user overload \texttt{few}.
+This avoids some of the pitfalls of dashboard design that can easily lead to user overload \citep{few}.
 
 Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary.
-{\textcolor{red}{As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest.}} {\textcolor{red}{This design is based on research into visualization sensemaking}}\citep{lee}, {\textcolor{red}{in that we allow users to explore outward from the familiar to the unknown}}.
-The map visuals were built using OSRM route functions in R to amplify the accuracy of the distances from necessary services in town-centric point.
-{\textcolor{red}{XXX cite R, use consistent name for the app, cite OSRM and expand the acronym, and exaplain that in some communities, "as the crow flies" distance isn't a good metric...
-road distance is.
-XXX}}
+As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest.
+This design is based on research into visualization sensemaking \citep{lee}, in that we allow users to explore outward from the familiar to the unknown.
+{\textcolor{blue}{The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point.
+OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.}}
 
-{\textcolor{red}{When faced with the next component, a parallel coordinate plot (PCP), a novice user}} will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
+When faced with the next component, a parallel coordinate plot (PCP), a novice user will be able to determine two basic components: Visual Object (textual objects and non-textual objects) and Frame (frame of content and frame of visual encoding).
 
 Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
-{\textcolor{red}{In version 2, as shown in}} \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.\} {\textcolor{red}{This relatively straightforward display is}} followed by a parallel coordinate plot that allows the {\textcolor{red}{user}} to see {\textcolor{red}{similar}} towns along dimensions such as economic indicators or population size.
+In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.
+This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.
 
-\hypertarget{initial-draft}{%
-\subsection{Initial Draft}\label{initial-draft}}
+\subsection{Initial Draft}
 
-{\textcolor{red}{The initial design sketch and implementation are shown in}} \autoref{fig:v1}.
+The initial design sketch and implementation are shown in \autoref{fig:v1}.
 
-{\textcolor{red}{Users' towns}} are at the center of our application, and comparisons to other, similar towns are secondary.
+{\textcolor{red}{Users' towns}} {\textcolor{blue}{Town Users' ??}} are at the center of our application, and comparisons to other, similar towns are secondary.
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
@@
 
-We {\textcolor{red}{implemented some suggested town comparisons using}} unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
-{\textcolor{red}{We initially focused on}} determining the next five to ten similar towns, based on distances to services.
-{\textcolor{red}{This feature became an important diagnostic for our data quality, as it became clear that towns which were grouped with big cities but which did not have a large population were so grouped because of missing data.}}{\textcolor{red}{Unfortunately, this clustering feature was not as useful to the application users, as they came to the dashboard with a pre-existing set of towns to compare to; our suggested comparisons were in the way.}}
+We implemented some suggested town comparisons using unsupervised clustering methods to help our towns make decisions that are informed in comparison to similar towns, for budget size, population size and location.
+We initially focused on determining the next five to ten similar towns, based on distances to services.
+This feature became an important diagnostic for our data quality, as it became clear that towns which were grouped with big cities but which did not have a large population were so grouped because of missing data.
+Unfortunately, this clustering feature was not as useful to the application users, as they came to the dashboard with a pre-existing set of towns to compare to; our suggested comparisons were in the way.
 
-{\textcolor{red}{The initial dashboard design featured several responsive maps showing the distance to the nearest hospital, fire department, post office, and school.
-These maps were ineffective for several reasons: }}
+The initial dashboard design featured several responsive maps showing the distance to the nearest hospital, fire department, post office, and school. These maps were ineffective for several reasons:
 
@@
 
-{\textcolor{red}{In addition, we received feedback on the parallel coordinate plot at the bottom of the app which was surprising: the viewers (in this case, other researchers on the team) were not as intimidated by the parallel coordinate plot as we had expected.
-They did need some explanation of how to read the plot, and these hints need to be included in the dashboard, but they grasped the fundamental idea of the plot very quickly.}}
+In addition, we received feedback on the parallel coordinate plot at the bottom of the app which was surprising: the viewers (in this case, other researchers on the team) were not as intimidated by the parallel coordinate plot as we had expected.
+They did need some explanation of how to read the plot, and these hints need to be included in the dashboard, but they grasped the fundamental idea of the plot very quickly.
 
-{\textcolor{red}{Our conclusion, based on this initial dashboard draft, was that we needed to restructure the application.
-Our attempt to show familiar information first to "build up" to the more unfamiliar structure of a parallel coordinate plot was not effective; there was too much clutter and not enough new information to draw users in.}}
+Our conclusion, based on this initial dashboard draft, was that we needed to restructure the application.
+Our attempt to show familiar information first to ``build up'' to the more unfamiliar structure of a parallel coordinate plot was not effective; there was too much clutter and not enough new information to draw users in.
 
-\hypertarget{redesign}{%
-\subsection{Redesign}\label{redesign}}
+\subsection{ Redesign }
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+{\textcolor{blue}{In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
+As a result, we cached map directions from OSRM for each service in our database, which drastically changed the response time for the user.
+Our initial design lack proper dashboard design, which would allow the user's eye to focus on the most important parts of the dashboard.
+The redesign allowed for a cleaner box from the top to the bottom.
+}}
 
-\hypertarget{future-work}{%
-\section{Future Work}\label{future-work}}
+{\textcolor{blue}{In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
+The rurality index allowed for a better classification index, the USDA source also allowed for the town to understand how much or little the closest major city is impacting their towns.
+Town members are commuting into the next town for work or resources, this could be helpful to know which services and resources to add into their own town.}}
+
+{\textcolor{blue}{In regards to the Parallel Coordinate Plots, we wanted to declutter the aspects of the x- and y-axis that will allow for a less overwhelming experience for our users.
+Our x-axis had a large number of variables that our team precevied to be the best variables.
+This sparked the need to make sure that only necessary variables are placed in the graph, along with the way they were placed along the x-axis.
+We decided to limit the variables to those open source variables that were related directly to QoL.
+The variables were then placed in a bulks based on QoL topic.}}
+
+\section{ Discussion}
+
+How did the philosophy work out?
+What worked and didn't (overall)?
+People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with?
+Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis.
+What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare.
+We also should provide much, much more user support for how to interact with and interpret the app and results.
+
+\section{ Future Work }
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section {Conclusions}
 

### 2022-07-31 SDSS Journal Paper/Paper Draft.Rmd
A few structural changes to the second version

@@
 \db{In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
-As a result, we cached map directions from OSRM for each service in our database, which drastically changed the response time for the user.
-Our initial design lack proper dashboard design, which would allow the user's eye to focus on the most important parts of the dashboard.
-The redesign allowed for a cleaner box from the top to the bottom.
+As a result, we cached map directions from OSRM for each service in our database, which drastically} \svp{reduced} \db{the response time for the user.
+Our initial design \svp{did not naturally focus} the user's eye on the most important parts of the dashboard; the redesign allowed for a cleaner \svp{flow} from the top to the bottom.
 }
@@
 \db{In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
-The rurality index allowed for a better classification index, the USDA source also allowed for the town to understand how much or little the closest major city is impacting their towns.
-Town members are commuting into the next town for work or resources, this could be helpful to know which services and resources to add into their own town.}
+The rurality index \svp{provided a better classification} and the USDA sources allowed for the town to understand \svp{the impact of} the closest major city \svp{due to commuting for work and shopping at larger stores not available within the town}.}
 
-\db{In regards to the Parallel Coordinate Plots, we wanted to declutter the aspects of the x- and y-axis that will allow for a less overwhelming experience for our users.
+\svp{We also modified the parallel coordinate plots in several ways:}
+
+- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with quality of life.
+However, there were still too many variables for users to successfully parse.
+We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by quality of life factor [@petersCommunityResiliencyDeclining2019].
+- Originally, parallel coordinate bands were scaled based on the selected comparison towns.
+This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set.
+We chose to show all towns in the data set in a very light $\alpha$ grey color to provide some information about the overall range of each variable.
+Unfortunately, even with the low-$\alpha$ value, this increased the visual complexity of the plot and confused users.
+Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user.
+This should strike a balance between visual complexity and representing the data accurately.
+- We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly.
+Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.
+
+<!-- \db{In regards to the Parallel Coordinate Plots, we wanted to declutter the aspects of the x- and y-axis that will allow for a less overwhelming experience for our users.
 Our x-axis had a large number of variables that our team precevied to be the best variables.
@@
 We decided to limit the variables to those open source variables that were related directly to QoL.
-The variables were then placed in a bulks based on QoL topic.}
+The variables were then placed in a bulks based on QoL topic.} -->
+
+\svp{While not all of these modifications were well received in our second round of user testing, the changes did incrementally move the dashboard display towards our goal of allowing users to explore the data and engage with it.
+We continued to be surprised with how well users reacted to the parallel coordinate plots, which we initially thought might be too abstract for users unfamiliar with multivariate data displays, but the ability to compare towns across multiple dimensions and examine the similarities and differences between their approaches to different services seemed to be intuitive for users once they understood that each vertical axis was a different variable.}
 

### 2022-07-31 SDSS Journal Paper/Paper-Draft.tex
A few structural changes to the second version

@@
 {\textcolor{blue}{In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
-As a result, we cached map directions from OSRM for each service in our database, which drastically changed the response time for the user.
-Our initial design lack proper dashboard design, which would allow the user's eye to focus on the most important parts of the dashboard.
-The redesign allowed for a cleaner box from the top to the bottom.
+As a result, we cached map directions from OSRM for each service in our database, which drastically}} {\textcolor{red}{reduced}} {\textcolor{blue}{the response time for the user.
+Our initial design {\textcolor{red}{did not naturally focus}} the user's eye on the most important parts of the dashboard; the redesign allowed for a cleaner {\textcolor{red}{flow}} from the top to the bottom.
 }}
@@
 {\textcolor{blue}{In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
-The rurality index allowed for a better classification index, the USDA source also allowed for the town to understand how much or little the closest major city is impacting their towns.
-Town members are commuting into the next town for work or resources, this could be helpful to know which services and resources to add into their own town.}}
+The rurality index {\textcolor{red}{provided a better classification}} and the USDA sources allowed for the town to understand {\textcolor{red}{the impact of}} the closest major city {\textcolor{red}{due to commuting for work and shopping at larger stores not available within the town}}.}}
 
-{\textcolor{blue}{In regards to the Parallel Coordinate Plots, we wanted to declutter the aspects of the x- and y-axis that will allow for a less overwhelming experience for our users.
-Our x-axis had a large number of variables that our team precevied to be the best variables.
-This sparked the need to make sure that only necessary variables are placed in the graph, along with the way they were placed along the x-axis.
-We decided to limit the variables to those open source variables that were related directly to QoL.
-The variables were then placed in a bulks based on QoL topic.}}
+{\textcolor{red}{We also modified the parallel coordinate plots in several ways:}}
+
+\begin{itemize}
+\tightlist
+\item
+  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with quality of life.
+However, there were still too many variables for users to successfully parse.
+We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by quality of life factor \citep{petersCommunityResiliencyDeclining2019}.
+\item
+  Originally, parallel coordinate bands were scaled based on the selected comparison towns.
+This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set.
+We chose to show all towns in the data set in a very light \(\alpha\) grey color to provide some information about the overall range of each variable.
+Unfortunately, even with the low-\(\alpha\) value, this increased the visual complexity of the plot and confused users.
+Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user.
+This should strike a balance between visual complexity and representing the data accurately.
+\item
+  We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly.
+Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.
+\end{itemize}
+
+{\textcolor{red}{While not all of these modifications were well received in our second round of user testing, the changes did incrementally move the dashboard display towards our goal of allowing users to explore the data and engage with it.
+We continued to be surprised with how well users reacted to the parallel coordinate plots, which we initially thought might be too abstract for users unfamiliar with multivariate data displays, but the ability to compare towns across multiple dimensions and examine the similarities and differences between their approaches to different services seemed to be intuitive for users once they understood that each vertical axis was a different variable.}}
 

### 2022-07-31 SDSS Journal Paper/Paper Draft.Rmd
7.31.2022

@@
 
-The Smart and Connected Community (SCC) dashaboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
+The Smart and Connected Community (SCC) dashboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
 Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level.
@@
 \item Identifying the dimensions of rurality: population size, density, remoteness, and built-up area.
-\item Selecting measureable variables to adequately represent each dimension:
+\item Selecting measurable variables to adequately represent each dimension:
     \begin{itemize}
@@
 
-\svp{XXX Denise, can you follow up with the initial design section and explain the 2nd design and approach?
-Why did we change things?
-How did those changes impact feedback from users?
-E.g.
-wanting to show the whole range of values in the PCP vs.
-being visually overwhelmed.
-Incorporate the user feedback into the narrative.
-XXX}
+In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
+As a result, we cached map directions from OSRM for each service in our database, which drastically reduced the response time for the user.
+Our initial design did not naturally focus the user's eye on the most important parts of the dashboard; the redesign allowed for a cleaner flow from the top to the bottom.
 
-\db{In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
-As a result, we cached map directions from OSRM for each service in our database, which drastically} \svp{reduced} \db{the response time for the user.
-Our initial design \svp{did not naturally focus} the user's eye on the most important parts of the dashboard; the redesign allowed for a cleaner \svp{flow} from the top to the bottom.
-}
+In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
+The rurality index provided a better classification and the USDA sources allowed for the town to understand the impact of the closest major city due to commuting for work and shopping at larger stores not available within the town.
 
-\db{In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
-The rurality index \svp{provided a better classification} and the USDA sources allowed for the town to understand \svp{the impact of} the closest major city \svp{due to commuting for work and shopping at larger stores not available within the town}.}
-
-\svp{We also modified the parallel coordinate plots in several ways:}
+We also modified the parallel coordinate plots in several ways:
 
@@
 
-\svp{While not all of these modifications were well received in our second round of user testing, the changes did incrementally move the dashboard display towards our goal of allowing users to explore the data and engage with it.
-We continued to be surprised with how well users reacted to the parallel coordinate plots, which we initially thought might be too abstract for users unfamiliar with multivariate data displays, but the ability to compare towns across multiple dimensions and examine the similarities and differences between their approaches to different services seemed to be intuitive for users once they understood that each vertical axis was a different variable.}
+While not all of these modifications were well received in our second round of user testing, the changes did incrementally move the dashboard display towards our goal of allowing users to explore the data and engage with it.
+We continued to be surprised with how well users reacted to the parallel coordinate plots, which we initially thought might be too abstract for users unfamiliar with multivariate data displays, but the ability to compare towns across multiple dimensions and examine the similarities and differences between their approaches to different services seemed to be intuitive for users once they understood that each vertical axis was a different variable.
 
@@
 How did the philosophy work out?  
+\db{The design philosophy worked well in the ways to promote guided discovery learning for the user.
+The blocks that were created helped create a way for the user to scroll to the next section.
+This framework in the design allowed for the highlighted pieces of the town in generally.
+This design allowed the user to see the PCP after understanding how the dashboard allowed for their town to centered.
+The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.}
+
 What worked and didn't (overall)?
+\db{The idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user.
+The need to explain a visual means that the user is more likely to not use the plot effectively.
+Our hope was that the PCP would be a bit more "self-explanatory" than it was initially.
+The number of helpful hints clutters the UI and may be a distraction for a particular user.
+}
+
 People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with?
 Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis.
-What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare.
+What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare
+
+\db{Many small towns don't like to see that their information is being displayed publicly or used to be compared to another city.
+This could a huge barrier for our resarch and work.
+To make sure that this does not become a problem, we could make sure that when allowing the user to compare towns in the PCP, we make sure to de-idenitify the towns in the UI.
+For example, the town user selects similiar population size, the each town that is populated will have a collection of ten similiar towns using a label, such as "Town 1", "Town 2", etc.
+}
+\db{Along with the de-idenitifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
+This login system can then be used as a filter to setup the dashboard there forward.
+This will hopefully remove the lack of ability to be interactive.
+This will seem as if the dashbaord is developed for their needs particularly.}
+\db{Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
+Suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.}
+
 We also should provide much, much more user support for how to interact with and interpret the app and results. 
+\db{Support for the dashboard should have a help email for towns to feel supported with interpreting the app and results.
+This helpful email will be connected to our research group.
+In which the help will not only have statisticians but also city planners that can help answer questions.
+As town users are looking to make changes for their towns, it could be a QoL variable that will change the status of the town, which can be addressed with the statisticians.
+We will be able to help them with making sure that the right variables are being used in the PCPs.
+}
 
 \section{ Future Work }
+Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
+\db{In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
+In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.}
 
-Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
-Transition to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
 Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.
+
+\db{EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data.
+Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions.
+However, EDA can be used in a data discovery, such as the work displayed in this article.
+Data Discovery can help with the quality of the data, in order to find new resources.
+EDA can help with potential casuality of town related issues.
+For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploaring the data could suggest that more people are using ride share companies.
+This could as suggest that public transportation may not offer the needs for the people in the town.
+EDA will help identify those issues or even allow for the town leaders to look for answers.}
 
@@
 dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
+\db{In conclusion, }
+
 Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
@@
 As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
-This direction will be limited but may be useful foor the sake of ease on the town leaders.
+This direction will be limited but may be useful for the sake of ease on the town leaders.
 
@@
 \subsection{Feedback}
-We will continue to engage with twon leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
+We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.

### 2022-07-31 SDSS Journal Paper/Paper-Draft.tex
7.31.2022

@@
 
-The Smart and Connected Community (SCC) dashaboard data are primarily assembled from \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
+The Smart and Connected Community (SCC) dashboard data are primarily assembled from \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
 Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level.
@@
 
-{\textcolor{red}{XXX Denise, can you follow up with the initial design section and explain the 2nd design and approach?
-Why did we change things?
-How did those changes impact feedback from users?
-E.g.
-wanting to show the whole range of values in the PCP vs.
-being visually overwhelmed.
-Incorporate the user feedback into the narrative.
-XXX}}
+In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
+As a result, we cached map directions from OSRM for each service in our database, which drastically reduced the response time for the user.
+Our initial design did not naturally focus the user's eye on the most important parts of the dashboard; the redesign allowed for a cleaner flow from the top to the bottom.
 
-{\textcolor{blue}{In the initial design, we included a map for each vital service, this initially created a lag for the users' experience.
-As a result, we cached map directions from OSRM for each service in our database, which drastically}} {\textcolor{red}{reduced}} {\textcolor{blue}{the response time for the user.
-Our initial design {\textcolor{red}{did not naturally focus}} the user's eye on the most important parts of the dashboard; the redesign allowed for a cleaner {\textcolor{red}{flow}} from the top to the bottom.
-}}
+In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
+The rurality index provided a better classification and the USDA sources allowed for the town to understand the impact of the closest major city due to commuting for work and shopping at larger stores not available within the town.
 
-{\textcolor{blue}{In addition to the timing due to the map loading slowly, we added the vital statistics at the county level to allow for a more robust understanding of the town and it's surroundings.
-The rurality index {\textcolor{red}{provided a better classification}} and the USDA sources allowed for the town to understand {\textcolor{red}{the impact of}} the closest major city {\textcolor{red}{due to commuting for work and shopping at larger stores not available within the town}}.}}
-
-{\textcolor{red}{We also modified the parallel coordinate plots in several ways:}}
+We also modified the parallel coordinate plots in several ways:
 
@@
 
-{\textcolor{red}{While not all of these modifications were well received in our second round of user testing, the changes did incrementally move the dashboard display towards our goal of allowing users to explore the data and engage with it.
-We continued to be surprised with how well users reacted to the parallel coordinate plots, which we initially thought might be too abstract for users unfamiliar with multivariate data displays, but the ability to compare towns across multiple dimensions and examine the similarities and differences between their approaches to different services seemed to be intuitive for users once they understood that each vertical axis was a different variable.}}
+While not all of these modifications were well received in our second round of user testing, the changes did incrementally move the dashboard display towards our goal of allowing users to explore the data and engage with it.
+We continued to be surprised with how well users reacted to the parallel coordinate plots, which we initially thought might be too abstract for users unfamiliar with multivariate data displays, but the ability to compare towns across multiple dimensions and examine the similarities and differences between their approaches to different services seemed to be intuitive for users once they understood that each vertical axis was a different variable.
 
@@
 
-How did the philosophy work out?
+How did the philosophy work out?\\
+{\textcolor{blue}{The design philosophy worked well in the ways to promote guided discovery learning for the user.
+The blocks that were created helped create a way for the user to scroll to the next section.
+This framework in the design allowed for the highlighted pieces of the town in generally.
+This design allowed the user to see the PCP after understanding how the dashboard allowed for their town to centered.
+The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.}}
+
 What worked and didn't (overall)?
+{\textcolor{blue}{The idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user.
+The need to explain a visual means that the user is more likely to not use the plot effectively.
+Our hope was that the PCP would be a bit more "self-explanatory" than it was initially.
+The number of helpful hints clutters the UI and may be a distraction for a particular user.
+}}
+
 People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with?
 Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis.
-What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare.
+What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare
+
+{\textcolor{blue}{Many small towns don't like to see that their information is being displayed publicly or used to be compared to another city.
+This could a huge barrier for our resarch and work.
+To make sure that this does not become a problem, we could make sure that when allowing the user to compare towns in the PCP, we make sure to de-idenitify the towns in the UI.
+For example, the town user selects similiar population size, the each town that is populated will have a collection of ten similiar towns using a label, such as "Town 1", "Town 2", etc.
+}}
+{\textcolor{blue}{Along with the de-idenitifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
+This login system can then be used as a filter to setup the dashboard there forward.
+This will hopefully remove the lack of ability to be interactive.
+This will seem as if the dashbaord is developed for their needs particularly.}}
+{\textcolor{blue}{Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
+Suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.}}
+
 We also should provide much, much more user support for how to interact with and interpret the app and results.
+{\textcolor{blue}{Support for the dashboard should have a help email for towns to feel supported with interpreting the app and results.
+This helpful email will be connected to our research group.
+In which the help will not only have statisticians but also city planners that can help answer questions.
+As town users are looking to make changes for their towns, it could be a QoL variable that will change the status of the town, which can be addressed with the statisticians.
+We will be able to help them with making sure that the right variables are being used in the PCPs.
+}}
 
@@
 Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
-Transition to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
+{\textcolor{blue}{In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
+In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.}}
+
 Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.
+
+{\textcolor{blue}{EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data.
+Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions.
+However, EDA can be used in a data discovery, such as the work displayed in this article.
+Data Discovery can help with the quality of the data, in order to find new resources.
+EDA can help with potential casuality of town related issues.
+For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploaring the data could suggest that more people are using ride share companies.
+This could as suggest that public transportation may not offer the needs for the people in the town.
+EDA will help identify those issues or even allow for the town leaders to look for answers.}}
 
@@
 Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
+{\textcolor{blue}{In conclusion, }}
+
 Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
@@
 
-We will continue to engage with twon leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
+We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
 

### 2022-07-31 SDSS Journal Paper/Paper Draft.Rmd
Conclusion stuff

@@
 dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
-\db{In conclusion, }
+
+\db{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
+This direction will be limited but may be useful for the sake of ease on the town leaders.}
+
+\db{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}
+
+\db{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
+This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}
+
+\db{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention.
+This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
+This is a generalization not only for our research but along all types of dashboard design.
+We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}
 
@@
 
-\subsection{User Testing}
-In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
+\db{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm.
+Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
+At times, collaborators could feel as if there suggestions were not being heard or dismissed.
+Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often.
+As a result both groups become fustrated.
+}
+
+<!-- \subsection{User Testing} -->
+<!-- In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
 We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.
+-->
 
-Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+<!-- Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
 As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
+-->
 
-Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@lee], we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
+<!-- Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@lee], we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
 We will adapt these questions to better help the town leaders communicate clear feedback for our team to use.
-We will collect how the users are experiencing the app with these set questions:
-\begin{itemize}
-\item Is the information the dashboard presents clear? \textit{(Scope)}
-\item Is the dashboard easy to navigate? \textit{(Usability)}
-\item Is the dashboard used in a socially responsible manner? \textit{(Ethics)}
-\item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)}
-\item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)}
-\end{itemize}
+We will collect how the users are experiencing the app with these set questions: -->
+<!-- \begin{itemize} -->
+<!-- \item Is the information the dashboard presents clear? \textit{(Scope)} -->
+<!-- \item Is the dashboard easy to navigate? \textit{(Usability)} -->
+<!-- \item Is the dashboard used in a socially responsible manner? \textit{(Ethics)} -->
+<!-- \item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)} -->
+<!-- \item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)} -->
+<!-- \end{itemize} -->
 
-\subsubsection{Beta Testing}
-Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups:
-\begin{itemize}
-    \item Graphics Group Members (Users know Statistical Graphics)
-    \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics)
-    \item  Others (Member of doctoral community that has background in educational adaptivity)
-\end{itemize}
+<!-- \subsubsection{Beta Testing} -->
+<!-- Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups: -->
+<!-- \begin{itemize} -->
+<!--     \item Graphics Group Members (Users know Statistical Graphics) -->
+<!--     \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics) -->
+<!--     \item  Others (Member of doctoral community that has background in educational adaptivity) -->
+<!-- \end{itemize} -->
 
-While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes.
+<!-- While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes. -->
 
-\section{Discussion and Future Research}
+<!-- \section{Discussion and Future Research} -->
 
-\subsection{Data Collection Revisited}
+<!-- \subsection{Data Collection Revisited} -->
 
-As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources.
+<!-- As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources.
 This work empowered our time to look for other open data sources to validate our vital statistics and maps.
@@
 It could be worth looking into developing a proxy from our data collected that would be reliable for the years in between the USDA's collected years.
+-->
 
-In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns.
+<!-- In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns.
 This may be important to identify the reasons for shrinking.
@@
 This could be the most important piece of truly identifying the use of data in towns with an aging population.
+-->
 
-\subsection{Visualization Revisited}
-In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+<!-- \subsection{Visualization Revisited} -->
+<!-- In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
 As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
 This direction will be limited but may be useful for the sake of ease on the town leaders.
+-->
 
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
+<!-- In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
 This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+-->
 
-\subsection{Feedback}
-We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
+<!-- \subsection{Feedback} -->
+<!-- We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
+-->

### 2022-07-31 SDSS Journal Paper/Paper-Draft.tex
Conclusion stuff

@@
 Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
-{\textcolor{blue}{In conclusion, }}
+
+{\textcolor{blue}{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
+This direction will be limited but may be useful for the sake of ease on the town leaders.}}
+
+{\textcolor{blue}{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}}
+
+{\textcolor{blue}{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
+This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}}
+
+{\textcolor{blue}{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention.
+This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
+This is a generalization not only for our research but along all types of dashboard design.
+We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}}
 
@@
 
-\subsection{User Testing}
-
-In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices.
-We examine which aspects of dashboard design ``common knowledge'\,' hold up in practice, and assess why these principles may not hold in our use-case.
-
-Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.
-
-Using the framework from ``Creating design guidelines for building city dashboards from a user's perspectives'' \citep{lee}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions.
-We will adapt these questions to better help the town leaders communicate clear feedback for our team to use.
-We will collect how the users are experiencing the app with these set questions:
-
-\begin{itemize}
-\item Is the information the dashboard presents clear? \textit{(Scope)}
-\item Is the dashboard easy to navigate? \textit{(Usability)}
-\item Is the dashboard used in a socially responsible manner? \textit{(Ethics)}
-\item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)}
-\item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)}
-\end{itemize}
-
-\subsubsection{Beta Testing}
-
-Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups:
-
-\begin{itemize}
-    \item Graphics Group Members (Users know Statistical Graphics)
-    \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics)
-    \item  Others (Member of doctoral community that has background in educational adaptivity)
-\end{itemize}
-
-While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes.
-
-\section{Discussion and Future Research}
-
-\subsection{Data Collection Revisited}
-
-As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources.
-This work empowered our time to look for other open data sources to validate our vital statistics and maps.
-Since the clustering methods were used differently, we decided to use the USDA's version of rural area definitions.
-This allowed for a more transparent look into towns that may be similar based on there rurality index.
-Similar to the Census, the rurality metrics are county based and updated every ten years.
-This could later cause a problem for towns that may be shrinking faster than the collection of data.
-It could be worth looking into developing a proxy from our data collected that would be reliable for the years in between the USDA's collected years.
-
-In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns.
-This may be important to identify the reasons for shrinking.
-In the event, that the towns are aging, will we be able to identify how fast the younger generation is moving out of the towns.
-This could be the most important piece of truly identifying the use of data in towns with an aging population.
-
-\subsection{Visualization Revisited}
-
-In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
-This direction will be limited but may be useful for the sake of ease on the town leaders.
-
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
-
-\subsection{Feedback}
-
-We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.
+{\textcolor{blue}{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm.
+Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
+At times, collaborators could feel as if there suggestions were not being heard or dismissed.
+Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often.
+As a result both groups become fustrated.
+}}
 

### 2022-07-31 SDSS Journal Paper/Paper Draft.Rmd
adding proper citations within paper

@@
 As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
-As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.
+As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form [@fisher].
 <!-- With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{fisher}.  -->
@@
 
-The Smart and Connected Community (SCC) dashboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
+The Smart and Connected Community (SCC) dashboard data are primarily assembled from  \url{data.iowa.gov} [@iowa_gov], with some additional datasets assembled from federal and private sources.
 Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level.
@@
 Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
-Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
-The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
-\db{[@petersCommunityResiliencyDeclining2019] -- Change the reference}
+Data collected from ELSI [@elsi] from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
+The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education [@zarecor2021rural].
 <!-- \svp{Cite Dave's papers about the ISTP QOL metrics, including education}
@@
 This design is based on research into visualization sensemaking [@lee], in that we allow users to explore outward from the familiar to the unknown.
-\db{The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point.
-OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.}
+The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point.
+OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.
 
@@
 
-\svp{Users' towns} \db{Town Users' ??} are at the center of our application, and comparisons to other, similar towns are secondary.
+Users' towns are at the center of our application, and comparisons to other, similar towns are secondary.
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.

### 2022-07-31 SDSS Journal Paper/Paper-Draft.tex
adding proper citations within paper

@@
 
-The Smart and Connected Community (SCC) dashboard data are primarily assembled from \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources.
+The Smart and Connected Community (SCC) dashboard data are primarily assembled from \url{data.iowa.gov} \citep{iowa_gov}, with some additional datasets assembled from federal and private sources.
 Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level.
@@
 Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
-Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
-The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education.
-{\textcolor{blue}{[@petersCommunityResiliencyDeclining2019] -- Change the reference}}
+Data collected from ELSI \citep{elsi} from \url{https://nces.ed.gov} were used to show the distance to any private or public school.
+The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education \citep{zarecor2021rural}.
 
@@
 This design is based on research into visualization sensemaking \citep{lee}, in that we allow users to explore outward from the familiar to the unknown.
-{\textcolor{blue}{The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point.
-OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.}}
+The map visuals were built using Open Source Routing Machine (OSRM) route functions \citep{luxen-vetter-2011} in R \citep{r} to amplify the accuracy of the distances from necessary services in town-centric point.
+OSRM allows for finding the ``As the Crow Flies'' distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.
 
@@
 
-{\textcolor{red}{Users' towns}} {\textcolor{blue}{Town Users' ??}} are at the center of our application, and comparisons to other, similar towns are secondary.
+Users' towns are at the center of our application, and comparisons to other, similar towns are secondary.
 As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.

### 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd
Edits to remove the color edits

@@
 \section{ Discussion}
-How did the philosophy work out?  
-\db{The design philosophy worked well in the ways to promote guided discovery learning for the user.
-The blocks that were created helped create a way for the user to scroll to the next section.
-This framework in the design allowed for the highlighted pieces of the town in generally.
-This design allowed the user to see the PCP after understanding how the dashboard allowed for their town to centered.
-The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.}
+Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary.
+This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally.
+Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status.
+The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.
 
-What worked and didn't (overall)?
-\db{The idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user.
+While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
 The need to explain a visual means that the user is more likely to not use the plot effectively.
 Our hope was that the PCP would be a bit more "self-explanatory" than it was initially.
-The number of helpful hints clutters the UI and may be a distraction for a particular user.
-}
+The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.
 
-People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with?
-Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis.
-What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare
+Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town.
+As a result, small towns don't like to see that their information being displayed publicly or compared to another city.
+This could a huge barrier for our research and work.
+To make sure that this does not become a problem, in the future work making sure that when allowing the user to compare towns in a visual graphic with a de-identified towns in the UI.
+For example, the town user selects similar population size, the each town that is populated will have a collection of ten similar towns using a label, such as "Town 1", "Town 2", etc.
+Along with the de-identifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
+This login system can then be used as a filter to setup the dashboard there forward.
+This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.
 
-\db{Many small towns don't like to see that their information is being displayed publicly or used to be compared to another city.
-This could a huge barrier for our resarch and work.
-To make sure that this does not become a problem, we could make sure that when allowing the user to compare towns in the PCP, we make sure to de-idenitify the towns in the UI.
-For example, the town user selects similiar population size, the each town that is populated will have a collection of ten similiar towns using a label, such as "Town 1", "Town 2", etc.
-}
-\db{Along with the de-idenitifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
-This login system can then be used as a filter to setup the dashboard there forward.
-This will hopefully remove the lack of ability to be interactive.
-This will seem as if the dashbaord is developed for their needs particularly.}
-\db{Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
-Suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.}
+Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
+By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.
 
-We also should provide much, much more user support for how to interact with and interpret the app and results. 
-\db{Support for the dashboard should have a help email for towns to feel supported with interpreting the app and results.
-This helpful email will be connected to our research group.
-In which the help will not only have statisticians but also city planners that can help answer questions.
-As town users are looking to make changes for their towns, it could be a QoL variable that will change the status of the town, which can be addressed with the statisticians.
-We will be able to help them with making sure that the right variables are being used in the PCPs.
-}
+Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear.
+In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions.
+Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.
 
 \section{ Future Work }
-Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
-\db{In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
-In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.}
+In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
+In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
 
-Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.
+In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+This direction will be limited but may be useful for the sake of ease on the town leaders.
+In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot.
+This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
 
-\db{EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
 While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data.
@@
 Data Discovery can help with the quality of the data, in order to find new resources.
-EDA can help with potential casuality of town related issues.
-For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploaring the data could suggest that more people are using ride share companies.
+EDA can help with potential causality of town related issues.
+For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies.
 This could as suggest that public transportation may not offer the needs for the people in the town.
-EDA will help identify those issues or even allow for the town leaders to look for answers.}
+EDA will help identify those issues or even allow for the town leaders to look for answers.
 
 \section {Conclusions}
-Specific takeaways - visualization research (broad) vs.
-dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
 
-\db{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
 As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
-This direction will be limited but may be useful for the sake of ease on the town leaders.}
+This direction will be limited but may be useful for the sake of ease on the town leaders.
 
-\db{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}
+Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
 
-\db{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}
+In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
+This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
 
-\db{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention.
+With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
 This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
 This is a generalization not only for our research but along all types of dashboard design.
-We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}
+We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.
 
-Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
-Test early, test often...
-
-\db{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm.
+When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm.
 Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
 At times, collaborators could feel as if there suggestions were not being heard or dismissed.
-Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often.
-As a result both groups become fustrated.
-}
+Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often.
+As a result both groups become frustrated.
 

### 2022-08-01 SDSS Journal Paper/Paper-Draft.tex
Edits to remove the color edits

@@
 
-How did the philosophy work out?\\
-{\textcolor{blue}{The design philosophy worked well in the ways to promote guided discovery learning for the user.
-The blocks that were created helped create a way for the user to scroll to the next section.
-This framework in the design allowed for the highlighted pieces of the town in generally.
-This design allowed the user to see the PCP after understanding how the dashboard allowed for their town to centered.
-The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.}}
+Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary.
+This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally.
+Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status.
+The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.
 
-What worked and didn't (overall)?
-{\textcolor{blue}{The idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user.
+While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
 The need to explain a visual means that the user is more likely to not use the plot effectively.
-Our hope was that the PCP would be a bit more "self-explanatory" than it was initially.
-The number of helpful hints clutters the UI and may be a distraction for a particular user.
-}}
+Our hope was that the PCP would be a bit more ``self-explanatory'' than it was initially.
+The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.
 
-People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with?
-Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis.
-What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare
+Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town.
+As a result, small towns don't like to see that their information being displayed publicly or compared to another city.
+This could a huge barrier for our research and work.
+To make sure that this does not become a problem, in the future work making sure that when allowing the user to compare towns in a visual graphic with a de-identified towns in the UI.
+For example, the town user selects similar population size, the each town that is populated will have a collection of ten similar towns using a label, such as ``Town 1'', ``Town 2'', etc.
+Along with the de-identifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
+This login system can then be used as a filter to setup the dashboard there forward.
+This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.
 
-{\textcolor{blue}{Many small towns don't like to see that their information is being displayed publicly or used to be compared to another city.
-This could a huge barrier for our resarch and work.
-To make sure that this does not become a problem, we could make sure that when allowing the user to compare towns in the PCP, we make sure to de-idenitify the towns in the UI.
-For example, the town user selects similiar population size, the each town that is populated will have a collection of ten similiar towns using a label, such as "Town 1", "Town 2", etc.
-}}
-{\textcolor{blue}{Along with the de-idenitifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
-This login system can then be used as a filter to setup the dashboard there forward.
-This will hopefully remove the lack of ability to be interactive.
-This will seem as if the dashbaord is developed for their needs particularly.}}
-{\textcolor{blue}{Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
-Suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.}}
+Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
+By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.
 
-We also should provide much, much more user support for how to interact with and interpret the app and results.
-{\textcolor{blue}{Support for the dashboard should have a help email for towns to feel supported with interpreting the app and results.
-This helpful email will be connected to our research group.
-In which the help will not only have statisticians but also city planners that can help answer questions.
-As town users are looking to make changes for their towns, it could be a QoL variable that will change the status of the town, which can be addressed with the statisticians.
-We will be able to help them with making sure that the right variables are being used in the PCPs.
-}}
+Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear.
+In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions.
+Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.
 
@@
 
-Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
-{\textcolor{blue}{In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
-In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.}}
+In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
+In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
 
-Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.
+In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+This direction will be limited but may be useful for the sake of ease on the town leaders.
+In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot.
+This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
 
-{\textcolor{blue}{EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
 While working with a variety of disciplines because EDA can be hard to understand the use cases of exploration of data.
@@
 Data Discovery can help with the quality of the data, in order to find new resources.
-EDA can help with potential casuality of town related issues.
-For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploaring the data could suggest that more people are using ride share companies.
+EDA can help with potential causality of town related issues.
+For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies.
 This could as suggest that public transportation may not offer the needs for the people in the town.
-EDA will help identify those issues or even allow for the town leaders to look for answers.}}
+EDA will help identify those issues or even allow for the town leaders to look for answers.
 
@@
 
-Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
+In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
+This direction will be limited but may be useful for the sake of ease on the town leaders.
 
-{\textcolor{blue}{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
-This direction will be limited but may be useful for the sake of ease on the town leaders.}}
+Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.
 
-{\textcolor{blue}{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}}
+In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
+This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
 
-{\textcolor{blue}{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}}
-
-{\textcolor{blue}{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention.
+With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
 This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
 This is a generalization not only for our research but along all types of dashboard design.
-We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}}
+We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.
 
-Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
-Test early, test often\ldots{}
-
-{\textcolor{blue}{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm.
+When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm.
 Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
 At times, collaborators could feel as if there suggestions were not being heard or dismissed.
-Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often.
-As a result both groups become fustrated.
-}}
+Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often.
+As a result both groups become frustrated.
 

### 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd
more edits from reread

@@
 
-In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-This direction will be limited but may be useful for the sake of ease on the town leaders.
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+Parallel Coordinate Plots (PCPs) allowed the user to look at multidimensional variables, while the Guided Discovery Learning was used to make sure that users were informed of the information that was presented, it did not provide the town leaders any more ease in understanding our statistical graphic.
+Our new proposed direction will be limited but may be useful for the sake of ease on the town leaders and allow for the exploration into other pieces of data, by using a flower plot method.
+This method may not represent that data in the way that could be misleading of the whole picture of the town, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
 
-EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
-While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data.
-Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions.
+While working with a variety of disciplines, EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+Because EDA can be hard to understand when the idea of EDA is only to prepare for modeling.
 However, EDA can be used in a data discovery, such as the work displayed in this article.
 Data Discovery can help with the quality of the data, in order to find new resources.
-EDA can help with potential causality of town related issues.
-For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies.
-This could as suggest that public transportation may not offer the needs for the people in the town.
-EDA will help identify those issues or even allow for the town leaders to look for answers.
 
-\section {Conclusions}
+\section{Conclusions}
 
-In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
+In conclusion, using a less known statistical graphic, PCPs, to the novice users was used to make sure that the users were informed of the multidimensional aspect of their information to help data driven decision making.
+As a result, we decided to look for easier ways to display data, such as boxplots, barplots, or flower plots.
 This direction will be limited but may be useful for the sake of ease on the town leaders.
@@
 With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
-This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
-This is a generalization not only for our research but along all types of dashboard design.
-We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.
+This lack of clarity outside of dashboard pitfalls by Stephen Few could lend the dashboard to miss key components that are useful.
+This is a generalization not only for our research but across all types of dashboard design.
+We are encouraged to minimize an interface that will not be overwhelming the user, but in what order should these components should be used can be a struggle for our research.
 
-When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm.
-Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
+When showing our applications versions to internal research members, it was difficult to collective incorporate necessary components and empower collaborators to help brainstorm.
+Once a design was presented, there were polarizing responses from no feedback to a dramatic change of the dashboard layout.
 At times, collaborators could feel as if there suggestions were not being heard or dismissed.
@@
 As a result both groups become frustrated.
+As a result our finalizing motto will be to "test often and test again," this will help ease the feeling of exclusivity with our collaborators.
 

### 2022-08-01 SDSS Journal Paper/Paper-Draft.tex
more edits from reread

@@
 
-In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-This direction will be limited but may be useful for the sake of ease on the town leaders.
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+Parallel Coordinate Plots (PCPs) allowed the user to look at multidimensional variables, while the Guided Discovery Learning was used to make sure that users were informed of the information that was presented, it did not provide the town leaders any more ease in understanding our statistical graphic.
+Our new proposed direction will be limited but may be useful for the sake of ease on the town leaders and allow for the exploration into other pieces of data, by using a flower plot method.
+This method may not represent that data in the way that could be misleading of the whole picture of the town, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
 
-EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
-While working with a variety of disciplines because EDA can be hard to understand the use cases of exploration of data.
-Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions.
+While working with a variety of disciplines, EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+Because EDA can be hard to understand when the idea of EDA is only to prepare for modeling.
 However, EDA can be used in a data discovery, such as the work displayed in this article.
 Data Discovery can help with the quality of the data, in order to find new resources.
-EDA can help with potential causality of town related issues.
-For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies.
-This could as suggest that public transportation may not offer the needs for the people in the town.
-EDA will help identify those issues or even allow for the town leaders to look for answers.
 
-\section {Conclusions}
+\section{Conclusions}
 
-In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
+In conclusion, using a less known statistical graphic, PCPs, to the novice users was used to make sure that the users were informed of the multidimensional aspect of their information to help data driven decision making.
+As a result, we decided to look for easier ways to display data, such as boxplots, barplots, or flower plots.
 This direction will be limited but may be useful for the sake of ease on the town leaders.
@@
 As a result both groups become frustrated.
+As a result our finalizing motto will be to ``test often and test again,'' this will help ease the feeling of exclusivity with our collaborators.
 

### 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd
My modifications

--- 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd
@@
 Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary.
-This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally.
-Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status.
-The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.
+This approach created a way for the user to see their town information at the top of the page and to explore the PCP after reviewing their own town's essential statistics.
+The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.
+\svp{While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start.
+We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.}
 
-While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
+One factor that we have briefly considered and have seen hints of in our user feedback is that towns may not want to be compared negatively with other towns.
+While users have very definite ideas about which towns they would like to compare to, we can always mask the town names and move back to comparisons based on town size and other factors (for instance, whether or not a town is the county seat is a factor that is important outside of population).
+Using this approach, we would label each town as "Town 1", "Town 2", and so on, which would eliminate some of the fears about negative comparisons, but would also remove some of the novelty of the data dashboard for our users and would prevent users from drawing on their own outside knowledge about each of the comparison towns.
+
+We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it.
+We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.
+<!-- While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
 The need to explain a visual means that the user is more likely to not use the plot effectively.
@@
 The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.
+-->
 
-Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town.
+<!-- Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town.
 As a result, small towns don't like to see that their information being displayed publicly or compared to another city.
@@
 This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.
+-->
 
-Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
+<!-- Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
 By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.
+-->
 
-Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear.
+<!-- Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear.
 In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions.
 Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.
+-->
 
 \section{ Future Work }
-In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
+One avenue we will explore in future iterations of the dashboard is to incorporate other dashboards generated by different groups within this project.
+This will create a wider field of information to explore: for instance, some of the additional work will focus on the 99 towns featured in the Iowa Small Town Poll; this will allow us to showcase survey-based measures of quality of life alongside the more objective measurements assembled in the dataset discussed in this paper.
+While at least one tab of this omni-dashboard will still focus on wider EDA and discovery, we hope to incorporate other information as well to provide a more well-rounded data display encompassing most of the facets of this complex project.
+
+We are also mindful of a distinction between "eye candy" and purpose-driven data visualization.
+While we have typically focused on the latter, there is certainly a place in our dashboard for the former as well.
+"Eye candy" visualization is intended to draw the viewer in and motivate them to explore; while these visualizations may not be particularly effective at communicating quantitative information, if they motivate the user to engage with the rest of the dashboard, they still serve a purpose.
+It is with this mindset that we intend to explore the use of flower plots - the artistic opportunities combined with the display of quantitative information (even in a form that isn't optimal for quantitative comparisons) may be useful to engage viewers before transitioning to more useful data visualizations intended to provide accurate quantitative comparisons.
+
+<!-- In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
 In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
+-->
 
-In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+<!-- In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
 This direction will be limited but may be useful for the sake of ease on the town leaders.
@@
 This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+-->
 
 EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
-While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data.
-Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions.
-However, EDA can be used in a data discovery, such as the work displayed in this article.
-Data Discovery can help with the quality of the data, in order to find new resources.
-EDA can help with potential causality of town related issues.
-For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies.
-This could as suggest that public transportation may not offer the needs for the people in the town.
-EDA will help identify those issues or even allow for the town leaders to look for answers.
+One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept.
+One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project.
+Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard.
+While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.
 
@@
 
-In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
+In this paper, we have documented the process of designing a dashboard for exploration and visualization of a large and complex data set assembled from many different sources.
+Our primary audience was leaders of small towns in Iowa, with a secondary audience of researchers in fields other than statistics collaborating on this project with us.
+Through the process of revising our dashboard, we found that the idea of guided discovery learning as implemented in our first version did not work as well as we had anticipated.
+It was more important to focus on allowing users to explore their questions about the dataset by facilitating user-driven comparisons and exploration, rather than attempting to anticipate user desires by providing comparison towns.
+In addition, we found that it would be more effective to draw users in with novel visual displays, as these seemed to attract more interest than providing known facts and an opportunity to explore outwards from an initial area of familiarity.
+
+While it is hard to apply the findings from one fairly specific visualization project more widely, there is a lack of resources in this area that provide both design philosophies and actual analysis of user feedback in a qualitative sense.
+We have attempted to address this dearth of information by providing the design strategies, user feedback, and our planned and executed modifications, in the hopes that others facing the daunting challenge of designing a dashboard for EDA may learn something from our experiences.
+
+<!-- In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
 As a result, we decided to look for easier ways to display data, such as boxplots and barplots.
 This direction will be limited but may be useful for the sake of ease on the town leaders.
+-->
 
-Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
+<!-- Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
 As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.
+-->
 
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
+<!-- In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
 This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+-->
 
-With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
+<!-- With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
 This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
@@
 We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.
+-->
 
-When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm.
+<!-- When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm.
 Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
@@
 As a result both groups become frustrated.
+-->
 

### 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd

@@
 
-<!-- In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding.
-This direction will be limited but may be useful for the sake of ease on the town leaders.
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+<!-- Parallel Coordinate Plots (PCPs) allowed the user to look at multidimensional variables, while the Guided Discovery Learning was used to make sure that users were informed of the information that was presented, it did not provide the town leaders any more ease in understanding our statistical graphic.
+Our new proposed direction will be limited but may be useful for the sake of ease on the town leaders and allow for the exploration into other pieces of data, by using a flower plot method.
+This method may not represent that data in the way that could be misleading of the whole picture of the town, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+-->
+
+<!-- While working with a variety of disciplines, EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+Because EDA can be hard to understand when the idea of EDA is only to prepare for modeling.
+However, EDA can be used in a data discovery, such as the work displayed in this article.
+Data Discovery can help with the quality of the data, in order to find new resources.
 -->
@@
 
-\section {Conclusions}
+\section{Conclusions}
 
@@
 We have attempted to address this dearth of information by providing the design strategies, user feedback, and our planned and executed modifications, in the hopes that others facing the daunting challenge of designing a dashboard for EDA may learn something from our experiences.
+
+<!-- In conclusion, using a less known statistical graphic, PCPs, to the novice users was used to make sure that the users were informed of the multidimensional aspect of their information to help data driven decision making.
+As a result, we decided to look for easier ways to display data, such as boxplots, barplots, or flower plots.
+This direction will be limited but may be useful for the sake of ease on the town leaders.
+-->
 
@@
 <!-- With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
-This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful.
-This is a generalization not only for our research but along all types of dashboard design.
-We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.
+This lack of clarity outside of dashboard pitfalls by Stephen Few could lend the dashboard to miss key components that are useful.
+This is a generalization not only for our research but across all types of dashboard design.
+We are encouraged to minimize an interface that will not be overwhelming the user, but in what order should these components should be used can be a struggle for our research.
 -->
 
-<!-- When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm.
-Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout.
+<!-- When showing our applications versions to internal research members, it was difficult to collective incorporate necessary components and empower collaborators to help brainstorm.
+Once a design was presented, there were polarizing responses from no feedback to a dramatic change of the dashboard layout.
 At times, collaborators could feel as if there suggestions were not being heard or dismissed.
@@
 As a result both groups become frustrated.
+As a result our finalizing motto will be to "test often and test again," this will help ease the feeling of exclusivity with our collaborators.
 -->

### 2022-08-01 SDSS Journal Paper/Paper-Draft.tex

@@
 Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary.
-This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally.
-Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status.
-The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.
+This approach created a way for the user to see their town information at the top of the page and to explore the PCP after reviewing their own town's essential statistics.
+The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.
+{\textcolor{red}{While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start.
+We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.}}
 
-While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
-The need to explain a visual means that the user is more likely to not use the plot effectively.
-Our hope was that the PCP would be a bit more ``self-explanatory'' than it was initially.
-The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.
+One factor that we have briefly considered and have seen hints of in our user feedback is that towns may not want to be compared negatively with other towns.
+While users have very definite ideas about which towns they would like to compare to, we can always mask the town names and move back to comparisons based on town size and other factors (for instance, whether or not a town is the county seat is a factor that is important outside of population).
+Using this approach, we would label each town as ``Town 1'', ``Town 2'', and so on, which would eliminate some of the fears about negative comparisons, but would also remove some of the novelty of the data dashboard for our users and would prevent users from drawing on their own outside knowledge about each of the comparison towns.
 
-Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town.
-As a result, small towns don't like to see that their information being displayed publicly or compared to another city.
-This could a huge barrier for our research and work.
-To make sure that this does not become a problem, in the future work making sure that when allowing the user to compare towns in a visual graphic with a de-identified towns in the UI.
-For example, the town user selects similar population size, the each town that is populated will have a collection of ten similar towns using a label, such as ``Town 1'', ``Town 2'', etc.
-Along with the de-identifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected.
-This login system can then be used as a filter to setup the dashboard there forward.
-This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.
-
-Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit.
-By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.
-
-Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear.
-In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions.
-Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.
+We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it.
+We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.
 
@@
 
-In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.).
-In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.
+One avenue we will explore in future iterations of the dashboard is to incorporate other dashboards generated by different groups within this project.
+This will create a wider field of information to explore: for instance, some of the additional work will focus on the 99 towns featured in the Iowa Small Town Poll; this will allow us to showcase survey-based measures of quality of life alongside the more objective measurements assembled in the dataset discussed in this paper.
+While at least one tab of this omni-dashboard will still focus on wider EDA and discovery, we hope to incorporate other information as well to provide a more well-rounded data display encompassing most of the facets of this complex project.
 
-Parallel Coordinate Plots (PCPs) allowed the user to look at multidimensional variables, while the Guided Discovery Learning was used to make sure that users were informed of the information that was presented, it did not provide the town leaders any more ease in understanding our statistical graphic.
-Our new proposed direction will be limited but may be useful for the sake of ease on the town leaders and allow for the exploration into other pieces of data, by using a flower plot method.
-This method may not represent that data in the way that could be misleading of the whole picture of the town, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
+We are also mindful of a distinction between ``eye candy'' and purpose-driven data visualization.
+While we have typically focused on the latter, there is certainly a place in our dashboard for the former as well.
+``Eye candy'' visualization is intended to draw the viewer in and motivate them to explore; while these visualizations may not be particularly effective at communicating quantitative information, if they motivate the user to engage with the rest of the dashboard, they still serve a purpose.
+It is with this mindset that we intend to explore the use of flower plots - the artistic opportunities combined with the display of quantitative information (even in a form that isn't optimal for quantitative comparisons) may be useful to engage viewers before transitioning to more useful data visualizations intended to provide accurate quantitative comparisons.
 
-While working with a variety of disciplines, EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
-Because EDA can be hard to understand when the idea of EDA is only to prepare for modeling.
-However, EDA can be used in a data discovery, such as the work displayed in this article.
-Data Discovery can help with the quality of the data, in order to find new resources.
+EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept.
+One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project.
+Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard.
+While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.
 
@@
 
-In conclusion, using a less known statistical graphic, PCPs, to the novice users was used to make sure that the users were informed of the multidimensional aspect of their information to help data driven decision making.
-As a result, we decided to look for easier ways to display data, such as boxplots, barplots, or flower plots.
-This direction will be limited but may be useful for the sake of ease on the town leaders.
+In this paper, we have documented the process of designing a dashboard for exploration and visualization of a large and complex data set assembled from many different sources.
+Our primary audience was leaders of small towns in Iowa, with a secondary audience of researchers in fields other than statistics collaborating on this project with us.
+Through the process of revising our dashboard, we found that the idea of guided discovery learning as implemented in our first version did not work as well as we had anticipated.
+It was more important to focus on allowing users to explore their questions about the dataset by facilitating user-driven comparisons and exploration, rather than attempting to anticipate user desires by providing comparison towns.
+In addition, we found that it would be more effective to draw users in with novel visual displays, as these seemed to attract more interest than providing known facts and an opportunity to explore outwards from an initial area of familiarity.
 
-Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions.
-As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.
-
-In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot.
-This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.
-
-With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention.
-This lack of clarity outside of dashboard pitfalls by Stephen Few could lend the dashboard to miss key components that are useful.
-This is a generalization not only for our research but across all types of dashboard design.
-We are encouraged to minimize an interface that will not be overwhelming the user, but in what order should these components should be used can be a struggle for our research.
-
-When showing our applications versions to internal research members, it was difficult to collective incorporate necessary components and empower collaborators to help brainstorm.
-Once a design was presented, there were polarizing responses from no feedback to a dramatic change of the dashboard layout.
-At times, collaborators could feel as if there suggestions were not being heard or dismissed.
-Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often.
-As a result both groups become frustrated.
-As a result our finalizing motto will be to ``test often and test again,'' this will help ease the feeling of exclusivity with our collaborators.
+While it is hard to apply the findings from one fairly specific visualization project more widely, there is a lack of resources in this area that provide both design philosophies and actual analysis of user feedback in a qualitative sense.
+We have attempted to address this dearth of information by providing the design strategies, user feedback, and our planned and executed modifications, in the hopes that others facing the daunting challenge of designing a dashboard for EDA may learn something from our experiences.
 

### 2022-08-02 SDSS Journal Paper/Paper Draft.Rmd
Updates for final submission

@@
 The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.
-\svp{While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start.
-We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.}
+
+While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start.
+We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.
 

### 2022-09-17 SDSS Journal Paper/Paper Draft.Rmd
Updates from revisions

@@
 preamble: >
-  \usepackage{amsfonts,amsmath,amssymb,amsthm}
+  \usepackage{amsfonts,amsmath,amssymb,amsthm,ulem}
   \usepackage{booktabs}
   \usepackage{lipsum}
-  \newcommand{\db}[1]{{\textcolor{blue}{#1}}}
+  \newcommand{\db}[1]{{\textcolor{green}{#1}}}
   \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
@@
 Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
-In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
+In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
 Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
@@
 
-# Abstract
+\db{\sout{Abstract}}
 
-Many small and rural places are shrinking.
+\db{\sout{Many small and rural places are shrinking.
 Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
@@
 Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
+}}
 
@@
 Data collected from the Index of Relative Rurality (IRR) [@Rural_classification]  were used in the SCC dashboard to help classify the towns.
-The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality.
+The \db{\sout{Index of Relative Rurality (}}IRR\db{\sout{)}} is a continuous, threshold-free, and unit-free measure of rurality.
 It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural).
@@
 At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
-This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
+This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework \db{that} leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
 
@@
 Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp].
-In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
+In this paper, we use the \texttt{ggpcp} package \db{in R [@Rlanguage]}, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
 
@@
 Taken together, the app is a single page; the initial "solid ground" which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
-In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.
+In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's \db{Quality of Life (QoL)} and financial metrics.
 This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.
@@
 
-- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with quality of life.
+- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with \db{QoL, grouped into four domains: physical, psychological, social, and environmental}.
 However, there were still too many variables for users to successfully parse.
-We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by quality of life factor [@petersCommunityResiliencyDeclining2019].
+We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by \db{\sout{quality of life} QoL} factor [@petersCommunityResiliencyDeclining2019].
 - Originally, parallel coordinate bands were scaled based on the selected comparison towns.
@@
 
-We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it.
-We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.
+We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the \db{applet} and teach towns how to use it.
+We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining \db{\sout{quality of life} QoL} amid shrinking populations.
 <!-- While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
@@
 
-EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+EDA can be \db{\sout{a}} difficult for a variety of groups of people, novice users and experienced researchers.
 One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept.

### 2022-09-17 SDSS Journal Paper/Paper-Draft.tex
Updates from revisions

@@
 
-\usepackage{amsfonts,amsmath,amssymb,amsthm} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{blue}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
+\usepackage{amsfonts,amsmath,amssymb,amsthm,ulem} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{green}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
 
@@
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+{\textcolor{green}{\sout{Abstract}}}
 
-Many small and rural places are shrinking.
+{\textcolor{green}{\sout{Many small and rural places are shrinking.
 Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
@@
 Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
+}}}
 
@@
 Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns.
-The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality.
+The {\textcolor{green}{\sout{Index of Relative Rurality (}}}IRR{\textcolor{green}{\sout{)}}} is a continuous, threshold-free, and unit-free measure of rurality.
 It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural).
@@
 At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
-This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
+This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework {\textcolor{green}{that}} leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
 
@@
 Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}.
-In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
+In this paper, we use the \texttt{ggpcp} package {\textcolor{green}{in R [@Rlanguage]}}, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
 
@@
 Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
-In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics.
+In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's {\textcolor{green}{Quality of Life (QoL)}} and financial metrics.
 This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.
@@
 \item
-  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with quality of life.
+  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with {\textcolor{green}{QoL, grouped into four domains: physical, psychological, social, and environmental}}.
 However, there were still too many variables for users to successfully parse.
-We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by quality of life factor \citep{petersCommunityResiliencyDeclining2019}.
+We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by {\textcolor{green}{\sout{quality of life} QoL}} factor \citep{petersCommunityResiliencyDeclining2019}.
 \item
@@
 The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.
-{\textcolor{red}{While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start.
-We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.}}
+
+While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start.
+We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.
 
@@
 
-We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it.
-We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.
+We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the {\textcolor{green}{applet}} and teach towns how to use it.
+We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining {\textcolor{green}{\sout{quality of life} QoL}} amid shrinking populations.
 
@@
 
-EDA can be a difficult for a variety of groups of people, novice users and experienced researchers.
+EDA can be {\textcolor{green}{\sout{a}}} difficult for a variety of groups of people, novice users and experienced researchers.
 One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept.

### 2022-09-20 SDSS Journal Paper/Paper Draft.Rmd
accepting changes

@@
   \newcommand{\db}[1]{{\textcolor{green}{#1}}}
-  \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
+  \newcommand{\svp}[1]{{\textcolor{blue}{#1}}}
 
@@
 ---
-
-\db{\sout{Abstract}}
-
-\db{\sout{Many small and rural places are shrinking.
-Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
-In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
-Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
-}}
 
@@
 Data collected from the Index of Relative Rurality (IRR) [@Rural_classification]  were used in the SCC dashboard to help classify the towns.
-The \db{\sout{Index of Relative Rurality (}}IRR\db{\sout{)}} is a continuous, threshold-free, and unit-free measure of rurality.
+The IRR is a continuous, threshold-free, and unit-free measure of rurality.
 It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural).
@@
 At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
-This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework \db{that} leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
+This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework that leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].
 
@@
 Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp].
-In this paper, we use the \texttt{ggpcp} package \db{in R [@Rlanguage]}, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
+In this paper, we use the \texttt{ggpcp} package in R [@Rlanguage,@ggpcp], which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
 
@@
 Taken together, the app is a single page; the initial "solid ground" which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
-In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's \db{Quality of Life (QoL)} and financial metrics.
+In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's Quality of Life (QoL) and financial metrics.
 This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.
@@
 
-- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with \db{QoL, grouped into four domains: physical, psychological, social, and environmental}.
+- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with QoL, grouped into four domains: physical, psychological, social, and environmental.
 However, there were still too many variables for users to successfully parse.
-We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by \db{\sout{quality of life} QoL} factor [@petersCommunityResiliencyDeclining2019].
+We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by QoL factor [@petersCommunityResiliencyDeclining2019].
 - Originally, parallel coordinate bands were scaled based on the selected comparison towns.
@@
 
-We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the \db{applet} and teach towns how to use it.
-We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining \db{\sout{quality of life} QoL} amid shrinking populations.
+We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the applet and teach towns how to use it.
+We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining QoL amid shrinking populations.
 <!-- While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP.
@@
 
-EDA can be \db{\sout{a}} difficult for a variety of groups of people, novice users and experienced researchers.
+EDA can be difficult for a variety of groups of people, novice users and experienced researchers.
 One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept.

### 2022-09-20 SDSS Journal Paper/Paper Draft.Rmd
one last \db comment

@@
 Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
-In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
+In our paper, we will use Iowa data to  explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
 Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.

### 2022-09-20 SDSS Journal Paper/Paper-Draft.tex
one last \db comment

@@
 
-\usepackage{amsfonts,amsmath,amssymb,amsthm,ulem} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{green}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}
+\usepackage{amsfonts,amsmath,amssymb,amsthm,ulem} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{green}{#1}}} \newcommand{\svp}[1]{{\textcolor{blue}{#1}}}
 
@@
 Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
-In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
+In our paper, we will use Iowa data to explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
 Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
@@
 \end{frontmatter}
-
-{\textcolor{green}{\sout{Abstract}}}
-
-{\textcolor{green}{\sout{Many small and rural places are shrinking.
-Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools.
-In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions.
-Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
-}}}
 
@@
 Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns.
-The {\textcolor{green}{\sout{Index of Relative Rurality (}}}IRR{\textcolor{green}{\sout{)}}} is a continuous, threshold-free, and unit-free measure of rurality.
+The IRR is a continuous, threshold-free, and unit-free measure of rurality.
 It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural).
@@
 At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
-This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework {\textcolor{green}{that}} leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
+This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework that leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.
 
@@
 Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}.
-In this paper, we use the \texttt{ggpcp} package {\textcolor{green}{in R [@Rlanguage]}}, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
+In this paper, we use the \texttt{ggpcp} package in R \citep[\citet{ggpcp}]{Rlanguage}, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.
 
@@
 Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals.
-In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's {\textcolor{green}{Quality of Life (QoL)}} and financial metrics.
+In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's Quality of Life (QoL) and financial metrics.
 This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.
@@
 \item
-  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with {\textcolor{green}{QoL, grouped into four domains: physical, psychological, social, and environmental}}.
+  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with QoL, grouped into four domains: physical, psychological, social, and environmental.
 However, there were still too many variables for users to successfully parse.
-We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by {\textcolor{green}{\sout{quality of life} QoL}} factor \citep{petersCommunityResiliencyDeclining2019}.
+We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by QoL factor \citep{petersCommunityResiliencyDeclining2019}.
 \item
@@
 
-We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the {\textcolor{green}{applet}} and teach towns how to use it.
-We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining {\textcolor{green}{\sout{quality of life} QoL}} amid shrinking populations.
+We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the applet and teach towns how to use it.
+We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining QoL amid shrinking populations.
 
@@
 
-EDA can be {\textcolor{green}{\sout{a}}} difficult for a variety of groups of people, novice users and experienced researchers.
+EDA can be difficult for a variety of groups of people, novice users and experienced researchers.
 One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept.

### 2022-09-20 SDSS Journal Paper/Paper Draft.Rmd
Fixing point #8 (no more specificity but at least avoiding the "small subset" language)

@@
 The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
-We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
+\svp{The user selects a single town of interest and we add summary variables reflecting QoL metrics such as distance to local schools and city budget.
+Then, the user can} in accordance with their interest.
 This avoids some of the pitfalls of dashboard design that can easily lead to user overload [@few].

### 2022-09-28 SDSS Journal Paper/Paper Draft.Rmd
Added the citation for the small definitions

@@
 This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades.
-As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations [@scc].
+As small rural towns, defined in the Iowa Small Town Poll as towns with between 500 and 10,000 people in the 1990 census [@peters2018using], do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations [@scc].
 We hope to allow towns to explore their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.

### 2022-09-28 SDSS Journal Paper/Paper-Draft.tex
Added the citation for the small definitions

@@
 This process, \emph{smart shrinkage}, is important for rural areas who have experienced shrinking populations for decades.
-As small rural towns do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}.
+As small rural towns, defined in the Iowa Small Town Poll as towns with between 500 and 10,000 people in the 1990 census \citep{peters2018using}, do not have access to data scientists or even the ability to easily leverage data collected locally to support decisions, our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations \citep{scc}.
 We hope to allow towns to explore their own data and compare to other similar towns, centering decision-making on data in the context of small-town Iowa life.
@@
 The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
-We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest.
+{\textcolor{blue}{The user selects a single town of interest and we add summary variables reflecting QoL metrics such as distance to local schools and city budget.
+Then, the user can}} in accordance with their interest.
 This avoids some of the pitfalls of dashboard design that can easily lead to user overload \citep{few}.

### 2022-10-02 SDSS Journal Paper/JDS supplemental items Resubmisson/JDS-Resubmisson.tex
Updates for resubmission to JDS

--- 2022-09-28 SDSS Journal Paper/Paper-Draft.tex
@@
 The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
-{\textcolor{blue}{The user selects a single town of interest and we add summary variables reflecting QoL metrics such as distance to local schools and city budget.
-Then, the user can}} in accordance with their interest.
+The user selects a single town of interest and we add summary variables reflecting QoL metrics such as distance to local schools and city budget.
+Then, the user can in accordance with their interest.
 This avoids some of the pitfalls of dashboard design that can easily lead to user overload \citep{few}.

### 2022-10-02 SDSS Journal Paper/JDS-Resubmisson.Rmd
Updates for resubmission to JDS

--- 2022-09-28 SDSS Journal Paper/Paper Draft.Rmd
@@
 The large set of publicly available data (primarily from \url{data.iowa.gov}) we have assembled is useful, but we must be careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture.
-\svp{The user selects a single town of interest and we add summary variables reflecting QoL metrics such as distance to local schools and city budget.
-Then, the user can} in accordance with their interest.
+The user selects a single town of interest and we add summary variables reflecting QoL metrics such as distance to local schools and city budget.
+Then, the user can in accordance with their interest.
 This avoids some of the pitfalls of dashboard design that can easily lead to user overload [@few].
